### 2024-11-25 paper/paper.qmd
Upload Initial Structure Files

---
title: "Microsoft Procurement"
subtitle: "SUBTITLE"
author: 
  - Deyi Kong
thanks: "Code and data are available at: [https://github.com/eeeee-cmd/Procurement/](https://github.com/eeeee-cmd/Procurement/)."
date: today
date-format: long
abstract: "This paper analizes the procurement supplier, Microsoft."
format: pdf
toc: true
toc-depth: 3
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Install Packages if not downloaded:
required_packages <- c(
  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra", "ggcorrplot", "modelsummary"
)
for (p in required_packages) {
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

# load libraries
library(tidyverse)
library(sf)
library(dplyr)
library(here)
library(arrow)
library(ggplot2)
library(janitor)
library(purrr)
library(knitr)
library(kableExtra)
library(ggcorrplot)
library(modelsummary)

# load data
cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
```

# Introduction

The main estimand is:

The structure of the paper is as follows: [@sec-data] outlines the data sources and variables considered, followed by the model setup in [@sec-modset] and justification in [@sec-modjust]. The results in [@sec-result] presents the key findings of the analysis, with a discussion on the implications. [@sec-discussion] then discusses potential limitations and suggestions for future research. [@sec-appx] provides additional detailed information about the data, model and methodology.

# Data {#sec-data}

## Overview

The data used in this analysis comes from a combination of publicly available procurement data [@IJF]. The analysis uses the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], `purrr` [@purrr], `sf` [@sf], and `here` [@here] for data manipulation. `ggplot2` [@ggplot], `ggcorrplot` [@ggcorrplot] and `kableExtra` [@kableExtra] for visualization. The dataset covers various polls conducted across multiple states, capturing the support for each major candidate—Donald Trump and Kamala Harris—along with detailed attributes of the polls.

## Measurement

The dataset: .

The measurement process begins with: .

## Data Cleaning

This paper first select and rename key variables from raw data to focus on relevant information. The key variables of interest in our analysis include Contract, StartDate, AwardDate, EndDate, Buyer, and Amount, with PreparatoryPhase and ContractTime being variables created.

More information on the data cleaning process can be found in @sec-appx.

## Outcome Variables

The main outcome variable of interest is: .

```{r}

```

## Predictor Variables

In this analysis, several key predictors were identified to evaluate proocurement data effectively.

# Model

## Model Set-Up {#sec-modset}

## Model Justification {#sec-modjust}

## Assumptions and Limitations

## Model Validation

# Results {#sec-result}

\newpage

# Discussion {#sec-discussion}

\newpage

# Appendix {#sec-appx}

## Data Cleaning Notes

## Additional Tables & Figures

## Methodology

## Idealized Methodology

## Idealized Survey

\newpage

# References

### 2024-11-25 scripts/00-simulate_data.R

@@ -1,0 +1,36 @@
+#### Preamble ####
+# Purpose: Simulate data 
+# Author: Deyi Kong
+# Date: November 24th, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `Procurement` rproj
+
+# load libraries
+library(tidyverse)
+library(arrow)
+
+# set seed for reproducability
+set.seed(123)
+
+# define possible values to simulate data
+contract <- c("Contract 1", "Contract 2", "Contract 3", "Contract 4", "Contract 5")
+buyer <- c("Buyer A", "Buyer B", "Buyer C", "Buyer D")
+n <- 100
+
+# generate random start and end dates for each poll
+start_dates <- sample(seq(as.Date("2024-01-01"), as.Date("2024-12-31"), by = "day"), n, replace = TRUE)
+end_dates <- start_dates + sample(1:30, n, replace = TRUE)
+
+# create a simulated data frame of polls with various attributes
+procurement_data <- data.frame(
+  contract = sample(contract, n, replace = TRUE),
+  buyer = sample(buyer, n, replace = TRUE),
+  start_date = start_dates,
+  end_date = end_dates
+)
+
+# write the cleaned and simulated polling data to a CSV file
+write_csv(procurement_data, here::here("data/simulated_data/simulated_data.csv"))
+write_parquet(procurement_data, here::here("data/simulated_data/simulated_data.parquet"))

### 2024-11-25 scripts/01-test_simulated_data.R

@@ -1,0 +1,39 @@
+#### Preamble ####
+# Purpose: This script reads in the simulated data, performs a series of validation tests
+# to ensure the data structure, column names, and values are correct, and reports the results.
+# Author: Deyi Kong
+# Date: November 24th, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse`, 'arrow', and `testthat` packages must be installed
+# Any other information needed? Make sure you are in the `Procurement` rproj
+
+# load libraries
+library(dplyr)
+library(arrow)
+library(testthat)
+
+# read the simulated data
+data <- read_parquet(here::here("data/simulated_data/simulated_data.parquet"))
+
+# create list to store test results
+test_results <- list()
+
+# test if the data structure is correct
+test_results$structure <- test_that("Data structure", {
+  expect_true(is.data.frame(data), info = "procurement_data is not a data frame.")
+  expect_equal(ncol(data), 25, info = "procurement_data does not have the correct number of columns.")
+  expect_equal(nrow(data), 100, info = "procurement_data does not have the correct number of rows.")
+})
+
+# test if the column names match the expected column names
+test_results$column_names <- test_that("Column names", {
+  expected_colnames <- c(
+    "contract", "buyer", "start_date", "end_date"
+  )
+  expect_equal(colnames(data), expected_colnames, info = "Column names do not match expected names.")
+})
+
+
+# print the test results
+print(test_results)

### 2024-11-25 scripts/02-download_data.R

@@ -1,0 +1,14 @@
+#### Preamble ####
+# Purpose: This script loads the raw procurement buyer data from a CSV file for further analysis.
+# Author: Deyi Kong
+# Date: November 24th, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `Procurement` rproj
+
+# load library
+library(readr)
+
+# Read the raw presidential polls data
+data <- read_csv(here::here("data/raw_data/procurement.csv"))

### 2024-11-25 scripts/03-clean_data.R

@@ -1,0 +1,73 @@
+#### Preamble ####
+# Purpose: This script reads and cleans raw procurement data,
+# performing various transformations, and outputs the cleaned data to a CSV file.
+# Author: Deyi Kong
+# Date: November 24th, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
+# Any other information needed? Make sure you are in the `Procurement` rproj
+
+# load libraries
+library(tidyverse)
+library(here)
+library(arrow)
+
+# read data
+data <- read_csv(here::here("data/raw_data/procurement.csv"))
+
+# Data cleaning by renaming and selecting specific columns, date formatting, filtering invalid entries
+cleaned_data <- tryCatch(
+  {
+    data %>%
+      select(
+        contract,
+        buyer,
+        supplier,
+        amount,
+        award_date,
+        start_date,
+        end_date
+      ) %>%
+      rename(
+        Supplier = supplier,
+        Contract = contract,
+        Buyer = buyer,
+        Amount = amount,
+        AwardDate = award_date,
+        StartDate = start_date,
+        EndDate = end_date
+      ) %>%
+      mutate(
+        # Specify the format of the date for proper parsing
+        StartDate = mdy(StartDate), # Use mdy() for MM/DD/YYYY format
+        AwardDate = mdy(AwardDate), 
+        EndDate = mdy(EndDate), 
+      ) %>%
+      filter(!is.na(StartDate), !is.na(AwardDate), !is.na(EndDate))
+  }, # error handling to mitigate any cleaning issues
+  error = function(e) {
+    message("An error occurred during data cleaning: ", e)
+    NULL
+  }
+)
+
+
+#unique(cleaned_data$Contract)
+
+# add preparatory phase and contract total days
+cleaned_data <- cleaned_data %>%
+  mutate(PreparatoryPhase = as.numeric(StartDate - AwardDate),
+         ContractTime = as.numeric(EndDate - StartDate))
+
+# MIGHT GOING TO ADD SOME CLEANING PROCESS ABOUT CONTRACT CATEGORY
+
+# save cleaned data as new csv and parquet if successful, print error if not
+if (!is.null(cleaned_data)) {
+  write_csv(cleaned_data, here::here("data/analysis_data/procurement_cleaned.csv"))
[…]

### 2024-11-25 scripts/04-test_analysis_data.R

@@ -1,0 +1,81 @@
+#### Preamble ####
+# Purpose: This script performs a series of validation tests on procurement data.
+# Author: Deyi Kong
+# Date: November 24th, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse`, `here`, `arrow`, and `testthat` packages must be installed.
+# Any other information needed? Make sure you are in the `Procurement` rproj
+
+# load libraries
+library(tidyverse)
+library(here)
+library(arrow)
+library(testthat)
+
+# load data
+cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
+
+# create list for results
+test_results <- list()
+
+# test structure of data
+test_results$structure <- test_that("Data structure", {
+  expect_true(is.data.frame(cleaned_data), info = "clean_data is not a data frame.")
+  expect_equal(ncol(cleaned_data), 18, info = "clean_data does not have the correct number of columns.")
+  expect_equal(nrow(cleaned_data), 15829, info = "clean_data does not have the correct number of rows.")
+})
+
+# test if column names match expected names
+test_results$column_names_test <- test_that("Column names", {
+  expected_colnames <- c(
+    "poll_id", "pollster_id", "Pollster", "Party", "sponsors",
+    "NumericGrade", "PollScore", "Methodology",
+    "State", "StartDate", "EndDate",
+    "SampleSize", "Population", "ElectionDate",
+    "Stage", "CandidateName", "Percentage", "PollRecency"
+  )
+  expect_equal(names(cleaned_data), expected_colnames, info = "Column names do not match expected names.")
+})
+
+# test if certain columns contain NA values
+test_results$na_values_test <- test_that("NA values", {
+  expect_false(any(is.na(cleaned_data$Percentage)), info = "Percentage contains NA values.")
+  expect_false(any(is.na(cleaned_data$CandidateName)), info = "CandidateName contains NA values.")
+})
+
+# test if NumericGrade and PollScore values are within expected range
+test_results$numeric_grade_pollscore_test <- test_that("Numeric grade and PollScore values", {
+  expect_true(all(cleaned_data$NumericGrade >= 0 & cleaned_data$NumericGrade <= 3, na.rm = TRUE),
+    info = "NumericGrade values are out of expected range (0 to 3)."
+  )
[…]

### 2024-11-25 scripts/05-exploratory_data_analysis.R

@@ -1,0 +1,41 @@
+#### Preamble ####
+# Purpose: Perform exploratory data analysis on cleaned data
+# Author: Deyi Kong
+# Date: November 24th, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse`, 'janitor', 'here', and 'reshape2' packages
+# must be installed
+# Any other information needed? Make sure you are in the `Procurement` rproj
+
+# load libraries
+library(tidyverse)
+library(janitor)
+library(here)
+library(reshape2)
+library(arrow)
+
+# read cleaned data
+cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet")) |>
+  clean_names()
+
+# quick overview of data structure
+glimpse(cleaned_data)
+
+# print first few rows of data
+head(cleaned_data)
+
+# print last few rows of data
+tail(cleaned_data)
+
+# randomly sample 6 rows
+cleaned_data |>
+  slice_sample(n = 6)
+
+# summary of the dataset
+summary(cleaned_data)
+
+# number of NA values in each column
+missing_data <- cleaned_data |>
+  summarise(across(everything(), ~ sum(is.na(.))))
+missing_data

### 2024-11-25 scripts/06-model_data.R

@@ -1,0 +1,18 @@
+#### Preamble ####
+# Purpose: Run a linear regression model to predict 
+# Author: Deyi Kong
+# Date: November 24th, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` and 'here' packages must be installed
+# Any other information needed? Make sure you are in the `Procurement` rproj
+
+# load libraries
+library(tidyverse)
+
+
+
+# save modified data for the model as a CSV + Parquet and save models as RDS
+write_csv(cleaned_data, here::here("data/analysis_data/data.csv"))
+write_parquet(cleaned_data, here::here("data/analysis_data/data.parquet"))
+saveRDS(lm_model_1, here::here("models/lm_model_1.rds"))

### 2024-11-25 scripts/07-validation.R

@@ -1,0 +1,22 @@
+#### Preamble ####
+# Purpose: Validate the regression models through out of sample testing and RMSE
+# Author: Deyi Kong
+# Date: November 24th, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse`, 'caret', and 'here' packages must be installed
+# Any other information needed? Make sure you are in the `Procurement` rproj
+
+
+# load libraries
+library(tidyverse)
+library(here)
+library(caret)
+
+# load data
+cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
+
+# data processing
+
+
+write_parquet(rmse_results, here::here("data/analysis_data/rmse_results.parquet"))

### 2024-11-25 scripts/00-simulate_data.R
Update Simulation

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulate data 
 # Author: Deyi Kong
-# Date: November 24th, 2024
+# Date: November 25th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
@@ -15,22 +15,30 @@ library(arrow)
 set.seed(123)
 
 # define possible values to simulate data
-contract <- c("Contract 1", "Contract 2", "Contract 3", "Contract 4", "Contract 5")
-buyer <- c("Buyer A", "Buyer B", "Buyer C", "Buyer D")
+contract <- c("License/Maintenance fees", "Computer services", "Consultants", "Application software", "Production and operations")
+buyer <- c("Employment and Social Development Canada", "Global Affairs Canada", "National Defence", "Fisheries and Oceans Canada")
+supplier <- c("MICROSOFT CANADA INC.", "MICROSOFT CORPORATION", "MICROSOFT CANADA CO.")
 n <- 100
 
 # generate random start and end dates for each poll
-start_dates <- sample(seq(as.Date("2024-01-01"), as.Date("2024-12-31"), by = "day"), n, replace = TRUE)
-end_dates <- start_dates + sample(1:30, n, replace = TRUE)
+award_dates <- sample(seq(as.Date("2024-01-01"), as.Date("2024-12-31"), by = "day"), n, replace = TRUE)
+start_dates <- award_dates + sample(1:30, n, replace = TRUE)
+end_dates <- start_dates + sample(1:300, n, replace = TRUE)
 
 # create a simulated data frame of polls with various attributes
 procurement_data <- data.frame(
   contract = sample(contract, n, replace = TRUE),
   buyer = sample(buyer, n, replace = TRUE),
+  supplier = sample(supplier, n, replace = TRUE),
+  amount = sample(40000:450000, n, replace = TRUE),
+  award_date = award_dates,
   start_date = start_dates,
-  end_date = end_dates
+  end_date = end_dates,
+  preparatory_phase = as.numeric(start_dates - award_dates),
+  contract_days = as.numeric(end_dates - start_dates)
 )
 
 # write the cleaned and simulated polling data to a CSV file
 write_csv(procurement_data, here::here("data/simulated_data/simulated_data.csv"))
 write_parquet(procurement_data, here::here("data/simulated_data/simulated_data.parquet"))
+

### 2024-11-25 scripts/01-test_simulated_data.R
Update Test Simulation

@@ -2,7 +2,7 @@
 # Purpose: This script reads in the simulated data, performs a series of validation tests
 # to ensure the data structure, column names, and values are correct, and reports the results.
 # Author: Deyi Kong
-# Date: November 24th, 2024
+# Date: November 25th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, 'arrow', and `testthat` packages must be installed
@@ -22,18 +22,37 @@ test_results <- list()
 # test if the data structure is correct
 test_results$structure <- test_that("Data structure", {
   expect_true(is.data.frame(data), info = "procurement_data is not a data frame.")
-  expect_equal(ncol(data), 25, info = "procurement_data does not have the correct number of columns.")
+  expect_equal(ncol(data), 9, info = "procurement_data does not have the correct number of columns.")
   expect_equal(nrow(data), 100, info = "procurement_data does not have the correct number of rows.")
 })
 
 # test if the column names match the expected column names
 test_results$column_names <- test_that("Column names", {
   expected_colnames <- c(
-    "contract", "buyer", "start_date", "end_date"
+    "contract", "buyer", "supplier", "amount", "award_date", "start_date", "end_date", "preparatory_phase", "contract_days"
   )
   expect_equal(colnames(data), expected_colnames, info = "Column names do not match expected names.")
 })
 
+# test if the amount values are within valid range (40000 to 450000)
+test_results$amount <- test_that("amount", {
+  expect_true(all(data$amount >= 40000 & data$amount <= 450000), info = "Contract amount values are out of range (40000 to 450000).")
+})
+
+# test if the preparatory phase (days) is within the valid range (1 to 30)
+test_results$preparatory_phase <- test_that("Preparatory Phase", {
+  expect_true(all(data$preparatory_phase >= 1 & data$preparatory_phase <= 30), info = "numeric_grade values are out of range (1 to 30).")
+})
+
+# test if the award dates are logically before the start dates
+test_results$award_logic <- test_that("Award Date logic", {
+  expect_true(all(data$award_date < data$start_date), info = "Some award_date are not before their corresponding start_dates")
+})
+
+# test if the start dates are logically before the end dates
+test_results$date_logic <- test_that("Date logic", {
+  expect_true(all(data$start_date < data$end_date), info = "Some start_dates are not before their corresponding end_dates.")
+})
 
 # print the test results
 print(test_results)

### 2024-11-25 scripts/03-clean_data.R
Update Data Cleaning Script

@@ -2,7 +2,7 @@
 # Purpose: This script reads and cleans raw procurement data,
 # performing various transformations, and outputs the cleaned data to a CSV file.
 # Author: Deyi Kong
-# Date: November 24th, 2024
+# Date: November 25th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
@@ -13,10 +13,13 @@ library(tidyverse)
 library(here)
 library(arrow)
 
-# read data
+# Read data
 data <- read_csv(here::here("data/raw_data/procurement.csv"))
 
-# Data cleaning by renaming and selecting specific columns, date formatting, filtering invalid entries
+# Check if the contract region are all federal, then we will exclude it from the cleaned data
+table(data$region)
+
+# Data cleasupplier# Data cleaning by renaming and selecting specific columns, date formatting, filtering invalid entries
 cleaned_data <- tryCatch(
   {
     data %>%
@@ -39,12 +42,19 @@ cleaned_data <- tryCatch(
         EndDate = end_date
       ) %>%
       mutate(
+        # To replace missing (NA) start date with the value of awarded date column
+        StartDate = ifelse(is.na(StartDate), AwardDate, StartDate)
+      ) %>%
+      mutate(
         # Specify the format of the date for proper parsing
         StartDate = mdy(StartDate), # Use mdy() for MM/DD/YYYY format
         AwardDate = mdy(AwardDate), 
-        EndDate = mdy(EndDate), 
+        EndDate = mdy(EndDate)
       ) %>%
-      filter(!is.na(StartDate), !is.na(AwardDate), !is.na(EndDate))
+      mutate(
+        # Change the $format of dollar amount into numeric numbers
+        Amount = as.numeric(gsub("[\\$,]", "", Amount))
+      )
   }, # error handling to mitigate any cleaning issues
   error = function(e) {
     message("An error occurred during data cleaning: ", e)
@@ -52,17 +62,62 @@ cleaned_data <- tryCatch(
   }
 )
 
-
-#unique(cleaned_data$Contract)
-
-# add preparatory phase and contract total days
+# Add preparatory phase and contract total days
 cleaned_data <- cleaned_data %>%
[…]

### 2024-11-26 scripts/03-clean_data.R
Update Analysis Data

@@ -16,8 +16,8 @@ library(arrow)
 # Read data
 data <- read_csv(here::here("data/raw_data/procurement.csv"))
 
-# Check if the contract region are all federal, then we will exclude it from the cleaned data
-table(data$region)
+# # Check if the contract region are all federal, then we will exclude it from the cleaned data
+# table(data$region)
 
 # Data cleasupplier# Data cleaning by renaming and selecting specific columns, date formatting, filtering invalid entries
 cleaned_data <- tryCatch(
@@ -43,7 +43,10 @@ cleaned_data <- tryCatch(
       ) %>%
       mutate(
         # To replace missing (NA) start date with the value of awarded date column
-        StartDate = ifelse(is.na(StartDate), AwardDate, StartDate)
+        StartDate = case_when(
+          is.na(StartDate) ~ AwardDate,
+          TRUE ~ StartDate
+        )
       ) %>%
       mutate(
         # Specify the format of the date for proper parsing
@@ -54,7 +57,9 @@ cleaned_data <- tryCatch(
       mutate(
         # Change the $format of dollar amount into numeric numbers
         Amount = as.numeric(gsub("[\\$,]", "", Amount))
-      )
+      ) %>%
+      # Drop contracts where EndDate is before StartDate or AwardDate due to error with unknown reason
+      filter(!(EndDate <= StartDate | EndDate <= AwardDate))
   }, # error handling to mitigate any cleaning issues
   error = function(e) {
     message("An error occurred during data cleaning: ", e)

### 2024-11-26 scripts/04-test_analysis_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: This script performs a series of validation tests on procurement data.
 # Author: Deyi Kong
-# Date: November 24th, 2024
+# Date: November 25th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, `here`, `arrow`, and `testthat` packages must be installed.
@@ -22,43 +22,24 @@ test_results <- list()
 # test structure of data
 test_results$structure <- test_that("Data structure", {
   expect_true(is.data.frame(cleaned_data), info = "clean_data is not a data frame.")
-  expect_equal(ncol(cleaned_data), 18, info = "clean_data does not have the correct number of columns.")
-  expect_equal(nrow(cleaned_data), 15829, info = "clean_data does not have the correct number of rows.")
+  expect_equal(ncol(cleaned_data), 9, info = "clean_data does not have the correct number of columns.")
+  expect_equal(nrow(cleaned_data), 950, info = "clean_data does not have the correct number of rows.")
 })
 
 # test if column names match expected names
 test_results$column_names_test <- test_that("Column names", {
   expected_colnames <- c(
-    "poll_id", "pollster_id", "Pollster", "Party", "sponsors",
-    "NumericGrade", "PollScore", "Methodology",
-    "State", "StartDate", "EndDate",
-    "SampleSize", "Population", "ElectionDate",
-    "Stage", "CandidateName", "Percentage", "PollRecency"
+     "Contract", "Buyer", "Supplier", "Amount", "AwardDate", "StartDate", "EndDate", "PreparatoryPhase", "ContractDays"
   )
   expect_equal(names(cleaned_data), expected_colnames, info = "Column names do not match expected names.")
 })
 
 # test if certain columns contain NA values
 test_results$na_values_test <- test_that("NA values", {
-  expect_false(any(is.na(cleaned_data$Percentage)), info = "Percentage contains NA values.")
-  expect_false(any(is.na(cleaned_data$CandidateName)), info = "CandidateName contains NA values.")
-})
-
-# test if NumericGrade and PollScore values are within expected range
-test_results$numeric_grade_pollscore_test <- test_that("Numeric grade and PollScore values", {
-  expect_true(all(cleaned_data$NumericGrade >= 0 & cleaned_data$NumericGrade <= 3, na.rm = TRUE),
[…]

### 2024-11-26 paper/paper.qmd
Update date cleaning and paper file

@@
 ---
-title: "Microsoft Procurement"
-subtitle: "SUBTITLE"
+title: "Procurement Dynamics: Analyzing Federal Contract Awards to Microsoft"
+subtitle: "A Comprehensive Study of Microsoft’s Role in Canadian Government Procurement (NEED a better SUBTITLE being more specific)"
 author: 
@@
 date-format: long
-abstract: "This paper analizes the procurement supplier, Microsoft."
+abstract: "This paper analyzes the federal procurement contracts awarded to Microsoft, focusing on trends in award values, the categories of the contract, and the relationships between Microsoft and government buyers.
+Using IJF procurement dataset, we found that Microsoft engages significant role in IT infrastructure and modernization efforts.
+The analysis reveals that Microsoft dominates high-value contracts and demonstrates consistent partnerships with key government departments.
+By examining these procurement dynamics, this paper providing insights into the broader implications of tech-industry participation in government procurement."
 format: pdf
@@
 
-The main estimand is:
+The main estimand is the dollar amount of federal contracts.
 
@@
 ```{r}
+#| label: fig-amount
+#| fig-cap: The histogram shows the distribution of the total amount($million) of each year's federal contract (since 2020).
+[An increasing trend may reflect the growing importance of technology in government operations, including software, cloud solutions, and IT infrastructure.]
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Extract year from AwardDate and filter the year from 2020 to 2024
+year_data <- cleaned_data %>%
+  mutate(
+    Year = as.integer(format(as.Date(AwardDate), "%Y"))) %>%
+  filter(
+    Year != 2019
+  )
+
+# Sum the total amount group by year
+yearly_amount <- year_data %>%
+  group_by(Year) %>%
+  summarize(TotalAmount = sum(Amount, na.rm = TRUE)) %>%
+  mutate(TotalAmount = round(TotalAmount / 1000000)) # Convert to millions
+
+# Plot the amount in each year
+ggplot(yearly_amount, aes(x = as.factor(Year), y = TotalAmount)) +
+  geom_bar(stat = "identity", fill = "darkorange", color = "black", alpha = 0.7) +
+  labs(
+    title = "Total Contract Amounts by Year for Microsoft",
+    x = "Year",
+    y = "Total Amount ($million)"
+  ) +
+  theme_minimal() +
+  theme(axis.text.y = element_text(size = 10), axis.text.x = element_text(size = 10))
 
@@
 In this analysis, several key predictors were identified to evaluate proocurement data effectively.
+
+```{r}
+#| label: fig-preparatory
+#| fig-cap: The histogram shows the distribution of preparatory period, which demonstrates the days between start and awarded date.
+A high concentration around 0 indicates rapid procurement processes or some emergency software services.
+There are a few outlying contracts with a mobilization period of over one year.
+These long mobilization periods suggest issues such as delays in the procurement process, a long negotiation period, or administrative approvals.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# create plot for Distribution of Poll Scores by candidate
+ggplot(cleaned_data, aes(x = PhaseDays)) +
+  geom_histogram(
+    bins = 30,
+    fill = "salmon",
+    color = "black",
+    alpha = 0.7
+  ) +
+  labs(
+    x = "Preparatory Phase (day)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+```
+
+```{r}
+#| label: fig-contract
+#| fig-cap: The histogram shows the distribution of performance period, which demonstrates the days between start and end date.
+The right skewed distribution indicates that the majority of contract duration is within a year.
+An over 5 years performance period implies a massive project that requires extensive preparation.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# create plot for Distribution of Poll Scores by candidate
+ggplot(cleaned_data, aes(x = ContractDays)) +
+  geom_histogram(
+    bins = 15,
+    fill = "steelblue",
+    color = "black",
+    alpha = 0.7
+  ) +
+  labs(
+    x = "Contract Performance Phase (day)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+```
 
@@
 
-## Idealized Survey
-
 \newpage

### 2024-11-26 scripts/03-clean_data.R

@@ -70,7 +70,9 @@ cleaned_data <- tryCatch(
 # Add preparatory phase and contract total days
 cleaned_data <- cleaned_data %>%
   mutate(PreparatoryPhase = as.numeric(StartDate - AwardDate),
-         ContractDays = as.numeric(EndDate - StartDate))
+         ContractDays = as.numeric(EndDate - StartDate),
+         PhaseDays = abs(PreparatoryPhase)
+         )
 
 # # Summary all the variables
 # sort(table(cleaned_data$Contract), decreasing = T)
@@ -118,7 +120,12 @@ clean_supplier_names <- function(supplier) {
 cleaned_data <- cleaned_data %>%
   mutate(Supplier = clean_supplier_names(Supplier))
 
-sort(table(cleaned_data$Supplier), decreasing = T)
+## Check variables
+# sort(table(cleaned_data$Supplier), decreasing = T)
+# sort(table(cleaned_data$PreparatoryPhase), decreasing = T)
+# summary(cleaned_data$PreparatoryPhase)
+# sort(table(cleaned_data$PhaseDays), decreasing = T)
+# summary(cleaned_data$PhaseDays)
 
 # MIGHT GOING TO ADD SOME CLEANING PROCESS ABOUT CONTRACT CATEGORY
 

### 2024-11-26 scripts/04-test_analysis_data.R

@@ -22,14 +22,14 @@ test_results <- list()
 # test structure of data
 test_results$structure <- test_that("Data structure", {
   expect_true(is.data.frame(cleaned_data), info = "clean_data is not a data frame.")
-  expect_equal(ncol(cleaned_data), 9, info = "clean_data does not have the correct number of columns.")
+  expect_equal(ncol(cleaned_data), 10, info = "clean_data does not have the correct number of columns.")
   expect_equal(nrow(cleaned_data), 950, info = "clean_data does not have the correct number of rows.")
 })
 
 # test if column names match expected names
 test_results$column_names_test <- test_that("Column names", {
   expected_colnames <- c(
-     "Contract", "Buyer", "Supplier", "Amount", "AwardDate", "StartDate", "EndDate", "PreparatoryPhase", "ContractDays"
+     "Contract", "Buyer", "Supplier", "Amount", "AwardDate", "StartDate", "EndDate", "PreparatoryPhase", "ContractDays", "PhaseDays"
   )
   expect_equal(names(cleaned_data), expected_colnames, info = "Column names do not match expected names.")
 })

### 2024-11-26 paper/paper.qmd
Update paper and data cleaning

@@
 title: "Procurement Dynamics: Analyzing Federal Contract Awards to Microsoft"
-subtitle: "A Comprehensive Study of Microsoft’s Role in Canadian Government Procurement (NEED a better SUBTITLE being more specific)"
+subtitle: "Trends and Insights of Microsoft’s Future in Canadian Government from 2020"
 author: 
@@
 date-format: long
-abstract: "This paper analyzes the federal procurement contracts awarded to Microsoft, focusing on trends in award values, the categories of the contract, and the relationships between Microsoft and government buyers.
-Using IJF procurement dataset, we found that Microsoft engages significant role in IT infrastructure and modernization efforts.
+abstract: "This paper analyzes the federal procurement contracts awarded to Microsoft, focusing on trends in award values, contract categories, and the relationships between buyers and Microsoft.
 The analysis reveals that Microsoft dominates high-value contracts and demonstrates consistent partnerships with key government departments.
+This paper utilizes a linear model to predict the future contract amount trend for Microsoft.
 By examining these procurement dynamics, this paper providing insights into the broader implications of tech-industry participation in government procurement."
@@
 
-The main estimand is the dollar amount of federal contracts.
+Government procurement serves as a critical avenue for public sector innovation, economic stimulus, and the fulfillment of strategic objectives.
+Within this landscape, the relationship between major technology providers and governmental bodies has become increasingly prominent.
+Among these, Microsoft has emerged as a pivotal partner in driving IT modernization and infrastructure development.
+Despite the growing reliance on tech giants, limited research has comprehensively examined the dynamics of such partnerships, particularly in the context of federal procurement contracts.
 
-The structure of the paper is as follows: [@sec-data] outlines the data sources and variables considered, followed by the model setup in [@sec-modset] and justification in [@sec-modjust].
+This paper addresses this gap by conducting a detailed analysis of federal procurement contracts awarded to Microsoft between 2020 and September 2024.
+Using publicly available data from the Investigative Journalism Foundation, the study explores trends in contract values, contract types, and Microsoft’s relationships with key government buyers.
+Emphasis is placed on identifying patterns in contract allocation and the implications of Microsoft's dominance in high-value contracts.
+
+The main estimand in this study is the amount of federal procurement contracts awarded to Microsoft.
+This metric is analyzed across multiple dimensions, including contract categories, durations, and partnerships, to quantify Microsoft's role in federal procurement comprehensively.
+The findings reveal that Microsoft plays a dominant role in IT infrastructure and modernization, securing the majority of high-value contracts and establishing consistent partnerships with departments such as Employment and Social Development Canada and National Defence.
+These insights highlight the broader implications of the private tech sector’s influence on public sector procurement, raising questions about competitiveness, dependency, and strategic alignment.
+
+This study contributes to understand the procurement landscape, providing valuable information for policymakers, industry stakeholders, and academics.
+It underscores the importance of transparency, strategic planning, and balancing innovation with equity in government contracts.
+
+The structure of the paper is as follows: [@sec-data] outlines the data sources and variables considered, followed by the model setup in [@sec-modset] and justification.
 The results in [@sec-result] presents the key findings of the analysis, with a discussion on the implications.
@@
 
-The data used in this analysis comes from a combination of publicly available procurement data [@IJF].
-The analysis uses the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], `purrr` [@purrr], `sf` [@sf], and `here` [@here] for data manipulation.
-`ggplot2` [@ggplot], `ggcorrplot` [@ggcorrplot] and `kableExtra` [@kableExtra] for visualization.
-The dataset covers various polls conducted across multiple states, capturing the support for each major candidate—Donald Trump and Kamala Harris—along with detailed attributes of the polls.
+The data used in this analysis comes from publicly available procurement data [@IJF].
+The analysis uses the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], `purrr` [@purrr], and `here` [@here] for data manipulation.
+`ggplot2` [@ggplot] and `kableExtra` [@kableExtra] for visualization.
+The dataset covers the information of chosen Federal Procurement Supplier - Microsoft - capturing the contracts from 2020 to 2024 (up to September).
 
@@
 
-The dataset: .
+The dataset used in this analysis originates from Canadian governments of all levels post solicitations for companies to search for business opportunities, as documented in @ProcurementIntro, then collects data from each of these platforms and unifies into standard format.
 
-The measurement process begins with: .
+The measurement process delegates contracts routed through Public Service and Procurement Canada (PSPC), and blanket disclosures organized by Treasury Board of Canada Secretariat (TBS) to understand the federal procurement.
+The methodologies of federal solicitations are either competitive or non-competitive.
+Competitive contracts accept bids from all or some subset of suppliers.
+Non-competitive contracts do not, citing certain exceptions, including a “pressing emergency in which delay would be injurious to the public interest,” or when the estimated cost to the buyer is below certain thresholds defined for some goods and services.
+
+More information about solicitation method can be found in @sec-appx.
+
+\newpage
 
@@
 
-This paper first select and rename key variables from raw data to focus on relevant information.
-The key variables of interest in our analysis include Contract, StartDate, AwardDate, EndDate, Buyer, and Amount, with PreparatoryPhase and ContractTime being variables created.
+After imported the raw dataset using `read_csv` function, I first select specific columns to focus on relevant information towards this study by omitting redundant information, such as the contract link and same value of region.
+Then I rename key variables from raw data for clarity.
+The key variables of interest in this analysis include Contract, StartDate, AwardDate, EndDate, Buyer, and Amount.
 
-More information on the data cleaning process can be found in @sec-appx.
+For any missing data (NA) of StartDate, I infer the StartDate from the AwardDate to replace the missing value as the value of AwardDate.
+Then I standardize the date format of all date related datas for proper parsing by using `mdy()` function and change the dollar format of amount into numeric numbers for future calculation purposes.
+Due to the error with unknown reason that EndDate is before StartDate or AwardDate, those entries were excluded from this paper.
+
+PreparatoryPhase, PhaseDays, and ContractDays variables were created for the purpose of analyzing preparatory phase and performance phase corresponding to each contract.
+PreparatoryPhase calculates the days from AwardDate to StartDate.
+ContractDays calculates the days from StartDate to EndDate.
+PhaseDays take the absolute value of PreparatoryPhase because of the minor effect of either prepare before contract being awarded or prepare after contract being awarded.
+
+This dataset contains only two entries for 2019 federal procurement contracts awarded to Microsoft.
+So, I filter out the contracts before 2020 to focus on the records between 2020 and September 2024 as desired.
+Then I check the character type variables, Supplier and Contract, using `table()` function for the purpose of manipulating input error.
+I standardize the supplier's name which combine the similar name due to capitalization or some notation differences, and format them with first letter of each word capitalized.
+
+After completing the cleaning, we saved the final dataset in both Parquet and CSV formats for later analysis.
 
@@
 
-The main outcome variable of interest is: .
+The main outcome variable of interest is the amount of federal procurement contracts awarded to Microsoft.
+This variable is need to represent the financial magnitude of Microsoft with the federal government and measurement of evaluating procurement dynamics.
+[@fig-amount] illustrates the annual total amounts of federal procurement contracts awarded to Microsoft from 2020 to 2024 (up to September).
+The histogram reveals a generally increasing trend of total value of each year's contract.
+This may reflect the growing of technology in government operations, including software, cloud solutions, and IT infrastructure.
+The amount in 2023 doubles the amount in 2021, which suggests a heightened investment in federal procurement funding and large-scale projects.
 
@@
 #| fig-cap: The histogram shows the distribution of the total amount($million) of each year's federal contract (since 2020).
-[An increasing trend may reflect the growing importance of technology in government operations, including software, cloud solutions, and IT infrastructure.]
 #| echo: false
@@
   mutate(
-    Year = as.integer(format(as.Date(AwardDate), "%Y"))) %>%
-  filter(
-    Year != 2019
+    Year = as.integer(format(as.Date(AwardDate), "%Y"))
     )
@@
 
-In this analysis, several key predictors were identified to evaluate proocurement data effectively.
+In this analysis, several key predictors were identified to evaluate procurement data effectively.
+The first variable, **Contract**, is a character-type variable that identifies each procurement contract awarded to Microsoft.
+[@tbl-contract] [SOME MODIFICATION OF CONTRACT CATEGORIES CLEANING PROCESS MAY APPLY]
+
+```{r}
+#| label: tbl-contract
+#| tbl-cap: The table summarizes the contract type which has over 25 contracts with the total dollar amount, and average dollar amount per contract. 
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Summarize the data and filter buyers with counts >= 25
+summary_table <- cleaned_data %>%
+  group_by(Contract) %>%
+  summarise(
+    Count = n(), 
+    TotalAmount = sum(Amount),  # Calculate the total contract value
+    AvgAmount = mean(Amount)    # Calculate the average contract value
+  ) %>%
+  filter(Count > 25) %>%
+  arrange(desc(Count))  # Arrange in descending order based on Count
+
+# Format the TotalAmount column with dollar format
+summary_table$TotalAmount <- scales::label_dollar()(summary_table$TotalAmount)
+summary_table$AvgAmount <- scales::label_dollar()(summary_table$AvgAmount)
+
+# Display the summary table using kableExtra for nicer formatting
+summary_table %>%
+  kable(col.names = c("Contract Type", "Counts", "Total Amount", "Average")) %>%
+  kable_styling(full_width = F, position = "center", bootstrap_options = c("striped", "hover")) %>%
+  column_spec(1, width = "22em")  # Set column width to wrap text
+
+```
+
+**Buyer** is a character-type variable representing the federal government department or agency who initiate and award the contract.
+[@tbl-buyer] shows the top 10 buyer's name who has over 25 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors.
+The table highlights that Employment and Social Development Canada (ESDC) and National Defence account over 50% of Microsoft's total awards.
+[CALCULATED 52.02%, BUT THINKING IF THIS SHOULD BE ADDED TO THE TABLE]
+
+```{r}
+#| label: tbl-buyer
+#| tbl-cap: The table summarizes top 10 buyer's name who has over 25 contracts with Microsoft in total with contract count, total dollar amount, and average dollar amount per contract. 
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Summarize the data and filter buyers with counts >= 25
+summary_table <- cleaned_data %>%
+  group_by(Buyer) %>%
+  summarise(
+    Count = n(), 
+    TotalAmount = sum(Amount),  # Calculate the total contract value
+    AvgAmount = mean(Amount)    # Calculate the average contract value
+  ) %>%
+  filter(Count > 25) %>%
+  arrange(desc(Count))  # Arrange in descending order based on Count
+
+# Format the TotalAmount column with dollar format
+summary_table$TotalAmount <- scales::label_dollar()(summary_table$TotalAmount)
+summary_table$AvgAmount <- scales::label_dollar()(summary_table$AvgAmount)
+
+# Display the summary table using kableExtra for nicer formatting
+summary_table %>%
+  kable(col.names = c("Buyer", "Contract", "Total Amount", "Average")) %>%
+  kable_styling(full_width = F, position = "center", bootstrap_options = c("striped", "hover")) %>%
+  column_spec(1, width = "22em")  # Set column width to wrap text
+
+```
+
+**Supplier** is a character-type variable includes all the affiliated subsidiary companies of Microsoft.
+[@tbl-supplier] shows the top 5 buyer's name who has awarded over 25 contracts.
+[DELETE THIS VARIABLE?
+OR MAYBE MOVE IT TO APPDX.]
+
+```{r}
+#| label: tbl-supplier
+#| tbl-cap: The table summarizes the main Microsoft supplier's name who has over 25 contracts with buyers in total with contract count, total dollar amount, and average dollar amount per contract. 
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Summarize the data and filter buyers with counts >= 25
+summary_table <- cleaned_data %>%
+  group_by(Supplier) %>%
+  summarise(
+    Count = n(), 
+    TotalAmount = sum(Amount),  # Calculate the total contract value
+    AvgAmount = mean(Amount)    # Calculate the average contract value
+  ) %>%
+  filter(Count > 25) %>%
+  arrange(desc(Count))  # Arrange in descending order based on Count
+
+# Format the TotalAmount column with dollar format
+summary_table$TotalAmount <- scales::label_dollar()(summary_table$TotalAmount)
+summary_table$AvgAmount <- scales::label_dollar()(summary_table$AvgAmount)
+
+# Display the summary table using kableExtra for nicer formatting
+summary_table %>%
+  kable(col.names = c("Suppllier", "Contract", "Total Amount", "Average")) %>%
+  kable_styling(full_width = F, position = "center", bootstrap_options = c("striped", "hover"))
+
+```
+
+This dataset contains a series of date-type variables capturing key milestones in the procurement process.
+**Award Date** is the date when the contract was officially granted to Microsoft.
+**Start Date** is the date when the contracted services or goods provision begins.
+**End Date** is the date when the contract is completed or terminated.
+These variables are essential for analyzing procurement timelines and trends over the study period.
+
+**PhaseDays** is a numeric variable demonstrates the days between AwardDate and StartDate.
+In [@fig-preparatory], there is a high concentration around 0 indicates rapid procurement processes or emergency software services.
+[@fig-preparatory] also shows a few outlying contracts with a mobilization period of over one year.
+These long mobilization periods reflects issues such as delays in the procurement process, a long negotiation period, or administrative approvals.
 
@@
 #| fig-cap: The histogram shows the distribution of preparatory period, which demonstrates the days between start and awarded date.
-A high concentration around 0 indicates rapid procurement processes or some emergency software services.
-There are a few outlying contracts with a mobilization period of over one year.
-These long mobilization periods suggest issues such as delays in the procurement process, a long negotiation period, or administrative approvals.
 #| echo: false
@@
 
+**ContractDays** is a numeric variable which is the days between StartDate and EndDate.
+[@fig-contract] shows the distribution of performance period.
+The right skewed distribution in [@fig-contract] indicates that the majority of contract duration is within a year.
+There is an over 5 years performance period in [@fig-contract] implies a massive project that requires extensive preparation.
+
 ```{r}
@@
 The right skewed distribution indicates that the majority of contract duration is within a year.
-An over 5 years performance period implies a massive project that requires extensive preparation.
 #| echo: false
@@
 
+\newpage
+
 # Model
@@
 
-## Model Justification {#sec-modjust}
++ Model Justification
 
@@
 
+\newpage
+
 # Results {#sec-result}
+
+```{r}
+#| label: fig-date
+#| fig-cap: The heatmap ....
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Create a new variable for "four-month periods"
+data <- cleaned_data %>%
+  mutate(
+    Year = year(AwardDate),
+    Month = month(AwardDate),
+    Period = case_when(
+      Month %in% 1:4 ~ "Jan-Apr",
+      Month %in% 5:8 ~ "May-Aug",
+      Month %in% 9:12 ~ "Sep-Dec"
+    ),
+    YearPeriod = paste(Year, Period, sep = "-")  # Combine Year and Period
+  )
+
+# Count the number of contracts in each four-month period
+contract_counts <- data %>%
+  group_by(YearPeriod) %>%
+  summarise(Count = n())
+
+# Heatmap plot
+ggplot(contract_counts, aes(x = YearPeriod, y = Count, fill = Count)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "red") +
+  labs(x = "Year-Period", y = "Frequency (Contract Count)", title = "Heatmap of Contracts by Four-Month Periods") +
+  theme_minimal() +
+  theme(axis.text.y = element_text(angle = 0))  # Keep y-axis labels readable
+```
+
+```{r}
+#| label: fig-mobilization
+#| fig-cap: The scatter plot shows the distribution of performance period, which demonstrates the days between start and end date.
+The right skewed distribution indicates that the majority of contract duration is within a year.
+An over 5 years performance period implies a massive project that requires extensive preparation.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| 
+ggplot(data, aes(x = AwardDate, y = StartDate)) +
+  geom_point(aes(color = Contract)) +
+  labs(x = "Award Date", y = "Start Date", title = "Award Date vs Start Date") +
+  theme(legend.position = "none")  # This removes the legend
+```
+
+```{r}
+#| label: fig-performance
+#| fig-cap: The scatter plot shows the distribution of performance period, which demonstrates the days between start and end date.
+The right skewed distribution indicates that the majority of contract duration is within a year.
+An over 5 years performance period implies a massive project that requires extensive preparation.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| 
+ggplot(data, aes(x = StartDate, y = EndDate)) +
+  geom_point(aes(color = Contract)) +
+  labs(x = "Start Date", y = "End Date", title = "Start Date vs End Date") +
+  theme(legend.position = "none")  # This removes the legend
+```
 
@@
 
+Federal solicitation methods are either competitive or non-competitive.
+Competitive contracts accept bids from all or some subset of suppliers.
+Non-competitive contracts do not, citing certain exceptions, including a “pressing emergency in which delay would be injurious to the public interest,” or when the estimated cost to the buyer is below certain thresholds defined for some goods and services.
+[@ProcurementSources]
+
 ## Idealized Methodology

### 2024-11-26 scripts/03-clean_data.R

@@ -2,7 +2,7 @@
 # Purpose: This script reads and cleans raw procurement data,
 # performing various transformations, and outputs the cleaned data to a CSV file.
 # Author: Deyi Kong
-# Date: November 25th, 2024
+# Date: November 26th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
@@ -69,10 +69,17 @@ cleaned_data <- tryCatch(
 
 # Add preparatory phase and contract total days
 cleaned_data <- cleaned_data %>%
-  mutate(PreparatoryPhase = as.numeric(StartDate - AwardDate),
-         ContractDays = as.numeric(EndDate - StartDate),
-         PhaseDays = abs(PreparatoryPhase)
-         )
+  mutate(
+    PreparatoryPhase = as.numeric(StartDate - AwardDate),
+    ContractDays = as.numeric(EndDate - StartDate),
+    PhaseDays = abs(PreparatoryPhase)
+    ) %>%
+  # Filter out the contracts before 2020
+  mutate(
+    Year = as.integer(format(as.Date(AwardDate), "%Y"))
+    ) %>%
+  filter(Year != 2019) %>%
+  select(-Year)  # Remove Year column after filtering
 
 # # Summary all the variables
 # sort(table(cleaned_data$Contract), decreasing = T)
@@ -121,6 +128,7 @@ cleaned_data <- cleaned_data %>%
   mutate(Supplier = clean_supplier_names(Supplier))
 
 ## Check variables
+# sort(table(cleaned_data$Buyer), decreasing = T)
 # sort(table(cleaned_data$Supplier), decreasing = T)
 # sort(table(cleaned_data$PreparatoryPhase), decreasing = T)
 # summary(cleaned_data$PreparatoryPhase)

### 2024-11-26 scripts/04-test_analysis_data.R

@@ -23,7 +23,7 @@ test_results <- list()
 test_results$structure <- test_that("Data structure", {
   expect_true(is.data.frame(cleaned_data), info = "clean_data is not a data frame.")
   expect_equal(ncol(cleaned_data), 10, info = "clean_data does not have the correct number of columns.")
-  expect_equal(nrow(cleaned_data), 950, info = "clean_data does not have the correct number of rows.")
+  expect_equal(nrow(cleaned_data), 948, info = "clean_data does not have the correct number of rows.")
 })
 
 # test if column names match expected names

### 2024-11-26 paper/paper.qmd
Update README.md and paper

@@
 #| label: fig-preparatory
-#| fig-cap: The histogram shows the distribution of preparatory period, which demonstrates the days between start and awarded date.
+#| fig-cap: The histogram shows the distribution of preparatory period.
 #| echo: false
@@
 #| message: false
+#| fig.width: 4    # Width of the figure in inches
+#| fig.height: 3   # Height of the figure in inches
 
@@
 #| fig-cap: The histogram shows the distribution of performance period, which demonstrates the days between start and end date.
-The right skewed distribution indicates that the majority of contract duration is within a year.
 #| echo: false
@@
 
-\newpage
+# Model
 
-# Model
+To understand the factors influencing the amount of federal procurement contracts awarded to Microsoft. I construct a linear regression model to explains variations in contract values.
 
@@
 
-+ Model Justification
+This paper utilizes a linear regression model to examine the relationships between the dependent variable.
+
+$$Amount = \beta_0 + \beta_1 Contract + \beta_2 Buyer + \beta_3 ContractDays + \beta_4 PhaseDays + \epsilon$$
+As defined in [@sec-data], each predictor variable has been chosen to reflect characteristics that influence the contract amount.
 
@@
 
-## Model Validation
+The model takes several assumptions, including the relationship between the predictors and the outcome is linear, each contract is independent, the variance of the error term is constant, and predictor variables are not highly correlated with each other.
 
-\newpage
+While the model presented offers valuable insights into the dynamics of Microsoft’s federal procurement contracts, there are some limitations including categorization of contract types, generalizability, data quality, potential for endogeneity, and unobserved variables.
+These limitations will be discussed in [@sec-discussion].
+
+## Model Validation?
+
+R-squared, adjusted R-squared
+
+tests (e.g. VIF, AIC, BIC).
 
@@
 #| label: fig-mobilization
-#| fig-cap: The scatter plot shows the distribution of performance period, which demonstrates the days between start and end date.
-The right skewed distribution indicates that the majority of contract duration is within a year.
-An over 5 years performance period implies a massive project that requires extensive preparation.
+#| fig-cap: The scatter plot shows ...
 #| echo: false
@@
 #| label: fig-performance
-#| fig-cap: The scatter plot shows the distribution of performance period, which demonstrates the days between start and end date.
-The right skewed distribution indicates that the majority of contract duration is within a year.
-An over 5 years performance period implies a massive project that requires extensive preparation.
+#| fig-cap: The scatter plot shows ....
 #| echo: false
@@
 
-## Additional Tables & Figures
-
 ## Methodology

### 2024-11-26 paper/paper.qmd
Update Methodology

@@
 
-## Data Cleaning Notes
-
 ## Methodology
 
-Federal solicitation methods are either competitive or non-competitive.
-Competitive contracts accept bids from all or some subset of suppliers.
-Non-competitive contracts do not, citing certain exceptions, including a “pressing emergency in which delay would be injurious to the public interest,” or when the estimated cost to the buyer is below certain thresholds defined for some goods and services.
-[@ProcurementSources]
+The methodology employed in this study involves analyzing federal procurement contracts awarded to Microsoft, leveraging data from two primary sources: Public Services and Procurement Canada (PSPC) and the Treasury Board of Canada Secretariat (TBS).
+These two datasets provide information on government procurement, but they come with distinct limitations and reporting structures [@ProcurementSources].
 
-## Idealized Methodology
+1. Data Sources and Structure:
+
+Federal procurement data is categorized into two main streams:
+
++ PSPC serves as the central procurement body for most federal government departments and is responsible for managing a significant portion of federal contracts.
+It maintains a public platform called CanadaBuys, where procurement information is published.
+However, due to migration from the older BuyAndSell platform and the incomplete nature of pre-2022 data, some variables, such as contract types and trade agreement information, are missing or inconsistently reported.
+The PSPC data is particularly useful for understanding contracts where the PSPC is the managing body, but only contracts above a certain dollar threshold are publicly available.
+
++ TBS consolidates procurement data from all government departments and requires proactive publication under the Access to Information Act.
+The TBS dataset is more comprehensive in terms of the number of contracts, but it suffers from inconsistencies and gaps in the reported fields.
+TBS publishes data quarterly, and records before 2017 should not be considered complete.
+Unlike PSPC, TBS data includes the full details of contract awards but does not incorporate tender or solicitation data.
+
+2. Data Integration and Challenges:
+
+There is an overlap between TBS and PSPC data, such that many contracts reported by PSPC are also included in the TBS dataset.
+However, the two datasets do not always align perfectly.
+Contracts reported by TBS may contain different details from the PSPC records.
+
+This mismatch is exacerbated by the lack of a consistent unique identifier that would allow for easy linking between the two datasets.
+While about 20% of PSPC awards have matching TBS records, many contracts do not, making it impossible to perfectly merge the two datasets.
+
+To address this, both datasets were kept separate in [@ProcurementSources].
+PSPC data was used primarily for its comprehensive tender information, while TBS data was relied upon for its broader coverage of contract awards, spanning more years and a greater volume of contracts.
+
+3. Solicitation Methods and Contract Types:
+
+Federal procurement processes are governed by various solicitation methods, which determine how contracts are awarded. These methods are categorized into competitive and non-competitive processes.
+
+Four competitive options:
+
++ Open bidding: Any supplier can bid in response to an online solicitation.
++ Selective tendering: Only some prequalified suppliers can bid.
++ Limited tendering: Deviation from requirements of the relevant trade agreements (see Trade Agreements), still allowing bids.
++ Traditional: Many suppliers can bid but not in response to a public, online solicitation; for example, responding to an email sent to suppliers by the government.
+
+Two non-competitive options:
+
++ Advanced Contract Award Notice: Notice to suppliers that the buyer intends to award a contract to a pre-identified supplier, while accepting a Statement of Capabilities, not a bid, from challenging suppliers.
++ Non-competitive: Blanket category for all other non-competitive processes.
 

### 2024-11-26 paper/paper.qmd
Update Paper

@@
 required_packages <- c(
-  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra", "ggcorrplot", "modelsummary"
+  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra"
 )
@@
 library(kableExtra)
-library(ggcorrplot)
-library(modelsummary)
 
@@
 
+## Limitations of the Model
+
+One limitation of the model is that it may fail to account for certain external factors or unobserved variables such as political factors, budgetary constraints, and departmental priorities.
+These external factors could lead to omitted variable bias, which also influence procurement decision.
+
+The model assumes a linear relationship between the independent variables and the dependent variable.
+However, procurement dynamics may be more complex, with non-linear interactions between variables.
+[The model categorizes contract types and buyers into broad groups, which may oversimplify the true diversity of contracts awarded.
+
+There may be issues of endogeneity, where causality runs in both directions between some predictors and the outcome variable.
+The amount of procurement from a buyer may influence the allocation of future budgets to that department, which in turn affects future contract amounts awarded to Microsoft.
+This reciprocal relationship could introduce bias in the coefficient estimates if not properly addressed.
+
 \newpage

### 2024-11-26 scripts/06-model_data.R
Update model

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Run a linear regression model to predict 
 # Author: Deyi Kong
-# Date: November 24th, 2024
+# Date: November 26th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` and 'here' packages must be installed
@@ -9,10 +9,24 @@
 
 # load libraries
 library(tidyverse)
+library(here)
+library(arrow)
 
+# convert some columns to the appropriate data types
+cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
+cleaned_data <- cleaned_data %>%
+  mutate(
+    Contract = as.factor(Contract),
+    Buyer = as.factor(Buyer),
+    Supplier = as.factor(Supplier)
+  )
 
+# linear model for contract dollar amount based on various predictors
+lm_model <- lm(Amount ~ Contract + Buyer + Supplier + ContractDays + PhaseDays,
+                     data = cleaned_data
+               )
 
 # save modified data for the model as a CSV + Parquet and save models as RDS
 write_csv(cleaned_data, here::here("data/analysis_data/data.csv"))
 write_parquet(cleaned_data, here::here("data/analysis_data/data.parquet"))
-saveRDS(lm_model_1, here::here("models/lm_model_1.rds"))
+saveRDS(lm_model, here::here("models/lm_model.rds"))

### 2024-11-26 paper/paper.qmd
Update Discussion

@@
 
-## Limitations of the Model
+##
+
+## Limitations
 
@@
 This reciprocal relationship could introduce bias in the coefficient estimates if not properly addressed.
+
+## Next Steps
+
+This paper provides a foundation for a deeper understanding of federal contract awards to Microsoft, but it also suggests several directions for future exploration.
+First, the study can expand to include comparisons with other major suppliers.
+By examining whether Microsoft's procurement trends are unique or reflective of broader patterns across industries.
+This gives a better understanding on assess whether certain dynamics are systemic within federal contracting practices.
+
+Data quality and consistency issues are also of concern.
+The anomalies identified in this study, such as incomplete or inconsistent contract dates, highlight the need for improved data reporting standards.
+These discrepancies can be addressed in collaboration with government agencies to improve the reliability of procurement datasets and ensure that future analyses are more reliable and actionable.
 

### 2024-11-26 paper/paper.qmd
Update paper

@@
 
+### Model Justification
+We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
+
 ## Assumptions and Limitations
@@
 
-## Model Validation?
+## Model Validation [REMOVE?]
 
@@
 
-##
+## Data
+
+The completeness and reliability of the data. For the missing or erroneous records, despite the data cleaning process, introduce bias and distort the findings.
+
+Since the data of the procurement contracts awarded to Microsoft in Canada is between 2020 and 2024 (up to September), the analysis provides a snapshot of recent procurement trends during this period.
+However, the data do not capture longer-term patterns or emerging shifts in procurement policies.
 
@@
 However, procurement dynamics may be more complex, with non-linear interactions between variables.
-[The model categorizes contract types and buyers into broad groups, which may oversimplify the true diversity of contracts awarded.
+The model categorizes contract types and buyers into broad groups, which may oversimplify the true diversity of contracts awarded.
 
@@
 
-## Next Steps
+## Weaknesses and Next Steps
 
@@
 
-1. Data Sources and Structure:
+### Data Sources and Structure
 
@@
 
-2. Data Integration and Challenges:
+### Data Integration and Challenges:
 
@@
 
-3. Solicitation Methods and Contract Types:
+### Solicitation Methods and Contract Types:
 

### 2024-11-26 scripts/07-validation.R

@@ -17,6 +17,24 @@ library(caret)
 cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
 
 # data processing
+cleaned_data <- cleaned_data %>%
+  mutate(
+    Contract = as.factor(Contract),
+    Buyer = as.factor(Buyer),
+    Supplier = as.factor(Supplier),
+    AwardDate = as.Date(AwardDate),
+    StartDate = as.Date(StartDate),
+    EndDate = as.Date(EndDate)
+  ) %>%
+  mutate(PollRecency = as.numeric(Sys.Date() - StartDate))
+  
+
+# split the data into training (70%) and testing (30%) sets
+set.seed(123) # Set seed for reproducibility
+train_index <- createDataPartition(cleaned_data$CandidateName, p = 0.7, list = FALSE)
+train_data <- cleaned_data[train_index, ]
+test_data <- cleaned_data[-train_index, ]
+
 
 
 write_parquet(rmse_results, here::here("data/analysis_data/rmse_results.parquet"))

### 2024-11-26 scripts/07-replications.R
Add files via upload

@@ -1,0 +1,19 @@
+#### Preamble ####
+# Purpose: Replicated graphs from... [...UPDATE THIS...]
+# Author: Deyi Kong
+# Date: November 24th, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` packages must be installed
+# Any other information needed? Make sure you are in the `Procurement` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+
+#### Load data ####
+# [...ADD CODE HERE...]
+
+
+

### 2024-11-26 scripts/07-validation.R

@@ -17,24 +17,6 @@ library(caret)
 cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
 
 # data processing
-cleaned_data <- cleaned_data %>%
-  mutate(
-    Contract = as.factor(Contract),
-    Buyer = as.factor(Buyer),
-    Supplier = as.factor(Supplier),
-    AwardDate = as.Date(AwardDate),
-    StartDate = as.Date(StartDate),
-    EndDate = as.Date(EndDate)
-  ) %>%
-  mutate(PollRecency = as.numeric(Sys.Date() - StartDate))
-  
-
-# split the data into training (70%) and testing (30%) sets
-set.seed(123) # Set seed for reproducibility
-train_index <- createDataPartition(cleaned_data$CandidateName, p = 0.7, list = FALSE)
-train_data <- cleaned_data[train_index, ]
-test_data <- cleaned_data[-train_index, ]
-
 
 
 write_parquet(rmse_results, here::here("data/analysis_data/rmse_results.parquet"))

### 2024-11-27 scripts/02-download_data.R
Add files via upload

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: This script loads the raw procurement buyer data from a CSV file for further analysis.
 # Author: Deyi Kong
-# Date: November 24th, 2024
+# Date: November 22th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
@@ -10,5 +10,5 @@
 # load library
 library(readr)
 
-# Read the raw presidential polls data
+# Read the raw federal procurement data
 data <- read_csv(here::here("data/raw_data/procurement.csv"))

### 2024-11-28 scripts/00-simulate_data.R
Update raw data and simulation

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulate data 
 # Author: Deyi Kong
-# Date: November 25th, 2024
+# Date: November 28th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
@@ -15,9 +15,9 @@ library(arrow)
 set.seed(123)
 
 # define possible values to simulate data
-contract <- c("License/Maintenance fees", "Computer services", "Consultants", "Application software", "Production and operations")
-buyer <- c("Employment and Social Development Canada", "Global Affairs Canada", "National Defence", "Fisheries and Oceans Canada")
-supplier <- c("MICROSOFT CANADA INC.", "MICROSOFT CORPORATION", "MICROSOFT CANADA CO.")
+contract <- c("License/Maintenance fees", "Communication Services", "Computer Services", "Application software", "Production and operations")
+buyer <- c("Shared Services Canada", "Employment and Social Development Canada", "Royal Canadian Mounted Police", "Global Affairs Canada", "National Defence")
+supplier <- c("MICROSOFT CANADA INC.", "BELL TELEPHONE CO. OF CANADA", "MICROSOFT CANADA CO.", "BELL CANADA")
 n <- 100
 
 # generate random start and end dates for each poll
@@ -36,7 +36,10 @@ procurement_data <- data.frame(
   end_date = end_dates,
   preparatory_phase = as.numeric(start_dates - award_dates),
   contract_days = as.numeric(end_dates - start_dates)
-)
+) %>%
+  mutate(
+    phase_days = abs(preparatory_phase)
+  )
 
 # write the cleaned and simulated polling data to a CSV file
 write_csv(procurement_data, here::here("data/simulated_data/simulated_data.csv"))

### 2024-11-28 scripts/01-test_simulated_data.R

@@ -22,14 +22,14 @@ test_results <- list()
 # test if the data structure is correct
 test_results$structure <- test_that("Data structure", {
   expect_true(is.data.frame(data), info = "procurement_data is not a data frame.")
-  expect_equal(ncol(data), 9, info = "procurement_data does not have the correct number of columns.")
+  expect_equal(ncol(data), 10, info = "procurement_data does not have the correct number of columns.")
   expect_equal(nrow(data), 100, info = "procurement_data does not have the correct number of rows.")
 })
 
 # test if the column names match the expected column names
 test_results$column_names <- test_that("Column names", {
   expected_colnames <- c(
-    "contract", "buyer", "supplier", "amount", "award_date", "start_date", "end_date", "preparatory_phase", "contract_days"
+    "contract", "buyer", "supplier", "amount", "award_date", "start_date", "end_date", "preparatory_phase", "contract_days", "phase_days"
   )
   expect_equal(colnames(data), expected_colnames, info = "Column names do not match expected names.")
 })

### 2024-11-28 scripts/02-download_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: This script loads the raw procurement buyer data from a CSV file for further analysis.
 # Author: Deyi Kong
-# Date: November 22th, 2024
+# Date: November 28th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
@@ -10,5 +10,6 @@
 # load library
 library(readr)
 
-# Read the raw federal procurement data
-data <- read_csv(here::here("data/raw_data/procurement.csv"))
+# Read the raw federal procurement data of Microsoft and Bell
+microsoft_data <- read_csv(here::here("data/raw_data/procurement_microsoft.csv"))
+bell_data <- read_csv(here::here("data/raw_data/procurement_bell.csv"))

### 2024-11-28 paper/paper.qmd
Add files via upload

@@
 ---
-title: "Procurement Dynamics: Analyzing Federal Contract Awards to Microsoft"
+title: "Procurement Dynamics: Analyzing Federal Contract Awards to Microsoft and Bell"
 subtitle: "Trends and Insights of Microsoft’s Future in Canadian Government from 2020"
@@
 required_packages <- c(
-  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra"
+  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra", "modelsummary"
 )
@@
 library(kableExtra)
+library(modelsummary)
 
@@
 summary_table <- cleaned_data %>%
-  group_by(Contract) %>%
+  group_by(ContractType) %>%
   summarise(
@@
   ) %>%
-  filter(Count > 25) %>%
+  filter(Count > 50) %>%
   arrange(desc(Count))  # Arrange in descending order based on Count
@@
 **Buyer** is a character-type variable representing the federal government department or agency who initiate and award the contract.
-[@tbl-buyer] shows the top 10 buyer's name who has over 25 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors.
+[@tbl-buyer] shows the top 10 buyer's name who has over 50 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors.
 The table highlights that Employment and Social Development Canada (ESDC) and National Defence account over 50% of Microsoft's total awards.
@@
 #| label: tbl-buyer
-#| tbl-cap: The table summarizes top 10 buyer's name who has over 25 contracts with Microsoft in total with contract count, total dollar amount, and average dollar amount per contract. 
+#| tbl-cap: The table summarizes top 10 buyer's name who has over 50 contracts with Microsoft in total with contract count, total dollar amount, and average dollar amount per contract. 
 #| echo: false
@@
 
-# Summarize the data and filter buyers with counts >= 25
+# Summarize the data and filter buyers with counts >= 50
 summary_table <- cleaned_data %>%
@@
 **Supplier** is a character-type variable includes all the affiliated subsidiary companies of Microsoft.
-[@tbl-supplier] shows the top 5 buyer's name who has awarded over 25 contracts.
+[@tbl-supplier] shows the top 5 buyer's name who has awarded over 50 contracts.
 [DELETE THIS VARIABLE?
@@
 
-$$Amount = \beta_0 + \beta_1 Contract + \beta_2 Buyer + \beta_3 ContractDays + \beta_4 PhaseDays + \epsilon$$
+$$Amount_i = \beta_0 + \beta_1 ContractType + \beta_2 Buyer + \beta_3 ContractDays + \beta_4 PhaseDays + \epsilon_i$$
 As defined in [@sec-data], each predictor variable has been chosen to reflect characteristics that influence the contract amount.
@@
 tests (e.g. VIF, AIC, BIC).
+
+```{r}
+#| label: tbl-keyestimates
+#| tbl-cap: Summary of key model estimates for Microsoft, including coefficients for predictors like xxxx, with standard errors for each estimate.
+Model performance statistics, such as sample size, R², and adjusted R², are also displayed.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# load data and model
+lm_model_m <- readRDS(here::here("models/lm_model_microsoft.rds"))
+# lm_model_b <- readRDS(here::here("models/lm_model_bell.rds"))
+
+# model_list <- list("Microsoft Model" = lm_model_m, "Bell Model" = lm_model_b)
+modelsummary(model_list, 
+             output = "kableExtra", 
+             statistic = "std.error", 
+             coef_map = c("ContractType" = "Contract Type", 
+                          "Buyer" = "Buyer", 
+                          "ContractDays" = "ContractDays",
+                          "PhaseDays" = "PhaseDays")) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r}
+#| label: tbl-rfkeyestimates
+#| tbl-cap: Summary of key model estimates for Microsoft, including coefficients for predictors like xxxx, with standard errors for each estimate.
+Model performance statistics, such as sample size, R², and adjusted R², are also displayed.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# load data and model
+rf_model<- readRDS(here::here("models/rf_model.rds"))
+
+# Summary of the Random Forest model
+summary(rf_model)
+# Print the model summary
+print(rf_model)
+# Error rates
+rf_model$err.rate  # Error rates for each class (classification)
+# Access the OOB error rate
+rf_model$oob.times  # Times each observation was used in the model
+rf_model$oob.err    # Out-of-bag error estimate
+
+```
 

### 2024-11-28 scripts/03-clean_data.R

@@ -2,7 +2,7 @@
 # Purpose: This script reads and cleans raw procurement data,
 # performing various transformations, and outputs the cleaned data to a CSV file.
 # Author: Deyi Kong
-# Date: November 26th, 2024
+# Date: November 28th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
@@ -14,7 +14,14 @@ library(here)
 library(arrow)
 
 # Read data
-data <- read_csv(here::here("data/raw_data/procurement.csv"))
+microsoft_data <- read_csv(here::here("data/raw_data/procurement_microsoft.csv"))
+# bell_data <- read_csv(here::here("data/raw_data/procurement_bell.csv"))
+# 
+# # Combine data
+# data <- rbind(microsoft_data, bell_data)
+
+
+data <- microsoft_data
 
 # # Check if the contract region are all federal, then we will exclude it from the cleaned data
 # table(data$region)
@@ -59,7 +66,9 @@ cleaned_data <- tryCatch(
         Amount = as.numeric(gsub("[\\$,]", "", Amount))
       ) %>%
       # Drop contracts where EndDate is before StartDate or AwardDate due to error with unknown reason
-      filter(!(EndDate <= StartDate | EndDate <= AwardDate))
+      filter(!(EndDate <= StartDate | EndDate <= AwardDate)) %>%
+      # Drop Na for amount
+      filter(!is.na(Amount))
   }, # error handling to mitigate any cleaning issues
   error = function(e) {
     message("An error occurred during data cleaning: ", e)
@@ -82,60 +91,156 @@ cleaned_data <- cleaned_data %>%
   select(-Year)  # Remove Year column after filtering
 
 # # Summary all the variables
-# sort(table(cleaned_data$Contract), decreasing = T)
 # sort(table(cleaned_data$Buyer), decreasing = T)
 # sort(table(cleaned_data$Supplier), decreasing = T)
-# summary(cleaned_data$Amount)
-# summary(cleaned_data$StartDate)
-# summary(cleaned_data$AwardDate)
-# summary(cleaned_data$EndDate)
+# summary(cleaned_data)
 
-# Standardize the supplier's name due to capitalization and notation differences
[…]

### 2024-11-28 scripts/04-test_analysis_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: This script performs a series of validation tests on procurement data.
 # Author: Deyi Kong
-# Date: November 25th, 2024
+# Date: November 28th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, `here`, `arrow`, and `testthat` packages must be installed.
@@ -22,14 +22,14 @@ test_results <- list()
 # test structure of data
 test_results$structure <- test_that("Data structure", {
   expect_true(is.data.frame(cleaned_data), info = "clean_data is not a data frame.")
-  expect_equal(ncol(cleaned_data), 10, info = "clean_data does not have the correct number of columns.")
-  expect_equal(nrow(cleaned_data), 948, info = "clean_data does not have the correct number of rows.")
+  expect_equal(ncol(cleaned_data), 11, info = "clean_data does not have the correct number of columns.")
+  expect_equal(nrow(cleaned_data), 1632, info = "clean_data does not have the correct number of rows.")
 })
 
 # test if column names match expected names
 test_results$column_names_test <- test_that("Column names", {
   expected_colnames <- c(
-     "Contract", "Buyer", "Supplier", "Amount", "AwardDate", "StartDate", "EndDate", "PreparatoryPhase", "ContractDays", "PhaseDays"
+     "Contract", "Buyer", "Supplier", "Amount", "AwardDate", "StartDate", "EndDate", "PreparatoryPhase", "ContractDays", "PhaseDays", "ContractType"
   )
   expect_equal(names(cleaned_data), expected_colnames, info = "Column names do not match expected names.")
 })

### 2024-11-28 scripts/06-model_data.R

@@ -1,32 +1,46 @@
 #### Preamble ####
 # Purpose: Run a linear regression model to predict 
 # Author: Deyi Kong
-# Date: November 26th, 2024
+# Date: November 28th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` and 'here' packages must be installed
+# Pre-requisites: The `tidyverse`, `arrow`, `randomForest` and 'here' packages must be installed
 # Any other information needed? Make sure you are in the `Procurement` rproj
 
 # load libraries
 library(tidyverse)
 library(here)
 library(arrow)
+library(randomForest)
 
 # convert some columns to the appropriate data types
 cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
 cleaned_data <- cleaned_data %>%
   mutate(
-    Contract = as.factor(Contract),
+    ContractType = as.factor(ContractType),
     Buyer = as.factor(Buyer),
     Supplier = as.factor(Supplier)
   )
 
-# linear model for contract dollar amount based on various predictors
-lm_model <- lm(Amount ~ Contract + Buyer + Supplier + ContractDays + PhaseDays,
-                     data = cleaned_data
-               )
+# linear model for microsoft contract dollar amount based on various predictors
+lm_model_m <- lm(Amount ~ ContractType + Buyer + ContractDays + PhaseDays,
+                 data = cleaned_data
+                 %>% filter(Supplier == "Microsoft")
+                 )
+# linear model for bell contract dollar amount based on various predictors
+lm_model_b <- lm(Amount ~ ContractType + Buyer + ContractDays + PhaseDays,
+                  data = cleaned_data
+                 %>% filter(Supplier == "Bell")
+                 )
+
+# Train a Random Forest model for regression (predicting a continuous outcome)
+rf_model <- randomForest(Amount ~ ContractType + Buyer + ContractDays + PhaseDays,
+                         data = cleaned_data, ntree = 500)
 
 # save modified data for the model as a CSV + Parquet and save models as RDS
 write_csv(cleaned_data, here::here("data/analysis_data/data.csv"))
 write_parquet(cleaned_data, here::here("data/analysis_data/data.parquet"))
-saveRDS(lm_model, here::here("models/lm_model.rds"))
+saveRDS(lm_model_m, here::here("models/lm_model_microsoft.rds"))
[…]

### 2024-11-28 scripts/07-validation.R

@@ -4,7 +4,7 @@
 # Date: November 24th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse`, 'caret', and 'here' packages must be installed
+# Pre-requisites: The `tidyverse`, `here`, 'randomForest', and 'arrow' packages must be installed
 # Any other information needed? Make sure you are in the `Procurement` rproj
 
 
@@ -12,11 +12,68 @@
 library(tidyverse)
 library(here)
 library(caret)
+library(arrow)
+library(tibble)
+library(randomForest)
 
 # load data
 cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
 
 # data processing
+cleaned_data <- cleaned_data %>%
+  mutate(
+    Contract = as.factor(Contract),
+    Buyer = as.factor(Buyer),
+    Supplier = as.factor(Supplier),
+    AwardDate = as.Date(AwardDate),
+    StartDate = as.Date(StartDate),
+    EndDate = as.Date(EndDate)
+  )
 
+# split the data into training (80%) and testing (20%) sets
+set.seed(123) # Set seed for reproducibility
+train_index <- createDataPartition(cleaned_data$Amount, p = 0.8, list = FALSE)
+train_data <- cleaned_data[train_index, ]
+test_data <- cleaned_data[-train_index, ]
 
-write_parquet(rmse_results, here::here("data/analysis_data/rmse_results.parquet"))
+# linear model for microsoft contract dollar amount based on various predictors
+lm_model <- lm(Amount ~ ContractType + Buyer + ContractDays + PhaseDays,
+                 data = cleaned_data)
+
+# Make predictions on the testing set
+predictions <- predict(lm_model, newdata = test_data)
+
+# Extract actual values for RMSE calculation
+actual_values <- test_data$Amount
+
+# Calculate RMSE (Root Mean Squared Error)
+rmse_value <- sqrt(mean((actual_values - predictions)^2, na.rm = TRUE))
+
+# Store RMSE results in a tibble
+rmse_results <- tibble(
+  Model = "Linear Regression",
+  RMSE = rmse_value
+)
+
+# Train a Random Forest model on the training set
+rf_model <- randomForest(Amount ~ ContractType + Buyer + ContractDays + PhaseDays, data = train_data, ntree = 500)
+
+# Make predictions on the testing set
[…]

### 2024-12-03 paper/paper.qmd
Add files via upload

@@
 ---
-title: "Procurement Dynamics: Analyzing Federal Contract Awards to Microsoft and Bell"
-subtitle: "Trends and Insights of Microsoft’s Future in Canadian Government from 2020"
+title: "Procurement Dynamics: Analyzing Federal Contract Awards to Microsoft"
+subtitle: "Trends of Microsoft’s Future in Canadian Government from 2020"
 author: 
@@
 The analysis reveals that Microsoft dominates high-value contracts and demonstrates consistent partnerships with key government departments.
-This paper utilizes a linear model to predict the future contract amount trend for Microsoft.
+This paper utilizes a linear model and random forest to predict the future contract amount trend for Microsoft.
 By examining these procurement dynamics, this paper providing insights into the broader implications of tech-industry participation in government procurement."
@@
 library(modelsummary)
+library(ggcorrplot)
+library(gridExtra)
+library(rstanarm)
 
@@
 The main estimand in this study is the amount of federal procurement contracts awarded to Microsoft.
-This metric is analyzed across multiple dimensions, including contract categories, durations, and partnerships, to quantify Microsoft's role in federal procurement comprehensively.
+This metric is analyzed across multiple dimensions, including contract categories, duration, and partnerships, to quantify Microsoft's role in federal procurement comprehensively.
 The findings reveal that Microsoft plays a dominant role in IT infrastructure and modernization, securing the majority of high-value contracts and establishing consistent partnerships with departments such as Employment and Social Development Canada and National Defence.
@@
 PhaseDays take the absolute value of PreparatoryPhase because of the minor effect of either prepare before contract being awarded or prepare after contract being awarded.
+Last but not least, I created a new variable called ContractType, which categorizes the contracts based on keywords.
 
@@
 In this analysis, several key predictors were identified to evaluate procurement data effectively.
-The first variable, **Contract**, is a character-type variable that identifies each procurement contract awarded to Microsoft.
-[@tbl-contract] [SOME MODIFICATION OF CONTRACT CATEGORIES CLEANING PROCESS MAY APPLY]
+The first variable, **ContractType**, is a character-type variable that identifies each procurement contract awarded to Microsoft into Categories.
+[@tbl-contract]
 
@@
   ) %>%
-  filter(Count > 50) %>%
+  filter(Count > 25) %>%
   arrange(desc(Count))  # Arrange in descending order based on Count
@@
 #| label: tbl-buyer
-#| tbl-cap: The table summarizes top 10 buyer's name who has over 50 contracts with Microsoft in total with contract count, total dollar amount, and average dollar amount per contract. 
+#| tbl-cap: The table summarizes top 10 buyer's name who has over 25 contracts with Microsoft in total with contract count, total dollar amount, and average dollar amount per contract. 
 #| echo: false
@@
 
-# Summarize the data and filter buyers with counts >= 50
+# Summarize the data and filter buyers with counts >= 25
 summary_table <- cleaned_data %>%
@@
   ) %>%
-  filter(Count > 25) %>%
+  filter(Count > 15) %>%
   arrange(desc(Count))  # Arrange in descending order based on Count
@@
   column_spec(1, width = "22em")  # Set column width to wrap text
-
-```
-
-**Supplier** is a character-type variable includes all the affiliated subsidiary companies of Microsoft.
-[@tbl-supplier] shows the top 5 buyer's name who has awarded over 50 contracts.
-[DELETE THIS VARIABLE?
-OR MAYBE MOVE IT TO APPDX.]
-
-```{r}
-#| label: tbl-supplier
-#| tbl-cap: The table summarizes the main Microsoft supplier's name who has over 25 contracts with buyers in total with contract count, total dollar amount, and average dollar amount per contract. 
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# Summarize the data and filter buyers with counts >= 50
-summary_table <- cleaned_data %>%
-  group_by(Supplier) %>%
-  summarise(
-    Count = n(), 
-    TotalAmount = sum(Amount),  # Calculate the total contract value
-    AvgAmount = mean(Amount)    # Calculate the average contract value
-  ) %>%
-  filter(Count > 50) %>%
-  arrange(desc(Count))  # Arrange in descending order based on Count
-
-# Format the TotalAmount column with dollar format
-summary_table$TotalAmount <- scales::label_dollar()(summary_table$TotalAmount)
-summary_table$AvgAmount <- scales::label_dollar()(summary_table$AvgAmount)
-
-# Display the summary table using kableExtra for nicer formatting
-summary_table %>%
-  kable(col.names = c("Suppllier", "Contract", "Total Amount", "Average")) %>%
-  kable_styling(full_width = F, position = "center", bootstrap_options = c("striped", "hover"))
 
@@
 
-# Model
-
-To understand the factors influencing the amount of federal procurement contracts awarded to Microsoft. I construct a linear regression model to explains variations in contract values.
-
-## Model Set-Up {#sec-modset}
-
-This paper utilizes a linear regression model to examine the relationships between the dependent variable.
-
-$$Amount_i = \beta_0 + \beta_1 ContractType + \beta_2 Buyer + \beta_3 ContractDays + \beta_4 PhaseDays + \epsilon_i$$
-As defined in [@sec-data], each predictor variable has been chosen to reflect characteristics that influence the contract amount.
-
-### Model Justification
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-## Assumptions and Limitations
-
-The model takes several assumptions, including the relationship between the predictors and the outcome is linear, each contract is independent, the variance of the error term is constant, and predictor variables are not highly correlated with each other.
-
-While the model presented offers valuable insights into the dynamics of Microsoft’s federal procurement contracts, there are some limitations including categorization of contract types, generalizability, data quality, potential for endogeneity, and unobserved variables.
-These limitations will be discussed in [@sec-discussion].
-
-## Model Validation [REMOVE?]
-
-R-squared, adjusted R-squared
-
-tests (e.g. VIF, AIC, BIC).
-
 ```{r}
-#| label: tbl-keyestimates
-#| tbl-cap: Summary of key model estimates for Microsoft, including coefficients for predictors like xxxx, with standard errors for each estimate.
-Model performance statistics, such as sample size, R², and adjusted R², are also displayed.
+#| label: fig-cor
+#| fig-cap: This correlation matrix visualizes the relationships between key numerical predictor variables and the outcome variable, Amount.
+The figure shows correlations include a weak negative correlations between PreparatoryPhase and ContractDays; and weak positive correlations between predictors and Amount.
 #| echo: false
@@
 
-# load data and model
-lm_model_m <- readRDS(here::here("models/lm_model_microsoft.rds"))
-# lm_model_b <- readRDS(here::here("models/lm_model_bell.rds"))
+# filter to numerical predictor variables
+numerical_data <- cleaned_data %>%
+  select(Amount, PreparatoryPhase, ContractDays)
 
-# model_list <- list("Microsoft Model" = lm_model_m, "Bell Model" = lm_model_b)
-modelsummary(model_list, 
-             output = "kableExtra", 
-             statistic = "std.error", 
-             coef_map = c("ContractType" = "Contract Type", 
-                          "Buyer" = "Buyer", 
-                          "ContractDays" = "ContractDays",
-                          "PhaseDays" = "PhaseDays")) %>%
-  kable_styling(latex_options = c("striped", "scale_down"))
+correlation_matrix <- cor(numerical_data, use = "complete.obs")
+
+# create correlation matrix
+ggcorrplot(correlation_matrix, lab = TRUE) + 
+  theme(
+    axis.text.x = element_text(size = 10),  # Adjust x-axis text size
+    axis.text.y = element_text(size = 10) # Adjust y-axis text size
+  )
 ```
@@
 ```{r}
-#| label: tbl-rfkeyestimates
-#| tbl-cap: Summary of key model estimates for Microsoft, including coefficients for predictors like xxxx, with standard errors for each estimate.
-Model performance statistics, such as sample size, R², and adjusted R², are also displayed.
+#| label: fig-box
+#| fig-cap: The Boxplots shows the relationship of ContractDays and PhaseDays with ContractType. 
+#| fig-subcap: ["Contract Duration in each Contract Type","Phase Duration in each Contract Type"]
 #| echo: false
-#| eval: true
 #| warning: false
-#| message: false
+#| layout-nrow: 2
 
-# load data and model
-rf_model<- readRDS(here::here("models/rf_model.rds"))
+# Clean and transform the data
+data <- cleaned_data %>%
+  mutate(
+    ContractType = gsub("Related", "", ContractType), # Remove "Related"
+    ContractType = str_wrap(gsub("/", " or ", ContractType), width = 10) # Wrap long labels
+  )
 
-# Summary of the Random Forest model
-summary(rf_model)
-# Print the model summary
-print(rf_model)
-# Error rates
-rf_model$err.rate  # Error rates for each class (classification)
-# Access the OOB error rate
-rf_model$oob.times  # Times each observation was used in the model
-rf_model$oob.err    # Out-of-bag error estimate
+# First graph: ContractType vs. ContractDays
+ggplot(data, aes(x = ContractType, y = ContractDays)) +
+  geom_boxplot() +
+  theme_minimal() +
+  labs(x = "Contract Type", y = "Contract Duration") +
+  theme(axis.text.x = element_text(size = 6))
+
+# Second graph: ContractType vs. PhaseDays
+ggplot(data, aes(x = ContractType, y = PhaseDays)) +
+  geom_boxplot() +
+  theme_minimal() +
+  labs(x = "Contract Type", y = "Phase Duration") +
+  theme(axis.text.x = element_text(size = 6))
 
@@
 
+# Model
+
+## Model Set-Up {#sec-modset}
+
+This paper utilizes a Bayesian model to examine the factors influencing the dollar amount of federal procurement contracts awarded to Microsoft.
+Define $Amount_i$ as the dollar amount of the federal procurement contract awarded to Microsoft and $\mu_i$ is the expected value of the contract amount modeled as a linear combination of the predictors.
+The model for predicting $Amount_i$ is specified as follows:
+\begin{align*} 
+Amount_i \mid \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma)\\
+\mu_i &= \alpha + \beta_1 \cdot \text{ContractType}_i + \beta_2 \cdot \text{Buyer}_i + \beta_3 \cdot \text{ContractDays}_i + \beta_4 \cdot \text{PhaseDays}_i \\
+\alpha &\sim \text{Normal}(0, 2) \\
+\beta_1, \beta_2, \beta_3, \beta_4 &\sim \text{Normal}(0, 2) \\
+\sigma &\sim \text{Exponential}(1)
+\end{align*}
+Where:\
+- $\alpha$ is the intercept term.\
+- $\beta_1, \beta_2, \beta_3, \beta_4$ are the coefficients associated with the predictors:\
+  - $ContractType_i$: A categorical variable representing the contract type.\
+  - $Buyer_i$: A categorical variable representing the buyer of the contract.\
+  - $ContractDays_i$: A numeric variable representing the duration of the contract is active.\
+  - $PhaseDays_i$: A numeric variable representing the days between AwardDate and StartDate.
+
+The model is fit using the `rstanarm` package of @rstanarm in R [@citeR], with the specified priors guiding the model while allowing it to adapt to the data.
+
+### Model Justification
+This paper expect that the factors - such as ContractType, Buyer, ContractDays and PhaseDays - will influence the outcome of contract amount. In particular, ...
+
+In this paper, I compared three different modeling approaches to examine the factors influencing the federal procurement contract amount: linear regression, random forest, and Bayesian regression.
+Each of these models offers different advantages and trade-offs.
+More detailed justification about model selection will be discussed in [@sec-discussion].
+The Bayesian model is selected due to better performance and flexibility.
+
+Bayesian approach can explicitly model uncertainty by incorporating prior knowledge and provides a full posterior distribution of the model parameters.
+This is useful when the data is sparse and prior knowledge about the contract types, buyers, or duration is available.
+The linear regression structure was selected due to its simplicity and interpretability.
+The numeric dependent variable, $Amount_i$, gives a chance of linear regression for modeling the relationship between the contract amount and its predictors.
+
+By including weakly informative priors for the coefficients and residual standard deviation, the model remains flexible and interpretable, while still preventing extreme parameter estimates.
+The use of an exponential prior for $\sigma$ allows the model to adapt a scale of residuals in the data, providing estimates of the variance.
+The normal prior for the coefficients $\beta_1, \beta_2, \beta_3, \beta_4$ with mean 0 and standard deviation 2 shows the expectation that the coefficients should not be too larg but flexible enough to adapt the data.
+
+This Bayesian linear regression model balances complexity with interpretability, and its flexibility allows us to draw meaningful observations from the data while quantifying uncertainty around the estimates, especially when the sample size is not large or when the predictors have a complex relationship with the outcome.
+[TOO GENERAL, CHANGE]
+
+## Assumptions and Limitations
+
+The model takes several assumptions, including linear relationship between predictors, normality of the residuals, independence of observations, homoscedasticity, and reasonable chosen priors.
+
+While the model presents some valuable observations of Microsoft’s federal procurement contracts, there are some model limitations including prior function selection, data quality, potential for endogeneity, generalizability, and unobserved variables.
+These limitations will be discussed more detailed in [@sec-discussion].
+
 # Results {#sec-result}
+
+## Data
 
@@
   mutate(
-    Year = year(AwardDate),
-    Month = month(AwardDate),
     Period = case_when(
@@
       Month %in% 9:12 ~ "Sep-Dec"
-    ),
-    YearPeriod = paste(Year, Period, sep = "-")  # Combine Year and Period
+      )
   )
@@
 contract_counts <- data %>%
-  group_by(YearPeriod) %>%
+  group_by(Period) %>%
   summarise(Count = n())
@@
 # Heatmap plot
-ggplot(contract_counts, aes(x = YearPeriod, y = Count, fill = Count)) +
+ggplot(contract_counts, aes(x = Period, y = Count, fill = Count)) +
   geom_tile() +
-  scale_fill_gradient(low = "white", high = "red") +
+  scale_fill_gradient(low = "lightblue", high = "darkblue") +
   labs(x = "Year-Period", y = "Frequency (Contract Count)", title = "Heatmap of Contracts by Four-Month Periods") +
@@
 
+## Model
+R-squared, adjusted R-squared
+
+tests (e.g. VIF, AIC, BIC).
+
+Our results are summarized in @tbl-modelresults.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+bayesian_model <-
+  readRDS(file = here::here("models/bayesian_model.rds"))
+```
+
+```{r}
+#| label: tbl-modelresults
+#| echo: false
+#| eval: true
+#| tbl-cap: "Explanatory models of contract amount based on ContractType, Buyer, ContractDays, PhaseDays and Month"
+#| warning: false
+#| message: false
+
+# Generate the model summary
+
+bayesian_model_full <- readRDS(here::here("models/bayesian_model_full.rds"))
+bayesian_model <- readRDS(here::here("models/bayesian_model.rds"))
+
+model_list <- list("Bayesian Model_Full" = bayesian_model_full, "Bayesian Model" = bayesian_model)
+
+modelsummary(
+  model_list,
+  output = "kableExtra", 
+  coef_map = c("ContractType" = "ContractType", "Buyer" = "Buyer", 
+               "ContractDays" = "ContractDays", "PhaseDays" = "PhaseDays")
+) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
+```
+
 \newpage
@@
 
+## Model Selection
+
 ## Weaknesses and Next Steps
@@
 
-# Appendix {#sec-appx}
+\appendix
 
-## Methodology
+# Appendix {-}
+
+# Additional data details
+
+# Model details {#sec-model-details}
+
+## Bayesian Model
+
+### Posterior predictive check
+
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+
+pp_check(bayesian_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+posterior_vs_prior(bayesian_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set3") +
+  theme(
+    legend.position = "none", 
+    axis.text.y = element_text(size = 7)
+    ) +
+  coord_flip()
+
+```
+
+### Diagnostics
+
+@fig-trace-1 is a trace plot. It shows... This suggests...
+
+@fig-trace-2 is a trace plot. It shows... This suggests...
+
+@fig-rhat is a Rhat plot. It shows... This suggests...
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-trace
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot - Contract Type","Trace plot - Buyers"]
+#| layout-nrow: 2
+
+# Extract the parameter names
+param_names <- colnames(as.matrix(bayesian_model))
+
+# Use grep to find all parameters that contain "ContractType"
+contract_type_params <- grep("ContractType", param_names, value = TRUE)
+
+# Use grep to find all parameters that contain "Buyer"
+buyer_params <- grep("Buyer", param_names, value = TRUE)
+
+# # Group 1: Buyers containing the word "the"
+# buyers_with_the <- grep("the", buyer_params, value = TRUE, ignore.case = TRUE)
+# 
+# # Group 2: Buyers containing the word "the"
+# buyers_with_and <- grep("and", buyer_params, value = TRUE, ignore.case = TRUE)
+# buyers_with_and <- buyers_with_and[!grepl("the", buyers_with_and, ignore.case = TRUE)]
+# 
+# # Group 3: Buyers containing the word "the"
+# buyers_with_of <- grep("of", buyer_params, value = TRUE, ignore.case = TRUE)
+# buyers_with_of <- buyers_with_of[!grepl("the|and", buyers_with_of, ignore.case = TRUE)]
+# 
+# # Group 4: Buyers containing the word "Canada"
+# buyers_with_canada <- grep("Canada", buyer_params, value = TRUE, ignore.case = TRUE)
+# buyers_with_canada <- buyers_with_canada[!grepl("the|and|of", buyers_with_canada, ignore.case = TRUE)]
+# 
+# # Group 5: Buyers that do not contain "the" or "Canada"
+# other_buyers <- grep("Canada|the|and|of", buyer_params, value = TRUE, invert = TRUE, ignore.case = TRUE)
+
+# Plot trace for ContractType
+plot(bayesian_model, "trace", pars = contract_type_params) +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",                # Place legend at the bottom
+    legend.text = element_text(size = 6),      # Adjust legend label size
+    legend.title = element_text(size = 8),     # Adjust legend title size
+    axis.text.x = element_text(size = 6),      # Adjust x-axis text size
+    axis.text.y = element_text(size = 6),      # Adjust y-axis text size
+    strip.text = element_text(size = 5)
+    )
+
+# Plot trace for ContractType
+plot(bayesian_model, "trace", pars = buyer_params) +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",                # Place legend at the bottom
+    legend.text = element_text(size = 6),      # Adjust legend label size
+    legend.title = element_text(size = 8),     # Adjust legend title size
+    axis.text.x = element_text(size = 4),      # Adjust x-axis text size
+    axis.text.y = element_text(size = 5),      # Adjust y-axis text size
+    strip.text = element_text(size = 5)
+    )
+
+# # Plot trace for buyers' name containing "the"
+# plot(bayesian_model, "trace", pars = buyers_with_the)
+# 
+# # Plot trace for buyers' name containing "and"
+# plot(bayesian_model, "trace", pars = buyers_with_and)
+# 
+# # Plot trace for buyers' name containing Canada
+# plot(bayesian_model, "trace", pars = buyers_with_canada)
+# 
+# # Plot trace for buyers' name containing "of"
+# plot(bayesian_model, "trace", pars = buyers_with_of)
+# 
+# # Plot trace for Other buyers
+# plot(bayesian_model, "trace", pars = other_buyers)
+```
+```{r}
+#| label: fig-rhat
+#| echo: false
+#| eval: true
+#| fig-cap: "Rhat plot for bayesian model"
+#| message: false
+#| warning: false
+#| fig-subcap: ["Rhat plot - Full model","Rhat plot - Reduced model"]
+#| layout-ncol: 2
+
+plot(bayesian_model_full, "rhat")
+plot(bayesian_model, "rhat")
+```
+## Linear Regression Model
+
+```{r}
+#| label: tbl-keyestimates
+#| tbl-cap: Summary of key model estimates for Microsoft, including coefficients for predictors like xxxx, with standard errors for each estimate.
+Model performance statistics, such as sample size, R², and adjusted R², are also displayed.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# load data and model
+lm_model_full <- readRDS(here::here("models/lm_model_full.rds"))
+lm_model <- readRDS(here::here("models/lm_model.rds"))
+model_list <- list("Linear Model_Full" = lm_model_full, "Linear Model" = lm_model)
+
+modelsummary(
+  model_list,
+  output = "kableExtra", 
+  statistic = "std.error", 
+  coef_map = c("ContractType" = "ContractType", "Buyer" = "Buyer", 
+               "ContractDays" = "ContractDays", "PhaseDays" = "PhaseDays")
+  ) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+## Random Forest Model
+
+```{r}
+#| label: tbl-rfkeyestimates
+#| tbl-cap: Summary of key model estimates for Microsoft, including coefficients for predictors like xxxx, with standard errors for each estimate.
+Model performance statistics, such as sample size, R², and adjusted R², are also displayed.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# load data and model
+rf_model<- readRDS(here::here("models/rf_model.rds"))
+
+# Summary of the Random Forest model
+summary(rf_model, output = "kableExtra")
+
+# # Error rates
+# rf_model$err.rate  # Error rates for each class (classification)
+# rf_model$oob.err    # Out-of-bag error estimate
+
+```
+
+# Methodology
 
@@
 
-### Data Sources and Structure
+## Data Sources and Structure
 
@@
 
-### Data Integration and Challenges:
+## Data Integration and Challenges:
 
@@
 
-### Solicitation Methods and Contract Types:
+## Solicitation Methods and Contract Types:
 

### 2024-12-03 models/server.R

@@ -1,0 +1,19 @@
+#### Preamble ####
+# Purpose: Initialize and run the server
+# Author: Deyi Kong
+# Date: December 2nd, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `plumber` packages must be installed
+# Any other information needed? No
+
+# Load libraries
+library(plumber)
+
+#### API Setup ####
+# Plumb the Plumber API from the plumber.R file, which contains route definitions
+serve_model <- plumb("models/plumber.R")
+
+#### Run the API ####
+# Start the API server to listen for incoming requests
+serve_model$run(port = 8000)

### 2024-12-03 scripts/00-simulate_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulate data 
 # Author: Deyi Kong
-# Date: November 28th, 2024
+# Date: November 30th, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
@@ -17,7 +17,7 @@ set.seed(123)
 # define possible values to simulate data
 contract <- c("License/Maintenance fees", "Communication Services", "Computer Services", "Application software", "Production and operations")
 buyer <- c("Shared Services Canada", "Employment and Social Development Canada", "Royal Canadian Mounted Police", "Global Affairs Canada", "National Defence")
-supplier <- c("MICROSOFT CANADA INC.", "BELL TELEPHONE CO. OF CANADA", "MICROSOFT CANADA CO.", "BELL CANADA")
+supplier <- c("MICROSOFT CANADA INC.", "MICROSOFT CANADA CO.")
 n <- 100
 
 # generate random start and end dates for each poll

### 2024-12-03 scripts/02-download_data.R

@@ -12,4 +12,3 @@ library(readr)
 
 # Read the raw federal procurement data of Microsoft and Bell
 microsoft_data <- read_csv(here::here("data/raw_data/procurement_microsoft.csv"))
-bell_data <- read_csv(here::here("data/raw_data/procurement_bell.csv"))

### 2024-12-03 scripts/03-clean_data.R

@@ -2,26 +2,21 @@
 # Purpose: This script reads and cleans raw procurement data,
 # performing various transformations, and outputs the cleaned data to a CSV file.
 # Author: Deyi Kong
-# Date: November 28th, 2024
+# Date: December 2nd, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
 # Any other information needed? Make sure you are in the `Procurement` rproj
 
 # load libraries
+library(dplyr)
 library(tidyverse)
 library(here)
 library(arrow)
+library(lubridate)
 
 # Read data
-microsoft_data <- read_csv(here::here("data/raw_data/procurement_microsoft.csv"))
-# bell_data <- read_csv(here::here("data/raw_data/procurement_bell.csv"))
-# 
-# # Combine data
-# data <- rbind(microsoft_data, bell_data)
-
-
-data <- microsoft_data
+data <- read_csv(here::here("data/raw_data/procurement_microsoft.csv"))
 
 # # Check if the contract region are all federal, then we will exclude it from the cleaned data
 # table(data$region)
@@ -49,13 +44,6 @@ cleaned_data <- tryCatch(
         EndDate = end_date
       ) %>%
       mutate(
-        # To replace missing (NA) start date with the value of awarded date column
-        StartDate = case_when(
-          is.na(StartDate) ~ AwardDate,
-          TRUE ~ StartDate
-        )
-      ) %>%
-      mutate(
         # Specify the format of the date for proper parsing
         StartDate = mdy(StartDate), # Use mdy() for MM/DD/YYYY format
         AwardDate = mdy(AwardDate), 
@@ -68,7 +56,9 @@ cleaned_data <- tryCatch(
       # Drop contracts where EndDate is before StartDate or AwardDate due to error with unknown reason
       filter(!(EndDate <= StartDate | EndDate <= AwardDate)) %>%
       # Drop Na for amount
-      filter(!is.na(Amount))
+      filter(!is.na(Amount) & !is.na(StartDate)) %>%
+      # drop the significant large contract
+      filter(Amount != 102260150)
[…]

### 2024-12-03 scripts/04-test_analysis_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: This script performs a series of validation tests on procurement data.
 # Author: Deyi Kong
-# Date: November 28th, 2024
+# Date: December 2nd, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, `here`, `arrow`, and `testthat` packages must be installed.
@@ -22,14 +22,14 @@ test_results <- list()
 # test structure of data
 test_results$structure <- test_that("Data structure", {
   expect_true(is.data.frame(cleaned_data), info = "clean_data is not a data frame.")
-  expect_equal(ncol(cleaned_data), 11, info = "clean_data does not have the correct number of columns.")
-  expect_equal(nrow(cleaned_data), 1632, info = "clean_data does not have the correct number of rows.")
+  expect_equal(ncol(cleaned_data), 14, info = "clean_data does not have the correct number of columns.")
+  expect_equal(nrow(cleaned_data), 943, info = "clean_data does not have the correct number of rows.")
 })
 
 # test if column names match expected names
 test_results$column_names_test <- test_that("Column names", {
   expected_colnames <- c(
-     "Contract", "Buyer", "Supplier", "Amount", "AwardDate", "StartDate", "EndDate", "PreparatoryPhase", "ContractDays", "PhaseDays", "ContractType"
+     "Contract", "Buyer", "Supplier", "Amount", "AwardDate", "StartDate", "EndDate", "PreparatoryPhase", "ContractDays", "PhaseDays", "Year", "Month", "ContractType", "BuyerCleaned"
   )
   expect_equal(names(cleaned_data), expected_colnames, info = "Column names do not match expected names.")
 })

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -39,3 +39,4 @@ summary(cleaned_data)
 missing_data <- cleaned_data |>
   summarise(across(everything(), ~ sum(is.na(.))))
 missing_data
+

### 2024-12-03 scripts/06-model_data.R

@@ -1,10 +1,10 @@
 #### Preamble ####
 # Purpose: Run a linear regression model to predict 
 # Author: Deyi Kong
-# Date: November 28th, 2024
+# Date: December 2nd, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse`, `arrow`, `randomForest` and 'here' packages must be installed
+# Pre-requisites: The `tidyverse`, `rstanarm`, `arrow`, `randomForest` and 'here' packages must be installed
 # Any other information needed? Make sure you are in the `Procurement` rproj
 
 # load libraries
@@ -12,6 +12,7 @@ library(tidyverse)
 library(here)
 library(arrow)
 library(randomForest)
+library(rstanarm)
 
 # convert some columns to the appropriate data types
 cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
@@ -22,25 +23,52 @@ cleaned_data <- cleaned_data %>%
     Supplier = as.factor(Supplier)
   )
 
-# linear model for microsoft contract dollar amount based on various predictors
-lm_model_m <- lm(Amount ~ ContractType + Buyer + ContractDays + PhaseDays,
+# full linear model for contract dollar amount based on various predictors
+lm_model_full <- lm(Amount ~ Contract + Buyer + ContractDays + PhaseDays + Month,
                  data = cleaned_data
-                 %>% filter(Supplier == "Microsoft")
-                 )
-# linear model for bell contract dollar amount based on various predictors
-lm_model_b <- lm(Amount ~ ContractType + Buyer + ContractDays + PhaseDays,
-                  data = cleaned_data
-                 %>% filter(Supplier == "Bell")
-                 )
+)
 
-# Train a Random Forest model for regression (predicting a continuous outcome)
-rf_model <- randomForest(Amount ~ ContractType + Buyer + ContractDays + PhaseDays,
+# linear model for contract dollar amount based on various predictors
+lm_model <- lm(Amount ~ ContractType + BuyerCleaned + ContractDays + PhaseDays + Month,
+                 data = cleaned_data
+               )
+
+# Random Forest model for regression (predicting a continuous outcome)
+rf_model <- randomForest(Amount ~ ContractType + BuyerCleaned + ContractDays + PhaseDays + Month,
[…]

### 2024-12-03 scripts/07-replications.R
Delete scripts/07-replications.R

@@ -1,19 +1,0 @@
-#### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
-# Author: Deyi Kong
-# Date: November 24th, 2024
-# Contact: deyi.kong@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: The `tidyverse` packages must be installed
-# Any other information needed? Make sure you are in the `Procurement` rproj
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-
-#### Load data ####
-# [...ADD CODE HERE...]
-
-
-

### 2024-12-03 paper/paper.qmd
Update Abstract and Introduction

@@
 date-format: long
-abstract: "This paper analyzes the federal procurement contracts awarded to Microsoft, focusing on trends in award values, contract categories, and the relationships between buyers and Microsoft.
-The analysis reveals that Microsoft dominates high-value contracts and demonstrates consistent partnerships with key government departments.
-This paper utilizes a linear model and random forest to predict the future contract amount trend for Microsoft.
-By examining these procurement dynamics, this paper providing insights into the broader implications of tech-industry participation in government procurement."
+abstract: "This paper analyzes federal procurement contracts awarded to Microsoft from 2020 to 2024, focusing on contract values, categories, and relationships between buyers and Microsoft.
+Multiple models were considered, with Bayesian modeling approach selected to predict Microsoft's contract amount.
+The analysis reveals Microsoft's dominant role in high-value contracts and its consistent partnerships with major government departments like National Defence.
+These findings provide informations evolving dynamics of federal procurement, raising questions about dependency and competitiveness."
 format: pdf
@@
 
-Government procurement serves as a critical avenue for public sector innovation, economic stimulus, and the fulfillment of strategic objectives.
+Government procurement refers to the acquisition of goods, services, and construction services by the government, spanning a wide range of purchases from office supplies to materials and services needed for large-scale infrastructure projects [@gac].
+This process enables public sector innovation, economic stimulus and fulfills strategic objectives.
 Within this landscape, the relationship between major technology providers and governmental bodies has become increasingly prominent.
-Among these, Microsoft has emerged as a pivotal partner in driving IT modernization and infrastructure development.
-Despite the growing reliance on tech giants, limited research has comprehensively examined the dynamics of such partnerships, particularly in the context of federal procurement contracts.
+Among the key in government procurement, large technology firms like Microsoft have become vital partners in IT supporting.
+Despite the growing reliance on tech giants, limited research examines the dynamics of such partnerships, particularly in the context of federal procurement contracts.
 
-This paper addresses this gap by conducting a detailed analysis of federal procurement contracts awarded to Microsoft between 2020 and September 2024.
-Using publicly available data from the Investigative Journalism Foundation, the study explores trends in contract values, contract types, and Microsoft’s relationships with key government buyers.
-Emphasis is placed on identifying patterns in contract allocation and the implications of Microsoft's dominance in high-value contracts.
+This paper aims to address this gap by analyzing federal procurement contracts awarded to Microsoft from 2020 up to September 2024.
+Using publicly available data from Investigative Journalism Foundation [@IJF], the study focusing on contract values, categories, and Microsoft’s relationships with key government buyers.
 
 The main estimand in this study is the amount of federal procurement contracts awarded to Microsoft.
-This metric is analyzed across multiple dimensions, including contract categories, duration, and partnerships, to quantify Microsoft's role in federal procurement comprehensively.
-The findings reveal that Microsoft plays a dominant role in IT infrastructure and modernization, securing the majority of high-value contracts and establishing consistent partnerships with departments such as Employment and Social Development Canada and National Defence.
-These insights highlight the broader implications of the private tech sector’s influence on public sector procurement, raising questions about competitiveness, dependency, and strategic alignment.
+This metric is analyzed across multiple dimensions, including contract categories, duration, and partnerships, to quantify Microsoft's role in federal procurement.
+A Bayesian modeling approach is employed to predict future trends in the contract amounts, integrating prior knowledge with observed data to address uncertainty and provide more robust estimates.
+The findings reveal that Microsoft plays a dominant role in IT infrastructure and modernization, securing the majority of high-value contracts and establishing consistent partnerships with departments such as Employment and Social Development Canada and National Defense.
+These findings highlight the broader implications of the private tech sector’s influence on public sector procurement, raising questions about competitiveness, dependency, and strategic alignment.
 
@@
 
-The structure of the paper is as follows: [@sec-data] outlines the data sources and variables considered, followed by the model setup in [@sec-modset] and justification.
+The structure of the paper is as follows: [@sec-data] outlines the data sources and variables considered, followed by the model setup and justificationa in [@sec-mod].
 The results in [@sec-result] presents the key findings of the analysis, with a discussion on the implications.
-[@sec-discussion] then discusses potential limitations and suggestions for future research.
-[@sec-appx] provides additional detailed information about the data, model and methodology.
+[@sec-discussion] then discusses potential limitations of data, model and suggestions for future research.
+[@sec-data-details] provides additional data detailed informations about data, model in [@sec-model-details] and methodology in [@sec-meth-details].
 
@@
 PhaseDays take the absolute value of PreparatoryPhase because of the minor effect of either prepare before contract being awarded or prepare after contract being awarded.
-Last but not least, I created a new variable called ContractType, which categorizes the contracts based on keywords.
+I also created new variables called ContractType and BuyerCleaned, which categorizes the most common buyers and contracts based on keywords.
 
@@
 So, I filter out the contracts before 2020 to focus on the records between 2020 and September 2024 as desired.
+By summary the data set, I noticed a significant outlier with value of $102,260,150, which largely impact the outcome.
+So, I decide to drop this entry.
 Then I check the character type variables, Supplier and Contract, using `table()` function for the purpose of manipulating input error.
@@
   ) %>%
-  filter(Count > 15) %>%
+  filter(Count > 25) %>%
   arrange(desc(Count))  # Arrange in descending order based on Count
@@
 
-**ContractDays** is a numeric variable which is the days between StartDate and EndDate.
+**ContractDays** is a numeric variable which is the days between $StartDate$ and $EndDate$.
 [@fig-contract] shows the distribution of performance period.
@@
 
+The correlation matrix in [@fig-cor] visualizes the relationships between key numerical predictor variables and the outcome variable, $Amount$.
+[@fig-cor] shows correlations include a weak negative correlations between $PreparatoryPhase$ and $ContractDays$; and weak positive correlations between predictors and $Amount$.
+
 ```{r}
 #| label: fig-cor
-#| fig-cap: This correlation matrix visualizes the relationships between key numerical predictor variables and the outcome variable, Amount.
-The figure shows correlations include a weak negative correlations between PreparatoryPhase and ContractDays; and weak positive correlations between predictors and Amount.
+#| fig-cap: Correlation matrix showing the relationships between numerical predictor variables and the outcome Amount variable
 #| echo: false
@@
 
-```{r}
-#| label: fig-box
-#| fig-cap: The Boxplots shows the relationship of ContractDays and PhaseDays with ContractType. 
-#| fig-subcap: ["Contract Duration in each Contract Type","Phase Duration in each Contract Type"]
-#| echo: false
-#| warning: false
-#| layout-nrow: 2
-
-# Clean and transform the data
-data <- cleaned_data %>%
-  mutate(
-    ContractType = gsub("Related", "", ContractType), # Remove "Related"
-    ContractType = str_wrap(gsub("/", " or ", ContractType), width = 10) # Wrap long labels
-  )
-
-# First graph: ContractType vs. ContractDays
-ggplot(data, aes(x = ContractType, y = ContractDays)) +
-  geom_boxplot() +
-  theme_minimal() +
-  labs(x = "Contract Type", y = "Contract Duration") +
-  theme(axis.text.x = element_text(size = 6))
-
-# Second graph: ContractType vs. PhaseDays
-ggplot(data, aes(x = ContractType, y = PhaseDays)) +
-  geom_boxplot() +
-  theme_minimal() +
-  labs(x = "Contract Type", y = "Phase Duration") +
-  theme(axis.text.x = element_text(size = 6))
-
-```
-
-# Model
+# Model {#sec-mod}
 
@@
 
+The boxplot @fig-box-1 shows the Contract Duration across various Contract Types.
+As we can see in @fig-box-1, Computer Services and Consulting contracts tend to have longer durations.
+This suggests these contracts often involve more complex or ongoing services that require extended timelines.
+In contrast, License/Maintenance Fees and Software contracts show shorter durations, with tighter interquartile ranges and lower medians, indicating that these contracts are likely less complex and are typically limited to a defined time period.
+
+The boxplot @fig-box-2 shows the Phase Duration for each Contract Type.
+@fig-box-2 reveals that certain contract types, such as Consulting and Distributed Computing Environment, have longer preparatory phases, with a higher median and a larger spread.
+This suggests these contracts require more time for negotiation, planning, or administrative procedures before the actual work begins.
+On the other hand, Computer Services and License/Maintenance Fees contracts exhibit shorter preparatory phases, as evidenced by the lower medians and narrower interquartile ranges in @fig-box-2.
+This suggests that these contracts are more straightforward and may require less preparation before implementation.
+```{r}
+#| label: fig-box
+#| fig-cap: The Boxplots shows the relationship of ContractDays and PhaseDays with ContractType. 
+#| fig-subcap: ["Contract Duration in each Contract Type","Phase Duration in each Contract Type"]
+#| echo: false
+#| warning: false
+#| layout-nrow: 2
+
+# Clean and transform the data
+data <- cleaned_data %>%
+  mutate(
+    ContractType = gsub("Related", "", ContractType), # Remove "Related"
+    ContractType = str_wrap(gsub("/", " or ", ContractType), width = 10) # Wrap long labels
+  )
+
+# First graph: ContractType vs. ContractDays
+ggplot(data, aes(x = ContractType, y = ContractDays)) +
+  geom_boxplot() +
+  theme_minimal() +
+  labs(x = "Contract Type", y = "Contract Duration") +
+  theme(axis.text.x = element_text(size = 6))
+
+# Second graph: ContractType vs. PhaseDays
+ggplot(data, aes(x = ContractType, y = PhaseDays)) +
+  geom_boxplot() +
+  theme_minimal() +
+  labs(x = "Contract Type", y = "Phase Duration") +
+  theme(axis.text.x = element_text(size = 6))
+```
+
 ```{r}
 #| label: fig-date
-#| fig-cap: The heatmap ....
+#| fig-cap: Heatmap of Contracts by Four-Month Periods (2020–2024).
+The heatmap shows the frequency of federal procurement contracts awarded to Microsoft in each four-month period, with darker colors representing higher contract frequencies.
 #| echo: false
@@
 
-# Count the number of contracts in each four-month period
+# Count the number of contracts for each four-month period and year
 contract_counts <- data %>%
-  group_by(Period) %>%
-  summarise(Count = n())
+  mutate(Year = as.factor(Year)) %>%
+  group_by(Year, Period) %>%
+  summarise(Count = n(), .groups = "drop")
 
 # Heatmap plot
-ggplot(contract_counts, aes(x = Period, y = Count, fill = Count)) +
-  geom_tile() +
-  scale_fill_gradient(low = "lightblue", high = "darkblue") +
-  labs(x = "Year-Period", y = "Frequency (Contract Count)", title = "Heatmap of Contracts by Four-Month Periods") +
-  theme_minimal() +
-  theme(axis.text.y = element_text(angle = 0))  # Keep y-axis labels readable
+ggplot(contract_counts, aes(x = Period, y = Year, fill = Count)) +
+  geom_tile(color = "white") +  # Add color between tiles for better visibility
+  scale_fill_gradient(low = "lightblue", high = "darkblue") +  # Adjust color scale
+  labs(x = "Four-Month Period", y = "Year", title = "Heatmap of Contracts by Four-Month Periods") +
+  theme_minimal()
 ```
@@
 ## Model
-R-squared, adjusted R-squared
-
-tests (e.g. VIF, AIC, BIC).
 
@@
 
-# Additional data details
+# Additional data details {#sec-data-details}
 
@@
 
-# Methodology
+# Methodology {#sec-meth-details}
 

### 2024-12-03 paper/paper.qmd
Update Data Section

@@
 
-More information about solicitation method can be found in @sec-appx.
+More information about solicitation method can be found in @sec-meth-details.
 
@@
 Then I rename key variables from raw data for clarity.
-The key variables of interest in this analysis include Contract, StartDate, AwardDate, EndDate, Buyer, and Amount.
+The key variables of interest in this analysis include $Contract$, $StartDate$, $AwardDate$, $EndDate$, $Buyer$, and $Amount$.
 
-For any missing data (NA) of StartDate, I infer the StartDate from the AwardDate to replace the missing value as the value of AwardDate.
-Then I standardize the date format of all date related datas for proper parsing by using `mdy()` function and change the dollar format of amount into numeric numbers for future calculation purposes.
-Due to the error with unknown reason that EndDate is before StartDate or AwardDate, those entries were excluded from this paper.
+I standardize the date format of all date related datas for proper parsing by using `mdy()` function and change the dollar format of amount into numeric numbers for future calculation purposes.
+Due to the error with unknown reason that $EndDate$ is before $StartDate$ or $AwardDate$, those entries were excluded from this paper.
 
-PreparatoryPhase, PhaseDays, and ContractDays variables were created for the purpose of analyzing preparatory phase and performance phase corresponding to each contract.
-PreparatoryPhase calculates the days from AwardDate to StartDate.
-ContractDays calculates the days from StartDate to EndDate.
-PhaseDays take the absolute value of PreparatoryPhase because of the minor effect of either prepare before contract being awarded or prepare after contract being awarded.
-I also created new variables called ContractType and BuyerCleaned, which categorizes the most common buyers and contracts based on keywords.
+$PreparatoryPhase$, $PhaseDays$, and $ContractDays$ variables were created for the purpose of analyzing preparatory phase and performance phase corresponding to each contract.
+$PreparatoryPhase$ calculates the days from $AwardDate$ to $StartDate$.
+$ContractDays$ calculates the days from $StartDate$ to $EndDate$.
+$PhaseDays$ take the absolute value of $PreparatoryPhase$ because of the minor effect of either prepare before contract being awarded or prepare after contract being awarded.
+I also created new variables called $ContractType$ and $BuyerCleaned$, which categorizes the most common buyers and contracts based on keywords.
 
@@
 So, I filter out the contracts before 2020 to focus on the records between 2020 and September 2024 as desired.
+Also, I filtered missing data (NA) of $StartDate$, which only has one entry in this dataset.
 By summary the data set, I noticed a significant outlier with value of $102,260,150, which largely impact the outcome.
 So, I decide to drop this entry.
-Then I check the character type variables, Supplier and Contract, using `table()` function for the purpose of manipulating input error.
+Then I check the character type variables, $Supplier$ and $Contract$, using `table()` function for the purpose of manipulating input error.
 I standardize the supplier's name which combine the similar name due to capitalization or some notation differences, and format them with first letter of each word capitalized.
@@
 **Buyer** is a character-type variable representing the federal government department or agency who initiate and award the contract.
-[@tbl-buyer] shows the top 10 buyer's name who has over 50 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors.
+[@tbl-buyer] shows the top 10 buyer's name who has over 25 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors.
 The table highlights that Employment and Social Development Canada (ESDC) and National Defence account over 50% of Microsoft's total awards.
-[CALCULATED 52.02%, BUT THINKING IF THIS SHOULD BE ADDED TO THE TABLE]
 

### 2024-12-03 paper/paper.qmd
Update version

@@
 A Bayesian modeling approach is employed to predict future trends in the contract amounts, integrating prior knowledge with observed data to address uncertainty and provide more robust estimates.
-The findings reveal that Microsoft plays a dominant role in IT infrastructure and modernization, securing the majority of high-value contracts and establishing consistent partnerships with departments such as Employment and Social Development Canada and National Defense.
+The findings reveal that Microsoft plays a dominant role in IT infrastructure and modernization, securing the majority of high-value contracts and establishing consistent partnerships with departments such as Employment and Social Development Canada and National Defence.
 These findings highlight the broader implications of the private tech sector’s influence on public sector procurement, raising questions about competitiveness, dependency, and strategic alignment.
@@
 
-The structure of the paper is as follows: [@sec-data] outlines the data sources and variables considered, followed by the model setup and justificationa in [@sec-mod].
+The structure of the paper is as follows: [@sec-data] outlines the data sources and variables considered, followed by the model setup and justifications in [@sec-mod].
 The results in [@sec-result] presents the key findings of the analysis, with a discussion on the implications.
 [@sec-discussion] then discusses potential limitations of data, model and suggestions for future research.
-[@sec-data-details] provides additional data detailed informations about data, model in [@sec-model-details] and methodology in [@sec-meth-details].
+The appendix provides additional detailed information about models in [@sec-model-details] and methodology in [@sec-meth-details].
 
@@
 More information about solicitation method can be found in @sec-meth-details.
-
-\newpage
 
@@
 This variable is need to represent the financial magnitude of Microsoft with the federal government and measurement of evaluating procurement dynamics.
-[@fig-amount] illustrates the annual total amounts of federal procurement contracts awarded to Microsoft from 2020 to 2024 (up to September).
-The histogram reveals a generally increasing trend of total value of each year's contract.
+[@fig-amount] is a histogram, which illustrates the annual total amounts of federal procurement contracts awarded to Microsoft from 2020 to 2024 (up to September).
+[@fig-amount] shows a generally increasing trend of total value of each year's contract.
 This may reflect the growing of technology in government operations, including software, cloud solutions, and IT infrastructure.
@@
 #| label: fig-amount
-#| fig-cap: The histogram shows the distribution of the total amount($million) of each year's federal contract (since 2020).
+#| fig-cap: The histogram shows the distribution of the total amount ($million) of each year's federal contract (since 2020).
 #| echo: false
@@
   theme_minimal() +
-  theme(axis.text.y = element_text(size = 10), axis.text.x = element_text(size = 10))
+  theme(
+    axis.text.y = element_text(size = 8), 
+    axis.title.y = element_text(size = 10),
+    axis.title.x = element_text(size = 10),
+    axis.text.x = element_text(size = 8),
+    plot.title = element_text(size = 12)
+    )
 
 ```
+
+\newpage
 
@@
 In this analysis, several key predictors were identified to evaluate procurement data effectively.
-The first variable, **ContractType**, is a character-type variable that identifies each procurement contract awarded to Microsoft into Categories.
-[@tbl-contract]
+
+**ContractType**, is a character-type variable that identifies each procurement contract awarded to Microsoft into Categories.
+License/Maintenance Fees has the highest number of contracts (495) in [@tbl-contract], suggesting the recurring need for software and system maintenance.
+Computer Services has the highest average contract value (\$1,763,432) in [@tbl-contract], reflecting the complexity and scope of IT services.
+Database Access Services also shows significant average values (\$1,680,453), indicating the importance of large-scale data systems in government operations.
 
@@
 **Buyer** is a character-type variable representing the federal government department or agency who initiate and award the contract.
-[@tbl-buyer] shows the top 10 buyer's name who has over 25 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors.
+[@tbl-buyer] shows the buyer who has over 25 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors.
 The table highlights that Employment and Social Development Canada (ESDC) and National Defence account over 50% of Microsoft's total awards.
@@
 #| label: tbl-buyer
-#| tbl-cap: The table summarizes top 10 buyer's name who has over 25 contracts with Microsoft in total with contract count, total dollar amount, and average dollar amount per contract. 
+#| tbl-cap: The table summarizes top 10 buyers with contract count, total dollar amount, and average dollar amount per contract. 
 #| echo: false
@@
 #| message: false
-#| fig.width: 4    # Width of the figure in inches
-#| fig.height: 3   # Height of the figure in inches
 
@@
 ```
+
+\newpage
 
@@
 ```
+More information about the data analysis can be found in @sec-result.
+
+\newpage
 
@@
 ### Model Justification
-This paper expect that the factors - such as ContractType, Buyer, ContractDays and PhaseDays - will influence the outcome of contract amount. In particular, ...
+This paper expect that the factors - such as $ContractType$, $Buyer$, $ContractDays$ and $PhaseDays$ - will influence the outcome of contract amount.
+In particular, $ContractType$ captures the service or product, such as software licenses or consulting services, which significantly influences contract value.
+$Buyer$ reflects the federal department or agency awarding the contract, as some departments have higher budgets and strategic priorities that result in larger contracts.
+$ContractDays$ and $PhaseDays$ quantify the duration and preparatory timelines of contracts, providing insight into their scope and complexity.
 
-In this paper, I compared three different modeling approaches to examine the factors influencing the federal procurement contract amount: linear regression, random forest, and Bayesian regression.
+In this paper, I compared linear regression and Bayesian regression to examine the factors influencing the federal procurement contract amount.
 Each of these models offers different advantages and trade-offs.
@@
 
-This Bayesian linear regression model balances complexity with interpretability, and its flexibility allows us to draw meaningful observations from the data while quantifying uncertainty around the estimates, especially when the sample size is not large or when the predictors have a complex relationship with the outcome.
-[TOO GENERAL, CHANGE]
+This Bayesian regression model balances complexity with interpretability, and its flexibility with quantifying uncertainty around the estimates, especially when the sample size is not large and the predictors have a complex relationship with the outcome.
 
@@
 While the model presents some valuable observations of Microsoft’s federal procurement contracts, there are some model limitations including prior function selection, data quality, potential for endogeneity, generalizability, and unobserved variables.
-These limitations will be discussed more detailed in [@sec-discussion].
+These limitations will be discussed more detailed in [@sec-discussion-model].
 
@@
 The boxplot @fig-box-2 shows the Phase Duration for each Contract Type.
-@fig-box-2 reveals that certain contract types, such as Consulting and Distributed Computing Environment, have longer preparatory phases, with a higher median and a larger spread.
+@fig-box-2 highlights Consulting and Distributed Computing Environment have longer preparatory phases, with a higher median and a larger spread.
 This suggests these contracts require more time for negotiation, planning, or administrative procedures before the actual work begins.
@@
 This suggests that these contracts are more straightforward and may require less preparation before implementation.
+
 ```{r}
@@
 ```
+
+The heatmap @fig-date shows the distribution of federal procurement contracts awarded to Microsoft over four-month periods from 2020 to 2024.
+The frequency tends to drop in the September–December period, it suggests the budgetary constraints or the fiscal year-end.
+A noticeable decline occurs in 2024 because the data only updates until September.
+The darker shades during early and mid-year in @fig-date align with expected procurement trends, where departments allocate funds proactively before the fiscal year closes.
+Understanding this distribution provides critical insights into procurement planning and Microsoft's role during different periods.
 
@@
 
+\newpage
+
+The scatter plot @fig-mobilization shows the relationship between the Award Date and the Start Date of federal procurement contracts awarded to Microsoft.
+Most points lie on the diagonal line indicating the numerous contracts commence immediately after being awarded.
+This is particularly common for time-sensitive contract types, such as License/Maintenance Fees Related.
+Contracts in 2020–2021 tend to exhibit a wider spread in start dates, possibly reflecting a slower procurement process during those years.
+More recent contracts in @fig-mobilization cluster closer to the diagonal, suggesting streamlined processes or more immediate implementation needs.
+
 ```{r}
 #| label: fig-mobilization
-#| fig-cap: The scatter plot shows ...
+#| fig-cap: The scatter plot shows the relationship between Award Date and Start Date for the top 10 federal procurement contracts types.
+Each point represents a contract, with colors indicating different contract types.
+The plot helps to identify trends in the timing between contract award and commencement.
 #| echo: false
@@
 #| message: false
-#| 
-ggplot(data, aes(x = AwardDate, y = StartDate)) +
-  geom_point(aes(color = Contract)) +
+
+ggplot(
+  data, 
+  aes(x = AwardDate, y = StartDate)) +
+  geom_point(aes(color = ContractType)) +
   labs(x = "Award Date", y = "Start Date", title = "Award Date vs Start Date") +
-  theme(legend.position = "none")  # This removes the legend
+  theme(
+    legend.position = "right",
+    legend.text = element_text(size = 6),      # Adjust legend label size
+    legend.title = element_text(size = 8)    # Adjust legend title size
+    )  # This removes the legend
 ```
+
+The scatter plot @fig-performance shows the diversity in contract durations across procurement categories.
+Most contracts ends within one year of their start dates, suggesting that most federal contracts are short-term.
+A smaller subset of contracts extends beyond three years, typically associated with large-scale projects or ongoing support agreements.
+@fig-performance also highlights that the most contract ends in the year of 2026.
+A few contracts with over 5 years stand out in 2021, suggesting that longer durations are prevalent in service-oriented categories.
 
@@
 #| label: fig-performance
-#| fig-cap: The scatter plot shows ....
+#| fig-cap: The scatter plot shows the relationship between Start Date and End Date for the top 10 Microsoft’s federal procurement contracts types.
+Each point represents a contract, and the color reflects the contract type.
+This visualization provides insights into the duration of the contracts and any deviations from expected timelines.
 #| echo: false
@@
 #| message: false
-#| 
-ggplot(data, aes(x = StartDate, y = EndDate)) +
-  geom_point(aes(color = Contract)) +
+
+ggplot(
+  data, 
+  aes(x = StartDate, y = EndDate)) +
+  geom_point(aes(color = ContractType)) +
   labs(x = "Start Date", y = "End Date", title = "Start Date vs End Date") +
-  theme(legend.position = "none")  # This removes the legend
+  theme(
+    legend.position = "right",
+    legend.text = element_text(size = 6),      # Adjust legend label size
+    legend.title = element_text(size = 8)      # Adjust legend title size
+  )
 ```
 
-## Model
+\newpage
 
-Our results are summarized in @tbl-modelresults.
+## Model Summary
+
+The results are summarized in @tbl-modelresults, including R², Log-Likelihood (Log.Lik.), Expected Log Pointwise Predictive Density (ELPD), Leave-One-Out Information Criterion (LOOIC), Watanabe-Akaike Information Criterion (WAIC), and Root Mean Square Error (RMSE).
 
@@
 #| eval: true
-#| tbl-cap: "Explanatory models of contract amount based on ContractType, Buyer, ContractDays, PhaseDays and Month"
+#| tbl-cap: "Explanatory models of contract amount based on ContractType, Buyer, ContractDays and PhaseDays"
 #| warning: false
@@
 
+The Full Bayesian Model has an $R^2$ of 0.289, indicating that approximately 28.9% of the variance in the contract amount is explained by the model.
+In comparison, the Reduced Bayesian Model explains only 8.2% of the variance in the contract amount, suggesting that the inclusion of additional predictors (specific buyers and contract names) improves the model’s explanatory power.
+However, both models have very low adjusted $R^2$, which suggests that the model is not capturing much of the unexplained variance in the data.
+
+The Full Bayesian Model has a Log-Likelihood of $-15,351.795$ and an ELPD of $-15,519.3$, while the Reduced Model has a Log-Likelihood of $-15,398.022$ and an ELPD of $-15,444.8$.
+These values suggest that the Full Bayesian Model fits the data slightly better than the Reduced Model, as it produces higher Log-Likelihood and ELPD values.
+
+The Full Bayesian Model has a lower LOOIC ($31,038.7$) and WAIC ($30,992.3$) compared to the Reduced Model, indicating that the Full Bayesian Model is better at out-of-sample predictive performance.
+Lower values of LOOIC and WAIC are indicative of better model fit and predictive accuracy.
+
+The Full Bayesian Model has a lower RMSE ($2,714,467.04$) compared to the Reduced Model ($2,971,135.35$), indicating better predictive performance.
+The lower RMSE value suggests that the Full Bayesian Model is more accurate in predicting the total contract amount.
+
 \newpage
@@
 
-## Data
+## Contract Complexity and Cost
+The positive association between $ContractDays$ and contract amount suggests that longer contracts are typically more expensive, reflecting the increased scope, complexity, and resource requirements of these agreements.
+This is particularly relevant for IT-related contracts, which often involve extensive development, maintenance, and support services.
 
-The completeness and reliability of the data. For the missing or erroneous records, despite the data cleaning process, introduce bias and distort the findings.
+## Negotiation and Efficiency
+The negative association between $PhaseDays$ and contract amount indicates that longer preparatory phases may lead to reduced costs, potentially due to more thorough planning, negotiation, or budgetary constraints.
+This insight can inform government strategies for managing procurement timelines to achieve cost efficiency.
 
-Since the data of the procurement contracts awarded to Microsoft in Canada is between 2020 and 2024 (up to September), the analysis provides a snapshot of recent procurement trends during this period.
-However, the data do not capture longer-term patterns or emerging shifts in procurement policies.
+## Sector-Specific Effects 
+The significant variation in contract amounts across different contract types (e.g., Consulting, Software) and $buyers$ (e.g., National Defence, Employment and Social Development Canada) underscores the importance of sector-specific factors in procurement dynamics.
+Departments with higher budgets or critical mandates, such as National Defence, tend to award larger contracts, highlighting their influence in shaping overall procurement trends.
 
-## Limitations
+## Model Performance and Limitations {#sec-discussion-model}
+
+The Bayesian model has low adjusted $R^2$ values.
+This means the model explains only a small portion (28.9%) of the variance in contract amounts.
+This suggests that important predictors may be missing.
+Also, missing or erroneous values in variables like $ContractDays$ or $PhaseDays$ could introduce bias and reduce the reliability of the results.
 
@@
 
-The model assumes a linear relationship between the independent variables and the dependent variable.
-However, procurement dynamics may be more complex, with non-linear interactions between variables.
-The model categorizes contract types and buyers into broad groups, which may oversimplify the true diversity of contracts awarded.
+The model assumes a linear relationship between the independent variables and the dependent variable, which oversimplify the dynamics of federal procurement.
+The model categorizes contract types and buyers into broad groups, which also oversimplify the true diversity of contracts awarded.
+Future studies could explore non-linear models, such as splines or machine learning approaches, to capture more complex patterns.
 
@@
 
-## Model Selection
+## Alternative Models Considered
+Linear regression model was evaluated in the paper.
+While the linear regression model was straightforward and interpretable, it lacked the flexibility to incorporate uncertainty in parameter estimates.
+The Bayesian regression model was chosen as a middle ground, balancing flexibility, interpretability, and the ability to quantify uncertainty.
+However, the limitations observed in model performance suggest that future work could benefit from integrating features of these alternative approaches.
+
+More results about these two models can be found in @sec-model-details.
 
@@
 
-# Additional data details {#sec-data-details}
-
-# Model details {#sec-model-details}
+# Additional Model details {#sec-model-details}
 
@@
 
-### Posterior predictive check
+### Posterior Predictive Check
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check, comparing the observed data, $y$ and the posterior predictive replicated data $y_{rep}$.
+This shows how well the Bayesian model replicates the observed data.
+@fig-ppcheckandposteriorvsprior-1 displays a pronounced spike at 0, where $y$ has a significantly higher frequency of values at $0e+00$ compared to the predictions from the model.
+This discrepancy indicates that the model fails to accurately capture the concentration of contracts with a value of exactly zero or near-zero, potentially due to missing covariates or unmodeled structural factors influencing the occurrence of zero-value contracts.
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+The predicted distribution $y_{rep}$ in @fig-ppcheckandposteriorvsprior-1 aligns more closely with the observed data at larger values, suggesting that the model performs well in capturing the behavior of non-zero contract amounts.
+However, the inability to replicate the sharp peak at 0 highlights a significant limitation in the model’s ability to account for zero-value contracts or extreme sparsity in the data.
+This suggests a potential need for refinements, such as incorporating a zero-inflated component or additional predictors to explain this phenomenon.
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
+This shows the influence of the observed data on parameter estimates.
+For most parameters, the posterior distributions are narrower and shifted away from the priors, indicating that the data provides substantial information to update the prior beliefs.
+Some parameters, however, show minimal deviation from their priors, which may suggest either weak data informativeness or a lack of association with the response variable.
 
@@
 #| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+#| fig-cap: "Examining how the model fits and affected by the data"
+#| fig-subcap: ["Posterior Prediction Check", "Comparing the Posterior with the Prior"]
 
@@
   coord_flip() 
-
 ```
@@
 
-@fig-trace-1 is a trace plot. It shows... This suggests...
+@fig-trace-1 is a trace plot related to $ContractType$.
+Each chain shows the sampled values of the parameters over the iterations of the Markov Chain Monte Carlo (MCMC) process.
+The chains for all contract type parameters appear to mix well and fluctuate around a consistent range, indicating convergence.
+There is no evidence of trends or drifts in the chains, which suggests that the MCMC process effectively explores the posterior distribution.
 
-@fig-trace-2 is a trace plot. It shows... This suggests...
-
-@fig-rhat is a Rhat plot. It shows... This suggests...
+@fig-trace-2 is a trace plot associated with $Buyers$.
+Similar to @fig-trace-1, the plot shows the sampled parameter values for multiple MCMC chains over iterations.
+The chains in @fig-trace-2 for most buyer parameters exhibit good mixing and stability, with fluctuations around a consistent range.
+This indicates effective convergence for the majority of buyer-related parameters.
+A few parameters in @fig-trace-2 exhibit slightly slower mixing, but they still stabilize, suggesting no major convergence issues.
 
@@
 ```
+
+@fig-rhat is Rhat plot.
+An Rhat value of 1.0 indicates that the MCMC chains have converged, meaning they are sampling from the same posterior distribution.
+If the Rhat value exceeds 1.05, it signals potential issues with convergence, indicating that the chains might not have mixed properly or that there is insufficient sampling to reach stable parameter estimates.
+
+@fig-rhat-1 is a Rhat plot for the Full Bayesian model.
+The Rhat values for most parameters are around 1.0, which suggests that convergence has been achieved for the majority of parameters.
+There are no parameters with Rhat values exceeding 1.05, indicating that the model's convergence is sound.
+
+@fig-rhat-2 is a Rhat plot for the reduced Bayesian model, which shows similar results.
+All parameters having Rhat values close to 1.0, further confirming that convergence is achieved across both models.
+There are no parameters with Rhat values exceeding 1.05, which supports the reliability of sampling process.
+
 ```{r}
@@
 ```
+
+\newpage
+
 ## Linear Regression Model
 
+The linear regression model is the most straightforward and interpretable among the models considered.
+It assumes a linear relationship between the outcome variable, $Amount$, and predictor variables - $Contract$, $Buyer$, $ContractDays$, $PhaseDays$.
+This makes it easy to understand how each predictor influences the contract amount.
+In @tbl-linear, linear regression model achieved an R² value of 0.220, shows that only 22% of the variance in contract amounts is explained by the predictors.
+The Root Mean Square Error (RMSE) was approximately \$2.71 million for the full model, indicating substantial prediction errors.
+
+Linear regression models are computationally inexpensive and require fewer resources compared to more complex models like random forests or Bayesian regression.
+
 ```{r}
-#| label: tbl-keyestimates
-#| tbl-cap: Summary of key model estimates for Microsoft, including coefficients for predictors like xxxx, with standard errors for each estimate.
+#| label: tbl-linear
+#| tbl-cap: Summary of key model estimates for Microsoft, including coefficients for numeric predictors like Contract Days and Phase Days, with standard errors for each estimate.
 Model performance statistics, such as sample size, R², and adjusted R², are also displayed.
@@
 # load data and model
-lm_model_full <- readRDS(here::here("models/lm_model_full.rds"))
 lm_model <- readRDS(here::here("models/lm_model.rds"))
-model_list <- list("Linear Model_Full" = lm_model_full, "Linear Model" = lm_model)
 
 modelsummary(
-  model_list,
+  list("Linear Model" = lm_model),
   output = "kableExtra", 
@@
 
-## Random Forest Model
-
-```{r}
-#| label: tbl-rfkeyestimates
-#| tbl-cap: Summary of key model estimates for Microsoft, including coefficients for predictors like xxxx, with standard errors for each estimate.
-Model performance statistics, such as sample size, R², and adjusted R², are also displayed.
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# load data and model
-rf_model<- readRDS(here::here("models/rf_model.rds"))
-
-# Summary of the Random Forest model
-summary(rf_model, output = "kableExtra")
-
-# # Error rates
-# rf_model$err.rate  # Error rates for each class (classification)
-# rf_model$oob.err    # Out-of-bag error estimate
-
-```
+\newpage
 
@@
 
+# Survey
+
+**Vendor Relationship Dynamics in Federal Procurement**
+
+**Introduction**
+
+We are conducting a survey to better understand the dynamics of vendor relationships in federal procurement processes, particularly with major technology providers like Microsoft.
+Your responses will be kept confidential and used only for research purposes.
+
+If you have any questions about this survey, please contact us: deyi.kong\@mail.utoronto.ca
+
+**Section 1: General Information**
+
+1. **What is your role in the procurement process?**  
+   - Procurement Officer  
+   - Project Manager  
+   - Technical Specialist  
+   - Other (please specify): _________
+
+2. **How many years of experience do you have in procurement?**  
+   - Less than 1 year  
+   - 1–5 years  
+   - 6–10 years  
+   - Over 10 years
+
+3. **Which sector or department do you primarily represent?**  
+   - National Defence  
+   - Employment and Social Development Canada  
+   - Natural Resources Canada  
+   - Other (please specify): _________
+
+**Section 2: Vendor Interaction and Communication**
+
+4. **How would you describe your department’s relationship with Microsoft?**  
+   - Excellent  
+   - Good  
+   - Neutral  
+   - Poor  
+   - Very Poor
+
+5. **How frequently does your department interact with Microsoft representatives?**  
+   - Weekly  
+   - Monthly  
+   - Quarterly  
+   - Annually  
+   - As needed
+
+6. **What is the primary mode of communication with Microsoft?**  
+   - Email  
+   - Phone calls  
+   - In-person meetings  
+   - Online meetings (e.g., Zoom, Teams)  
+   - Other (please specify): _________
+
+**Section 3: Relationship Dynamics**
+
+7. **To what extent does Microsoft involve your department in the decision-making process for solutions or services?**  
+   - Always  
+   - Often  
+   - Sometimes  
+   - Rarely  
+   - Never
+
+8. **How would you rate Microsoft’s responsiveness to your department’s needs or concerns?**  
+   - Very Responsive  
+   - Responsive  
+   - Neutral  
+   - Unresponsive  
+   - Very Unresponsive
+
+9. **Does your department feel valued as a client by Microsoft?**  
+   - Strongly Agree  
+   - Agree  
+   - Neutral  
+   - Disagree  
+   - Strongly Disagree
+
+**Section 4: Impact of the Vendor Relationship**
+
+10. **What is the biggest benefit of your department's relationship with Microsoft?**  
+    - Access to cutting-edge technology  
+    - Reliability of services  
+    - Cost efficiency  
+    - Strategic partnership opportunities  
+    - Other (please specify): _________
+
+11. **What challenges have you experienced in working with Microsoft? (Select all that apply)**  
+    - High costs  
+    - Limited customization options  
+    - Slow response times  
+    - Difficulty with contract negotiations  
+    - Other (please specify): _________
+
+12. **How likely are you to recommend Microsoft as a vendor to other departments?**  
+    - Very Likely  
+    - Likely  
+    - Neutral  
+    - Unlikely  
+    - Very Unlikely
+
+**Section 5: Suggestions for Improvement**
+
+13. **What changes would you recommend for improving the relationship with Microsoft?**  
+    (Open text field)
+
+14. **Are there alternative vendors you believe could meet your department’s needs better than Microsoft? If so, please specify.**  
+    (Open text field)
+
+-----
+
+Thank you for your participation in this survey.
+Your feedback is invaluable in understanding and improving vendor relationships in federal procurement.
+If you have additional comments or would like to receive updates on the research findings, please indicate below.
+   - [\ ] Yes, I’d like to receive updates.  
+   - [\ ] No, I do not wish to receive updates.
+
+\newpage
+
 # References

### 2024-12-03 models/plumber.R

@@ -1,0 +1,82 @@
+#### Preamble ####
+# Purpose: 
+# Author: Deyi Kong
+# Date: December 2nd, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `plumber`, `rstanarm`, and `tidyverse` packages must be installed
+# Any other information needed? No
+
+# Load libraries
+library(plumber)
+library(rstanarm)
+library(tidyverse)
+
+# Load model
+model <- readRDS("bayesian_model.rds")
+
+# Define model version
+version <- "0.0.1"
+
+# Define variables
+variables <- list(
+  ContractType = "The category of the contract (only License/Maintenance or Computer Services).",
+  BuyerCleaned = "The buyer of the contract (only Employment and Social Development Canada, or Global Affairs Canada).",
+  ContractDays = "The duration of the contract",
+  PhaseDays = "The duration of the preparatory phase."
+)
+
+#* @param ContractType
+#* @param BuyerCleaned
+#* @param ContractDays
+#* @param PhaseDays
+#* @get /predict_value
+
+predict_value <- function( 
+    ContractType = "License/Maintenance", 
+    BuyerCleaned = "Global Affairs Canada", 
+    ContractDays = 123, 
+    PhaseDays = 30
+    ) {
+  # Convert inputs to appropriate types
+  ContractType <- as.character(ContractType)
+  BuyerCleaned <- as.character(BuyerCleaned)
+  ContractDays <- as.numeric(ContractDays)
+  PhaseDays <- as.numeric(PhaseDays)
+  
+  #Prepare the payload as a data frame
+  payload <- data.frame(
+    ContractType = ContractType,
+    BuyerCleaned = BuyerCleaned,
+    ContractDays = ContractDays,
+    PhaseDays = PhaseDays
+  )
+  
+## Extract Posterior samples ##
+  # Convert to matrix for easier manipulation
+  posterior_samples <- as.matrix(model)
+  
+  # Define the generative process for prediction
+  alpha <- posterior_samples[, "(Intercept)"]
+  beta_ContractType <- posterior_samples[, "ContractType"]
+  beta_BuyerCleaned <- posterior_samples[, "BuyerCleaned"]
+  beta_ContractDays <- posterior_samples[, "ContractDays"]
+  beta_PhaseDays <- posterior_samples[, "PhaseDays"]
+  
+  # Compute the predicted value for the observation
[…]

### 2024-12-03 other/shiny_app/shiny.R

@@ -1,0 +1,151 @@
+#### Preamble ####
+# Purpose: This Shiny application provides an interactive analysis of federal procurement contracts.
+# Author: Deyi Kong
+# Date: December 3rd, 2024
+# Contact: deyi.kong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `shiny`, `ggplot2`, `DT`, `dplyr`, `rstanarm`, and `bayesplot` packages must be installed.
+# Any other information needed? Ensure that you are working within the `Procurement`.rproj.
+
+# Load necessary libraries
+library(shiny)
+library(ggplot2)
+library(DT)
+library(dplyr) 
+library(rstanarm)
+library(bayesplot)
+
+# Define the UI
+ui <- fluidPage(
+  titlePanel("Analysis of Federal Contracts"),
+  sidebarLayout(
+    sidebarPanel(
+      fileInput("file", "Upload Cleaned Data (CSV)", accept = ".csv"),
+      selectInput("plotType", "Select Plot Type", 
+                  choices = c("Yearly Contract Amounts", 
+                              "Trace Plot", 
+                              "Posterior Predictive Check")),
+      numericInput("titleSize", "Title Font Size", value = 14, min = 10, max = 30),
+      actionButton("refresh", "Refresh Plot"),
+      br(),
+      h3("Model Results"),
+      DTOutput("resultsTable")
+    ),
+    mainPanel(
+      plotOutput("mainPlot"),
+      h3("Description"),
+      textOutput("plotDescription")
+    )
+  )
+)
+
+# Define the server
+server <- function(input, output) {
+  # Reactive expression to load data
+  data <- reactive({
+    req(input$file)
+    df <- read.csv(input$file$datapath)
+    
+    # Print column names for debugging
+    print(colnames(df))
+    
+    # Simulate 'yrep' if missing
+    if (!("yrep" %in% colnames(df))) {
+      set.seed(123)
+      df$yrep <- rnorm(nrow(df), mean = mean(df$Amount), sd = sd(df$Amount))
+    }
+    
+    # Simulate 'Prior' if missing
+    if (!("Prior" %in% colnames(df))) {
+      df$Prior <- rnorm(nrow(df), mean = 0, sd = 1)
+    }
+    
+    # Simulate 'Posterior' if missing
+    if (!("Posterior" %in% colnames(df))) {
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Perform exploratory data analysis on cleaned data
 # Author: Deyi Kong
-# Date: November 24th, 2024
+# Date: December 2nd, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, 'janitor', 'here', and 'reshape2' packages
@@ -40,3 +40,47 @@ missing_data <- cleaned_data |>
   summarise(across(everything(), ~ sum(is.na(.))))
 missing_data
 
+# Sum the total amount group by year
+yearly_amount <- year_data %>%
+  group_by(Year) %>%
+  summarize(TotalAmount = sum(Amount, na.rm = TRUE)) %>%
+  mutate(TotalAmount = round(TotalAmount / 1000000)) # Convert to millions
+
+# Plot the amount in each year
+ggplot(yearly_amount, aes(x = as.factor(Year), y = TotalAmount)) +
+  geom_bar(stat = "identity", fill = "darkorange", color = "black", alpha = 0.7) +
+  labs(
+    title = "Total Contract Amounts by Year for Microsoft",
+    x = "Year",
+    y = "Total Amount ($million)"
+  ) +
+  theme_minimal() +
+  theme(axis.text.y = element_text(size = 10), axis.text.x = element_text(size = 10))
+
+# create plot for Distribution of Poll Scores by candidate
+ggplot(cleaned_data, aes(x = phase_days)) +
+  geom_histogram(
+    bins = 30,
+    fill = "salmon",
+    color = "black",
+    alpha = 0.7
+  ) +
+  labs(
+    x = "Preparatory Phase (day)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+# create plot for Distribution of Poll Scores by candidate
+ggplot(cleaned_data, aes(x = contract_days)) +
+  geom_histogram(
+    bins = 15,
+    fill = "steelblue",
+    color = "black",
+    alpha = 0.7
+  ) +
+  labs(
+    x = "Contract Performance Phase (day)",
+    y = "Frequency"
+  ) +
+  theme_minimal()

### 2024-12-03 scripts/06-model_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Run a linear regression model to predict 
 # Author: Deyi Kong
-# Date: December 2nd, 2024
+# Date: December 3rd, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, `rstanarm`, `arrow`, `randomForest` and 'here' packages must be installed
@@ -24,26 +24,17 @@ cleaned_data <- cleaned_data %>%
   )
 
 # full linear model for contract dollar amount based on various predictors
-lm_model_full <- lm(Amount ~ Contract + Buyer + ContractDays + PhaseDays + Month,
+lm_model <- lm(Amount ~ Contract + Buyer + ContractDays + PhaseDays,
                  data = cleaned_data
 )
 
-# linear model for contract dollar amount based on various predictors
-lm_model <- lm(Amount ~ ContractType + BuyerCleaned + ContractDays + PhaseDays + Month,
-                 data = cleaned_data
-               )
-
-# Random Forest model for regression (predicting a continuous outcome)
-rf_model <- randomForest(Amount ~ ContractType + BuyerCleaned + ContractDays + PhaseDays + Month,
-                         data = cleaned_data, ntree = 500)
-
 # full Bayesian model
 bayesian_model_full <- stan_glm(
-  Amount ~ Contract + Buyer + ContractDays + PhaseDays + Month,
+  Amount ~ Contract + Buyer + ContractDays + PhaseDays,
   data = cleaned_data,
   family = gaussian(),
-  prior = normal(location = 0, scale = 2, autoscale = TRUE),   # Prior for coefficients
-  prior_intercept = normal(location = 0, scale = 2, autoscale = TRUE),  # Prior for intercept
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),   # Prior for coefficients
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),  # Prior for intercept
   prior_aux = exponential(rate = 1, autoscale = TRUE),  # Prior for residual SD
   seed = 123,
   iter = 4000,  # Increase the total number of iterations
@@ -52,7 +43,7 @@ bayesian_model_full <- stan_glm(
 
 # Bayesian model
 bayesian_model <- stan_glm(
-  Amount ~ ContractType + BuyerCleaned + ContractDays + PhaseDays + Month,
[…]

### 2024-12-03 scripts/07-validation.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Validate the regression models through out of sample testing and RMSE
 # Author: Deyi Kong
-# Date: November 24th, 2024
+# Date: December 3rd, 2024
 # Contact: deyi.kong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse`, `here`, 'randomForest', and 'arrow' packages must be installed
@@ -14,7 +14,7 @@ library(here)
 library(caret)
 library(arrow)
 library(tibble)
-library(randomForest)
+library(rstanarm)
 
 # load data
 cleaned_data <- read_parquet(here::here("data/analysis_data/procurement_cleaned.parquet"))
@@ -37,43 +37,62 @@ train_data <- cleaned_data[train_index, ]
 test_data <- cleaned_data[-train_index, ]
 
 # linear model for microsoft contract dollar amount based on various predictors
-lm_model <- lm(Amount ~ ContractType + Buyer + ContractDays + PhaseDays,
-                 data = cleaned_data)
+lm_model <- lm(Amount ~ Contract + Buyer + ContractDays + PhaseDays,
+               data = cleaned_data)
 
-# Make predictions on the testing set
-predictions <- predict(lm_model, newdata = test_data)
+# Make predictions on the testing set using the linear model
+predictions_lm <- predict(lm_model, newdata = test_data)
 
 # Extract actual values for RMSE calculation
 actual_values <- test_data$Amount
 
-# Calculate RMSE (Root Mean Squared Error)
-rmse_value <- sqrt(mean((actual_values - predictions)^2, na.rm = TRUE))
+# Calculate RMSE (Root Mean Squared Error) for the linear model
+rmse_lm <- sqrt(mean((actual_values - predictions_lm)^2, na.rm = TRUE))
 
-# Store RMSE results in a tibble
-rmse_results <- tibble(
+# Save the linear model results into a parquet file
+lm_results <- tibble(
   Model = "Linear Regression",
-  RMSE = rmse_value
+  RMSE = rmse_lm,
+  Coefficients = list(coef(lm_model)),
+  Predictions = list(predictions_lm)
 )
 
-# Train a Random Forest model on the training set
-rf_model <- randomForest(Amount ~ ContractType + Buyer + ContractDays + PhaseDays, data = train_data, ntree = 500)
+# Save linear model results to a parquet file
[…]

### 2024-12-04 paper/paper.qmd
Refine Introduction and subtitle

@@
 title: "Procurement Dynamics: Analyzing Federal Contract Awards to Microsoft"
-subtitle: "Trends of Microsoft’s Future in Canadian Government from 2020"
+subtitle: "Microsoft's Dominance in IT Procurement and Relationships with Government"
 author: 
@@
 date-format: long
-abstract: "This paper analyzes federal procurement contracts awarded to Microsoft from 2020 to 2024, focusing on contract values, categories, and relationships between buyers and Microsoft.
-Multiple models were considered, with Bayesian modeling approach selected to predict Microsoft's contract amount.
-The analysis reveals Microsoft's dominant role in high-value contracts and its consistent partnerships with major government departments like National Defence.
-These findings provide informations evolving dynamics of federal procurement, raising questions about dependency and competitiveness."
+abstract: "This paper analyzes federal procurement contracts awarded to Microsoft from 2020 to 2024, focusing on contract values, categories, and relationships with key government buyers.
+A Bayesian modeling approach was employed to predict contract amounts.
+The study highlights Microsoft's significant role in securing high-value contracts and maintaining steady partnerships with major departments such as National Defence.
+These findings provide important trends in federal procurement, drawing attention to the potential risks of dependency on a single supplier and challenges to competition."
 format: pdf
@@
 
-Government procurement refers to the acquisition of goods, services, and construction services by the government, spanning a wide range of purchases from office supplies to materials and services needed for large-scale infrastructure projects [@gac].
-This process enables public sector innovation, economic stimulus and fulfills strategic objectives.
-Within this landscape, the relationship between major technology providers and governmental bodies has become increasingly prominent.
-Among the key in government procurement, large technology firms like Microsoft have become vital partners in IT supporting.
+Government procurement involves the acquisition of goods and services.
+The construction services by the government span a wide range of purchases, from office supplies to materials and services needed for large-scale infrastructure projects [@gac].
+This process enables public innovation, economic stimulus, and fulfills strategic objectives.
+Within this landscape, the relationship between technology providers and government operations has become increasingly prominent.
 Despite the growing reliance on tech giants, limited research examines the dynamics of such partnerships, particularly in the context of federal procurement contracts.
@@
 This paper aims to address this gap by analyzing federal procurement contracts awarded to Microsoft from 2020 up to September 2024.
-Using publicly available data from Investigative Journalism Foundation [@IJF], the study focusing on contract values, categories, and Microsoft’s relationships with key government buyers.
+Using publicly available data from the Investigative Journalism Foundation [@IJF], the study focuses on contract values, categories, and Microsoft’s relationships with key government buyers.
 
@@
 This metric is analyzed across multiple dimensions, including contract categories, duration, and partnerships, to quantify Microsoft's role in federal procurement.
+
 A Bayesian modeling approach is employed to predict future trends in the contract amounts, integrating prior knowledge with observed data to address uncertainty and provide more robust estimates.
-The findings reveal that Microsoft plays a dominant role in IT infrastructure and modernization, securing the majority of high-value contracts and establishing consistent partnerships with departments such as Employment and Social Development Canada and National Defence.
+The findings reveal that Microsoft plays a dominant role in IT infrastructure and modernization, securing most high-value contracts and establishing consistent partnerships with departments such as Employment and Social Development Canada and National Defence.
 These findings highlight the broader implications of the private tech sector’s influence on public sector procurement, raising questions about competitiveness, dependency, and strategic alignment.
@@
 This study contributes to understand the procurement landscape, providing valuable information for policymakers, industry stakeholders, and academics.
-It underscores the importance of transparency, strategic planning, and balancing innovation with equity in government contracts.
+The study  underscores the importance of transparency, strategic planning, and balancing innovation with equity in government contracts.
 

### 2024-12-04 paper/paper.qmd
Update Model Section

@@
 
+\newpage
+
 # Introduction
@@
 
-The dataset used in this analysis originates from Canadian governments of all levels post solicitations for companies to search for business opportunities, as documented in @ProcurementIntro, then collects data from each of these platforms and unifies into standard format.
+The dataset used in this analysis originates from Canadian governments of all levels post solicitations for companies to search for business opportunities, as documented in @ProcurementIntro.
+The data from these various platforms is then standardized to form a unified dataset for analysis.
 
@@
 The methodologies of federal solicitations are either competitive or non-competitive.
-Competitive contracts accept bids from all or some subset of suppliers.
-Non-competitive contracts do not, citing certain exceptions, including a “pressing emergency in which delay would be injurious to the public interest,” or when the estimated cost to the buyer is below certain thresholds defined for some goods and services.
-
-More information about solicitation method can be found in @sec-meth-details.
+Competitive contracts allow bids from all or a subset of suppliers, while non-competitive contracts are awarded without open bidding.
+Exceptions for non-competitive contracts include situations where delays could harm the public interest or when the cost to the buyer falls below certain thresholds for specific goods or services.
+
+More detailed information about solicitation methodology can be found in @sec-meth-details.
 
@@
 
-After imported the raw dataset using `read_csv` function, I first select specific columns to focus on relevant information towards this study by omitting redundant information, such as the contract link and same value of region.
-Then I rename key variables from raw data for clarity.
+After importing the raw dataset using `read_csv` function, I first select specific columns to focus on relevant information for this study by omitting redundant information, such as the contract link and the same value of region.
+Then I rename key variables from the raw data for clarity.
 The key variables of interest in this analysis include $Contract$, $StartDate$, $AwardDate$, $EndDate$, $Buyer$, and $Amount$.
 
-I standardize the date format of all date related datas for proper parsing by using `mdy()` function and change the dollar format of amount into numeric numbers for future calculation purposes.
-Due to the error with unknown reason that $EndDate$ is before $StartDate$ or $AwardDate$, those entries were excluded from this paper.
-
-$PreparatoryPhase$, $PhaseDays$, and $ContractDays$ variables were created for the purpose of analyzing preparatory phase and performance phase corresponding to each contract.
+I standardize the date format of all date-related data for proper parsing by using `mdy()` function and change the dollar format of the amount into numeric numbers for future calculation purposes.
+Due to the error with the unknown reason that $EndDate$ is before $StartDate$ or $AwardDate$, those entries were excluded from this paper.
+
+$PreparatoryPhase$, $PhaseDays$, and $ContractDays$ variables were created to analyze the preparatory phase and performance phase corresponding to each contract.
 $PreparatoryPhase$ calculates the days from $AwardDate$ to $StartDate$.
 $ContractDays$ calculates the days from $StartDate$ to $EndDate$.
-$PhaseDays$ take the absolute value of $PreparatoryPhase$ because of the minor effect of either prepare before contract being awarded or prepare after contract being awarded.
-I also created new variables called $ContractType$ and $BuyerCleaned$, which categorizes the most common buyers and contracts based on keywords.
+$PhaseDays$ takes the absolute value of $PreparatoryPhase$ because of the minor effect of either preparing before the contract is awarded or preparing after the contract is awarded.
+I also created new variables called $ContractType$ and $BuyerCleaned$, which categorize the most common buyers and contracts based on keywords.
 
 This dataset contains only two entries for 2019 federal procurement contracts awarded to Microsoft.
-So, I filter out the contracts before 2020 to focus on the records between 2020 and September 2024 as desired.
+So, I filtered out the contracts before 2020 to focus on the records between 2020 and September 2024 as desired.
 Also, I filtered missing data (NA) of $StartDate$, which only has one entry in this dataset.
-By summary the data set, I noticed a significant outlier with value of $102,260,150, which largely impact the outcome.
-So, I decide to drop this entry.
-Then I check the character type variables, $Supplier$ and $Contract$, using `table()` function for the purpose of manipulating input error.
-I standardize the supplier's name which combine the similar name due to capitalization or some notation differences, and format them with first letter of each word capitalized.
+By summary of the data set, I noticed a significant outlier with the value of $102,260,150, which largely impacted the outcome.
+So, I decided to drop this entry.
+Then I check the character type variables, $Supplier$ and $Contract$, using `table()` function to manipulate input error.
+I standardize the supplier's name which combines the similar name due to capitalization or some notation differences, and format them with the first letter of each word capitalized.
 
@@
 The main outcome variable of interest is the amount of federal procurement contracts awarded to Microsoft.
-This variable is need to represent the financial magnitude of Microsoft with the federal government and measurement of evaluating procurement dynamics.
+This variable is needed to represent the financial magnitude of Microsoft with the federal government and the measurement of evaluating procurement dynamics.
 [@fig-amount] is a histogram, which illustrates the annual total amounts of federal procurement contracts awarded to Microsoft from 2020 to 2024 (up to September).
-[@fig-amount] shows a generally increasing trend of total value of each year's contract.
-This may reflect the growing of technology in government operations, including software, cloud solutions, and IT infrastructure.
+[@fig-amount] shows a generally increasing trend of the total value of each year's contract.
+This may reflect the growth of technology in government operations, including software, cloud solutions, and IT infrastructure.
 The amount in 2023 doubles the amount in 2021, which suggests a heightened investment in federal procurement funding and large-scale projects.
@@
 
-**Buyer** is a character-type variable representing the federal government department or agency who initiate and award the contract.
-[@tbl-buyer] shows the buyer who has over 25 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors.
-The table highlights that Employment and Social Development Canada (ESDC) and National Defence account over 50% of Microsoft's total awards.
+**Buyer** is a character-type variable representing the federal government department or agency that initiates and awards the contract.
+[@tbl-buyer] shows the buyer who has over 25 contracts with Microsoft, which identifies key partners and understands the distribution of contracts across sectors.
+The table highlights that Employment and Social Development Canada (ESDC) and National Defence account for over 50% of Microsoft's total awards.
 
@@
 
-**PhaseDays** is a numeric variable demonstrates the days between AwardDate and StartDate.
-In [@fig-preparatory], there is a high concentration around 0 indicates rapid procurement processes or emergency software services.
+**PhaseDays** is a numeric variable that demonstrates the days between AwardDate and StartDate.
+In [@fig-preparatory], there is a high concentration of around 0 indicates rapid procurement processes or emergency software services.
 [@fig-preparatory] also shows a few outlying contracts with a mobilization period of over one year.
-These long mobilization periods reflects issues such as delays in the procurement process, a long negotiation period, or administrative approvals.
+These long mobilization periods reflect issues such as delays in the procurement process, a long negotiation period, or administrative approvals.
 
@@
 **ContractDays** is a numeric variable which is the days between $StartDate$ and $EndDate$.
-[@fig-contract] shows the distribution of performance period.
-The right skewed distribution in [@fig-contract] indicates that the majority of contract duration is within a year.
+[@fig-contract] shows the distribution of the performance period.
+The right-skewed distribution in [@fig-contract] indicates that the majority of contract duration is within a year.
 There is an over 5 years performance period in [@fig-contract] implies a massive project that requires extensive preparation.
@@
 The boxplot @fig-box-1 shows the Contract Duration across various Contract Types.
-As we can see in @fig-box-1, Computer Services and Consulting contracts tend to have longer durations.
+As we can see in @fig-box-1, Computer Services, and Consulting contracts tend to have longer durations.
 This suggests these contracts often involve more complex or ongoing services that require extended timelines.
-In contrast, License/Maintenance Fees and Software contracts show shorter durations, with tighter interquartile ranges and lower medians, indicating that these contracts are likely less complex and are typically limited to a defined time period.
+In contrast, License/Maintenance Fees and Software contracts show shorter durations, with tighter interquartile ranges and lower medians, indicating that these contracts are likely less complex and are typically limited to a defined period.
 
 The boxplot @fig-box-2 shows the Phase Duration for each Contract Type.
-@fig-box-2 highlights Consulting and Distributed Computing Environment have longer preparatory phases, with a higher median and a larger spread.
+@fig-box-2 highlights Consulting and Distributed Computing Environments have longer preparatory phases, with a higher median and a larger spread.
 This suggests these contracts require more time for negotiation, planning, or administrative procedures before the actual work begins.
@@
 The heatmap @fig-date shows the distribution of federal procurement contracts awarded to Microsoft over four-month periods from 2020 to 2024.
-The frequency tends to drop in the September–December period, it suggests the budgetary constraints or the fiscal year-end.
+The frequency tends to drop in the September–December period, which suggests the budgetary constraints or the fiscal year-end.
 A noticeable decline occurs in 2024 because the data only updates until September.
@@
 
-\newpage
-
 The scatter plot @fig-mobilization shows the relationship between the Award Date and the Start Date of federal procurement contracts awarded to Microsoft.
@@
     legend.text = element_text(size = 6),      # Adjust legend label size
-    legend.title = element_text(size = 8)    # Adjust legend title size
+    legend.title = element_text(size = 8),     # Adjust legend title size
+    axis.text.x = element_text(size = 6),      # Adjust x-axis text size
+    axis.text.y = element_text(size = 6),      # Adjust y-axis text size
     )  # This removes the legend
@@
 The scatter plot @fig-performance shows the diversity in contract durations across procurement categories.
-Most contracts ends within one year of their start dates, suggesting that most federal contracts are short-term.
+Most contracts end within one year of their start dates, suggesting that most federal contracts are short-term.
 A smaller subset of contracts extends beyond three years, typically associated with large-scale projects or ongoing support agreements.
-@fig-performance also highlights that the most contract ends in the year of 2026.
+@fig-performance also highlights that most contract ends in the year 2026.
 A few contracts with over 5 years stand out in 2021, suggesting that longer durations are prevalent in service-oriented categories.
@@
     legend.text = element_text(size = 6),      # Adjust legend label size
-    legend.title = element_text(size = 8)      # Adjust legend title size
+    legend.title = element_text(size = 8),     # Adjust legend title size
+    axis.text.x = element_text(size = 6),      # Adjust x-axis text size
+    axis.text.y = element_text(size = 6),      # Adjust y-axis text size
   )
@@
 #| eval: true
-#| tbl-cap: "Explanatory models of contract amount based on ContractType, Buyer, ContractDays and PhaseDays"
+#| tbl-cap: "Summary of explanatory models for predicting contract amounts."
 #| warning: false
@@
 
+The table compares two Bayesian models: the full model, `Bayesian Model_Full`, which includes all predictors (e.g., $Contract$, $Buyer$, $ContractDays$, and $PhaseDays$); and a simplified model, `Bayesian Model`, that adjusts for reduced predictors' complexity (e.g., $ContractType$, $BuyerCleaned$, $ContractDays$, and $PhaseDays$).
+Coefficients represent the estimated effect of each predictor on the contract amount, with credible intervals reflecting uncertainty.
+The results highlight the influence of contract-specific and buyer-specific variables on procurement amounts.
+
 The Full Bayesian Model has an $R^2$ of 0.289, indicating that approximately 28.9% of the variance in the contract amount is explained by the model.
@@
 The lower RMSE value suggests that the Full Bayesian Model is more accurate in predicting the total contract amount.
+
+More results about the Bayesian model can be found in @sec-model-details.
 
@@
 This reciprocal relationship could introduce bias in the coefficient estimates if not properly addressed.
-
-## Alternative Models Considered
-Linear regression model was evaluated in the paper.
-While the linear regression model was straightforward and interpretable, it lacked the flexibility to incorporate uncertainty in parameter estimates.
-The Bayesian regression model was chosen as a middle ground, balancing flexibility, interpretability, and the ability to quantify uncertainty.
-However, the limitations observed in model performance suggest that future work could benefit from integrating features of these alternative approaches.
-
-More results about these two models can be found in @sec-model-details.
 
@@
 
-The methodology employed in this study involves analyzing federal procurement contracts awarded to Microsoft, leveraging data from two primary sources: Public Services and Procurement Canada (PSPC) and the Treasury Board of Canada Secretariat (TBS).
-These two datasets provide information on government procurement, but they come with distinct limitations and reporting structures [@ProcurementSources].
+The methodology employed in this study focuses on analyzing federal procurement contracts awarded to Microsoft.
+The dataset used in this analysis comes from various Canadian government platforms that post solicitations for business opportunities, enabling companies to bid on government contracts.
+This data is gathered from multiple government sources and consolidated into a unified format to ensure consistency and comprehensiveness.
+
+The analysis relies on two primary data sources: Public Services and Procurement Canada (PSPC) and the Treasury Board of Canada Secretariat (TBS).
+Each dataset provides unique insights into the federal procurement process, but they come with distinct limitations that must be considered when interpreting the results [@ProcurementSources].
 
@@
 
-+ PSPC serves as the central procurement body for most federal government departments and is responsible for managing a significant portion of federal contracts.
-It maintains a public platform called CanadaBuys, where procurement information is published.
-However, due to migration from the older BuyAndSell platform and the incomplete nature of pre-2022 data, some variables, such as contract types and trade agreement information, are missing or inconsistently reported.
-The PSPC data is particularly useful for understanding contracts where the PSPC is the managing body, but only contracts above a certain dollar threshold are publicly available.
-
-+ TBS consolidates procurement data from all government departments and requires proactive publication under the Access to Information Act.
-The TBS dataset is more comprehensive in terms of the number of contracts, but it suffers from inconsistencies and gaps in the reported fields.
-TBS publishes data quarterly, and records before 2017 should not be considered complete.
-Unlike PSPC, TBS data includes the full details of contract awards but does not incorporate tender or solicitation data.
-
-## Data Integration and Challenges:
-
-There is an overlap between TBS and PSPC data, such that many contracts reported by PSPC are also included in the TBS dataset.
-However, the two datasets do not always align perfectly.
-Contracts reported by TBS may contain different details from the PSPC records.
-
-This mismatch is exacerbated by the lack of a consistent unique identifier that would allow for easy linking between the two datasets.
-While about 20% of PSPC awards have matching TBS records, many contracts do not, making it impossible to perfectly merge the two datasets.
-
-To address this, both datasets were kept separate in [@ProcurementSources].
-PSPC data was used primarily for its comprehensive tender information, while TBS data was relied upon for its broader coverage of contract awards, spanning more years and a greater volume of contracts.
-
-## Solicitation Methods and Contract Types:
+### Public Services and Procurement Canada (PSPC)
+
+PSPC serves as the central procurement body for most federal government departments, overseeing a significant portion of federal contracts.
+It manages the CanadaBuys platform, where procurement information is publicly published.
+However, due to the migration from the older BuyAndSell platform and incomplete records prior to 2022, some contract details, such as contract types and trade agreement information, may be missing or inconsistently reported.
+The PSPC dataset is valuable for understanding the procurement contracts managed by PSPC, though it only includes contracts above a certain dollar threshold.
+
+### Treasury Board of Canada Secretariat (TBS)
+
+TBS consolidates procurement data from all government departments, requiring proactive publication under the Access to Information Act.
+Unlike PSPC, the TBS dataset covers a larger number of contracts, including those from various departments and ministries.
+However, some records in this dataset, especially those prior to 2017, should not be considered complete, and there are inconsistencies in the fields reported.
+The TBS data provides full details of contract awards but lacks tender or solicitation information, limiting its use for certain types of analysis.
+
+## Data Integration and Challenges
+
+One of the key challenges of this study is the overlap between the PSPC and TBS datasets.
+Many contracts reported by PSPC are also present in the TBS dataset.
+However, these two datasets do not always match perfectly.
+Differences in contract details, such as award amounts, contract types, or buyer names, can make direct comparisons difficult.
+
+A significant issue is the lack of a consistent unique identifier across the two datasets, which would allow for easy linking of records.
+While approximately 20% of PSPC contracts have matching records in TBS, a large proportion do not, complicating the merging of the datasets.
+In this study, the two datasets were kept separate to preserve the integrity of each source.
+The PSPC data was used primarily for its comprehensive tender information, while the TBS dataset provided a broader scope of contracts, covering a longer time period and more contracts overall.
+
+## Solicitation Methods and Contract Types
 
@@
 
-Four competitive options:
-
-+ Open bidding: Any supplier can bid in response to an online solicitation.
-+ Selective tendering: Only some prequalified suppliers can bid.
-+ Limited tendering: Deviation from requirements of the relevant trade agreements (see Trade Agreements), still allowing bids.
-+ Traditional: Many suppliers can bid but not in response to a public, online solicitation; for example, responding to an email sent to suppliers by the government.
-
-Two non-competitive options:
-
-+ Advanced Contract Award Notice: Notice to suppliers that the buyer intends to award a contract to a pre-identified supplier, while accepting a Statement of Capabilities, not a bid, from challenging suppliers.
-+ Non-competitive: Blanket category for all other non-competitive processes.
+### Competitive Solicitation Methods
+
+1. Open bidding: Any supplier can bid in response to an online solicitation.
+2. Selective tendering: Only some prequalified suppliers can bid.
+3. Limited tendering: Deviation from requirements of the relevant trade agreements (see Trade Agreements), still allowing bids.
+4. Traditional: Many suppliers can bid but not in response to a public, online solicitation; for example, responding to an email sent to suppliers by the government.
+
+### Non-Competitive Solicitation Methods
+
+1.
+Advanced Contract Award Notice: Notice to suppliers that the buyer intends to award a contract to a pre-identified supplier, while accepting a Statement of Capabilities, not a bid, from challenging suppliers.
+2. Non-competitive: Blanket category for all other non-competitive processes.
 

### 2024-12-04 paper/paper.qmd
Final Rearrangement

@@
 
-```{r}
-#| label: fig-contract
-#| fig-cap: The histogram shows the distribution of performance period, which demonstrates the days between start and end date.
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# create plot for Distribution of Poll Scores by candidate
-ggplot(cleaned_data, aes(x = ContractDays)) +
-  geom_histogram(
-    bins = 15,
-    fill = "steelblue",
-    color = "black",
-    alpha = 0.7
-  ) +
-  labs(
-    x = "Contract Performance Phase (day)",
-    y = "Frequency"
-  ) +
-  theme_minimal()
-```
-
 The correlation matrix in [@fig-cor] visualizes the relationships between key numerical predictor variables and the outcome variable, $Amount$.
@@
 
-```{r}
-#| label: fig-cor
-#| fig-cap: Correlation matrix showing the relationships between numerical predictor variables and the outcome Amount variable
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# filter to numerical predictor variables
-numerical_data <- cleaned_data %>%
-  select(Amount, PreparatoryPhase, ContractDays)
-
-correlation_matrix <- cor(numerical_data, use = "complete.obs")
-
-# create correlation matrix
-ggcorrplot(correlation_matrix, lab = TRUE) + 
-  theme(
-    axis.text.x = element_text(size = 10),  # Adjust x-axis text size
-    axis.text.y = element_text(size = 10) # Adjust y-axis text size
-  )
-```
-More information about the data analysis can be found in @sec-result.
-
-\newpage
+More information about data analysis and figures are displayed in @sec-data-visualization.
 
@@
 
-## Data
-
-The boxplot @fig-box-1 shows the Contract Duration across various Contract Types.
-As we can see in @fig-box-1, Computer Services, and Consulting contracts tend to have longer durations.
-This suggests these contracts often involve more complex or ongoing services that require extended timelines.
-In contrast, License/Maintenance Fees and Software contracts show shorter durations, with tighter interquartile ranges and lower medians, indicating that these contracts are likely less complex and are typically limited to a defined period.
-
-The boxplot @fig-box-2 shows the Phase Duration for each Contract Type.
-@fig-box-2 highlights Consulting and Distributed Computing Environments have longer preparatory phases, with a higher median and a larger spread.
-This suggests these contracts require more time for negotiation, planning, or administrative procedures before the actual work begins.
-On the other hand, Computer Services and License/Maintenance Fees contracts exhibit shorter preparatory phases, as evidenced by the lower medians and narrower interquartile ranges in @fig-box-2.
-This suggests that these contracts are more straightforward and may require less preparation before implementation.
+The results are summarized in @tbl-modelresults, including R², Log-Likelihood (Log.Lik.), Expected Log Pointwise Predictive Density (ELPD), Leave-One-Out Information Criterion (LOOIC), Watanabe-Akaike Information Criterion (WAIC), and Root Mean Square Error (RMSE).
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+bayesian_model <-
+  readRDS(file = here::here("models/bayesian_model.rds"))
+```
+
+```{r}
+#| label: tbl-modelresults
+#| echo: false
+#| eval: true
+#| tbl-cap: "Summary of explanatory models for predicting contract amounts."
+#| warning: false
+#| message: false
+
+# Generate the model summary
+
+bayesian_model_full <- readRDS(here::here("models/bayesian_model_full.rds"))
+bayesian_model <- readRDS(here::here("models/bayesian_model.rds"))
+
+model_list <- list("Bayesian Model_Full" = bayesian_model_full, "Bayesian Model" = bayesian_model)
+
+modelsummary(
+  model_list,
+  output = "kableExtra", 
+  coef_map = c("ContractType" = "ContractType", "Buyer" = "Buyer", 
+               "ContractDays" = "ContractDays", "PhaseDays" = "PhaseDays")
+) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+The table compares two Bayesian models: the full model, `Bayesian Model_Full`, which includes all predictors (e.g., $Contract$, $Buyer$, $ContractDays$, and $PhaseDays$); and a simplified model, `Bayesian Model`, that adjusts for reduced predictors' complexity (e.g., $ContractType$, $BuyerCleaned$, $ContractDays$, and $PhaseDays$).
+Coefficients represent the estimated effect of each predictor on the contract amount, with credible intervals reflecting uncertainty.
+The results highlight the influence of contract-specific and buyer-specific variables on procurement amounts.
+
+The Full Bayesian Model has an $R^2$ of 0.289, indicating that approximately 28.9% of the variance in the contract amount is explained by the model.
+In comparison, the Reduced Bayesian Model explains only 8.2% of the variance in the contract amount, suggesting that the inclusion of additional predictors (specific buyers and contract names) improves the model’s explanatory power.
+However, both models have very low adjusted $R^2$, which suggests that the model is not capturing much of the unexplained variance in the data.
+
+The Full Bayesian Model has a Log-Likelihood of $-15,351.795$ and an ELPD of $-15,519.3$, while the Reduced Model has a Log-Likelihood of $-15,398.022$ and an ELPD of $-15,444.8$.
+These values suggest that the Full Bayesian Model fits the data slightly better than the Reduced Model, as it produces higher Log-Likelihood and ELPD values.
+
+The Full Bayesian Model has a lower LOOIC ($31,038.7$) and WAIC ($30,992.3$) compared to the Reduced Model, indicating that the Full Bayesian Model is better at out-of-sample predictive performance.
+Lower values of LOOIC and WAIC are indicative of better model fit and predictive accuracy.
+
+The Full Bayesian Model has a lower RMSE ($2,714,467.04$) compared to the Reduced Model ($2,971,135.35$), indicating better predictive performance.
+The lower RMSE value suggests that the Full Bayesian Model is more accurate in predicting the total contract amount.
+
+More results about the Bayesian model can be found in @sec-model-details.
+
+\newpage
+
+# Discussion {#sec-discussion}
+
+## Contract Complexity and Cost
+The positive association between $ContractDays$ and contract amount suggests that longer contracts tend to be more expensive.
+This pattern likely reflects the increased scope, complexity, and resource demands of these agreements, which are particularly prominent in IT-related contracts.
+These contracts often require ongoing development, maintenance, and support, making them more costly over time.
+Understanding this relationship helps stakeholders anticipate the financial requirements of long-term contracts and make more informed decisions when negotiating the terms of such agreements.
+
+## Negotiation and Efficiency
+The negative association between $PhaseDays$ and contract amount suggests that longer planning and negotiation periods may be associated with lower costs.
+This could be attributed to more thorough upfront planning, strategic negotiation, and budget management, which might lead to better cost control.
+For procurement agencies, this finding highlights the potential for improving procurement efficiency by streamlining the preparation phase.
+It also emphasizes the importance of early planning in securing cost-effective contracts, which can ultimately benefit both suppliers and government departments.
+
+## Sector-Specific Effects 
+The significant variation in contract amounts across different contract types (e.g., Consulting, Software) and buyers (e.g., National Defence, Employment and Social Development Canada) underscores the importance of sector-specific factors in federal procurement dynamics.
+Departments with higher budgets or those managing critical projects, such as National Defence, tend to award larger contracts.
+This observation suggests that the size and nature of the buyer's mandate play a substantial role in shaping procurement patterns.
+It also indicates that procurement decisions are not solely based on market competitiveness but are influenced by the strategic priorities of the departments involved.
+
+## Model Performance and Limitations {#sec-discussion-model}
+
+The Bayesian model used in this study has low adjusted $R^2$ values, indicating that it explains only a small portion of the variance in contract amounts (28.9%).
+This suggests that other important predictors, not captured by the model, may be influencing contract outcomes.
+The presence of missing or erroneous values in key variables like $ContractDays$ or $PhaseDays$ may introduce bias, further reducing the model's reliability.
+
+One notable limitation of the model is its inability to account for external factors that could significantly affect procurement decisions, such as political influences, departmental priorities, or budgetary constraints.
+These unobserved variables may introduce omitted variable bias, affecting the model's accuracy.
+Furthermore, the model assumes linear relationships between the independent variables and the dependent variable, which oversimplifies the complex nature of federal procurement.
+The categorization of contract types and buyers into broad groups also limits the model's ability to capture the full diversity of contracts awarded.
+
+Endogeneity is another issue that may affect the model's findings.
+There may be a bidirectional relationship between procurement amounts and buyer behavior—future procurement amounts could influence the allocation of future budgets, which in turn may impact future contracts awarded to Microsoft.
+This reciprocal relationship, if not properly addressed, can lead to biased coefficient estimates.
+
+## Weaknesses and Next Steps
+
+This paper provides a foundation for understanding federal procurement contracts awarded to Microsoft but also suggests several directions for future research.
+One potential extension of this study is to compare Microsoft's procurement trends with those of other major suppliers.
+By analyzing whether Microsoft's patterns are unique or consistent with broader trends across industries, researchers could better assess whether the observed dynamics are systemic within federal procurement practices.
+
+Additionally, the data quality issues identified—such as incomplete or inconsistent contract dates—highlight the need for improved data reporting standards.
+These inconsistencies could be addressed through collaboration with government agencies to ensure that future procurement datasets are more accurate and reliable.
+This would enhance the validity of future analyses and make procurement data more actionable for both policymakers and industry stakeholders.
+
+Finally, exploring more complex models, such as non-linear regression or machine learning approaches, could provide deeper insights into the factors influencing contract amounts.
+These methods may offer more flexibility in capturing non-linear relationships and uncovering hidden patterns within the data, improving the robustness and predictive power of the analysis.
+
+\newpage
+
+\appendix
+
+# Appendix {-}
+
+# Additional Data details {#sec-data-details}
+
+## Graphs {#sec-data-visualization}
+```{r}
+#| label: fig-contract
+#| fig-cap: The histogram shows the distribution of performance period, which demonstrates the days between start and end date.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# create plot for Distribution of Poll Scores by candidate
+ggplot(cleaned_data, aes(x = ContractDays)) +
+  geom_histogram(
+    bins = 15,
+    fill = "steelblue",
+    color = "black",
+    alpha = 0.7
+  ) +
+  labs(
+    x = "Contract Performance Phase (day)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+```
+
+```{r}
+#| label: fig-cor
+#| fig-cap: Correlation matrix showing the relationships between numerical predictor variables and the outcome Amount variable
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# filter to numerical predictor variables
+numerical_data <- cleaned_data %>%
+  select(Amount, PreparatoryPhase, ContractDays)
+
+correlation_matrix <- cor(numerical_data, use = "complete.obs")
+
+# create correlation matrix
+ggcorrplot(correlation_matrix, lab = TRUE) + 
+  theme(
+    axis.text.x = element_text(size = 10),  # Adjust x-axis text size
+    axis.text.y = element_text(size = 10) # Adjust y-axis text size
+  )
+```
 
@@
 ```
+
+\newpage
+
+## Data Visualization Analysis
+
+The boxplot @fig-box-1 shows the Contract Duration across various Contract Types.
+As we can see in @fig-box-1, Computer Services, and Consulting contracts tend to have longer durations.
+This suggests these contracts often involve more complex or ongoing services that require extended timelines.
+In contrast, License/Maintenance Fees and Software contracts show shorter durations, with tighter interquartile ranges and lower medians, indicating that these contracts are likely less complex and are typically limited to a defined period.
+
+The boxplot @fig-box-2 shows the Phase Duration for each Contract Type.
+@fig-box-2 highlights Consulting and Distributed Computing Environments have longer preparatory phases, with a higher median and a larger spread.
+This suggests these contracts require more time for negotiation, planning, or administrative procedures before the actual work begins.
+On the other hand, Computer Services and License/Maintenance Fees contracts exhibit shorter preparatory phases, as evidenced by the lower medians and narrower interquartile ranges in @fig-box-2.
+This suggests that these contracts are more straightforward and may require less preparation before implementation.
 
@@
 ```
-
-\newpage
-
-## Model Summary
-
-The results are summarized in @tbl-modelresults, including R², Log-Likelihood (Log.Lik.), Expected Log Pointwise Predictive Density (ELPD), Leave-One-Out Information Criterion (LOOIC), Watanabe-Akaike Information Criterion (WAIC), and Root Mean Square Error (RMSE).
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-bayesian_model <-
-  readRDS(file = here::here("models/bayesian_model.rds"))
-```
-
-```{r}
-#| label: tbl-modelresults
-#| echo: false
-#| eval: true
-#| tbl-cap: "Summary of explanatory models for predicting contract amounts."
-#| warning: false
-#| message: false
-
-# Generate the model summary
-
-bayesian_model_full <- readRDS(here::here("models/bayesian_model_full.rds"))
-bayesian_model <- readRDS(here::here("models/bayesian_model.rds"))
-
-model_list <- list("Bayesian Model_Full" = bayesian_model_full, "Bayesian Model" = bayesian_model)
-
-modelsummary(
-  model_list,
-  output = "kableExtra", 
-  coef_map = c("ContractType" = "ContractType", "Buyer" = "Buyer", 
-               "ContractDays" = "ContractDays", "PhaseDays" = "PhaseDays")
-) %>%
-  kable_styling(latex_options = c("striped", "scale_down"))
-```
-
-The table compares two Bayesian models: the full model, `Bayesian Model_Full`, which includes all predictors (e.g., $Contract$, $Buyer$, $ContractDays$, and $PhaseDays$); and a simplified model, `Bayesian Model`, that adjusts for reduced predictors' complexity (e.g., $ContractType$, $BuyerCleaned$, $ContractDays$, and $PhaseDays$).
-Coefficients represent the estimated effect of each predictor on the contract amount, with credible intervals reflecting uncertainty.
-The results highlight the influence of contract-specific and buyer-specific variables on procurement amounts.
-
-The Full Bayesian Model has an $R^2$ of 0.289, indicating that approximately 28.9% of the variance in the contract amount is explained by the model.
-In comparison, the Reduced Bayesian Model explains only 8.2% of the variance in the contract amount, suggesting that the inclusion of additional predictors (specific buyers and contract names) improves the model’s explanatory power.
-However, both models have very low adjusted $R^2$, which suggests that the model is not capturing much of the unexplained variance in the data.
-
-The Full Bayesian Model has a Log-Likelihood of $-15,351.795$ and an ELPD of $-15,519.3$, while the Reduced Model has a Log-Likelihood of $-15,398.022$ and an ELPD of $-15,444.8$.
-These values suggest that the Full Bayesian Model fits the data slightly better than the Reduced Model, as it produces higher Log-Likelihood and ELPD values.
-
-The Full Bayesian Model has a lower LOOIC ($31,038.7$) and WAIC ($30,992.3$) compared to the Reduced Model, indicating that the Full Bayesian Model is better at out-of-sample predictive performance.
-Lower values of LOOIC and WAIC are indicative of better model fit and predictive accuracy.
-
-The Full Bayesian Model has a lower RMSE ($2,714,467.04$) compared to the Reduced Model ($2,971,135.35$), indicating better predictive performance.
-The lower RMSE value suggests that the Full Bayesian Model is more accurate in predicting the total contract amount.
-
-More results about the Bayesian model can be found in @sec-model-details.
-
-\newpage
-
-# Discussion {#sec-discussion}
-
-## Contract Complexity and Cost
-The positive association between $ContractDays$ and contract amount suggests that longer contracts are typically more expensive, reflecting the increased scope, complexity, and resource requirements of these agreements.
-This is particularly relevant for IT-related contracts, which often involve extensive development, maintenance, and support services.
-
-## Negotiation and Efficiency
-The negative association between $PhaseDays$ and contract amount indicates that longer preparatory phases may lead to reduced costs, potentially due to more thorough planning, negotiation, or budgetary constraints.
-This insight can inform government strategies for managing procurement timelines to achieve cost efficiency.
-
-## Sector-Specific Effects 
-The significant variation in contract amounts across different contract types (e.g., Consulting, Software) and $buyers$ (e.g., National Defence, Employment and Social Development Canada) underscores the importance of sector-specific factors in procurement dynamics.
-Departments with higher budgets or critical mandates, such as National Defence, tend to award larger contracts, highlighting their influence in shaping overall procurement trends.
-
-## Model Performance and Limitations {#sec-discussion-model}
-
-The Bayesian model has low adjusted $R^2$ values.
-This means the model explains only a small portion (28.9%) of the variance in contract amounts.
-This suggests that important predictors may be missing.
-Also, missing or erroneous values in variables like $ContractDays$ or $PhaseDays$ could introduce bias and reduce the reliability of the results.
-
-One limitation of the model is that it may fail to account for certain external factors or unobserved variables such as political factors, budgetary constraints, and departmental priorities.
-These external factors could lead to omitted variable bias, which also influence procurement decision.
-
-The model assumes a linear relationship between the independent variables and the dependent variable, which oversimplify the dynamics of federal procurement.
-The model categorizes contract types and buyers into broad groups, which also oversimplify the true diversity of contracts awarded.
-Future studies could explore non-linear models, such as splines or machine learning approaches, to capture more complex patterns.
-
-There may be issues of endogeneity, where causality runs in both directions between some predictors and the outcome variable.
-The amount of procurement from a buyer may influence the allocation of future budgets to that department, which in turn affects future contract amounts awarded to Microsoft.
-This reciprocal relationship could introduce bias in the coefficient estimates if not properly addressed.
-
-## Weaknesses and Next Steps
-
-This paper provides a foundation for a deeper understanding of federal contract awards to Microsoft, but it also suggests several directions for future exploration.
-First, the study can expand to include comparisons with other major suppliers.
-By examining whether Microsoft's procurement trends are unique or reflective of broader patterns across industries.
-This gives a better understanding on assess whether certain dynamics are systemic within federal contracting practices.
-
-Data quality and consistency issues are also of concern.
-The anomalies identified in this study, such as incomplete or inconsistent contract dates, highlight the need for improved data reporting standards.
-These discrepancies can be addressed in collaboration with government agencies to improve the reliability of procurement datasets and ensure that future analyses are more reliable and actionable.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
 

### 2024-12-04 paper/paper.qmd
Final check

@@
 [@sec-discussion] then discusses potential limitations of data, model and suggestions for future research.
-The appendix provides additional detailed information about models in [@sec-model-details] and methodology in [@sec-meth-details].
+The appendix provides additional detailed information about data visualization in [@sec-data-details] models in [@sec-model-details] and methodology in [@sec-meth-details].
 
@@
 
-```{r}
-#| label: fig-amount
-#| fig-cap: The histogram shows the distribution of the total amount ($million) of each year's federal contract (since 2020).
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# Extract year from AwardDate and filter the year from 2020 to 2024
-year_data <- cleaned_data %>%
-  mutate(
-    Year = as.integer(format(as.Date(AwardDate), "%Y"))
-    )
-
-# Sum the total amount group by year
-yearly_amount <- year_data %>%
-  group_by(Year) %>%
-  summarize(TotalAmount = sum(Amount, na.rm = TRUE)) %>%
-  mutate(TotalAmount = round(TotalAmount / 1000000)) # Convert to millions
-
-# Plot the amount in each year
-ggplot(yearly_amount, aes(x = as.factor(Year), y = TotalAmount)) +
-  geom_bar(stat = "identity", fill = "darkorange", color = "black", alpha = 0.7) +
-  labs(
-    title = "Total Contract Amounts by Year for Microsoft",
-    x = "Year",
-    y = "Total Amount ($million)"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.y = element_text(size = 8), 
-    axis.title.y = element_text(size = 10),
-    axis.title.x = element_text(size = 10),
-    axis.text.x = element_text(size = 8),
-    plot.title = element_text(size = 12)
-    )
-
-```
-
-\newpage
-
 ## Predictor Variables
@@
 #| label: tbl-contract
-#| tbl-cap: The table summarizes the contract type which has over 25 contracts with the total dollar amount, and average dollar amount per contract. 
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# Summarize the data and filter buyers with counts >= 25
+#| tbl-cap: Summarizes the contract type which has over 30 contracts with the total dollar amount, and average dollar amount per contract. 
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Summarize the data and filter buyers with counts >= 30
 summary_table <- cleaned_data %>%
@@
   ) %>%
-  filter(Count > 25) %>%
+  filter(Count > 30) %>%
   arrange(desc(Count))  # Arrange in descending order based on Count
@@
 **Buyer** is a character-type variable representing the federal government department or agency that initiates and awards the contract.
-[@tbl-buyer] shows the buyer who has over 25 contracts with Microsoft, which identifies key partners and understands the distribution of contracts across sectors.
+[@tbl-buyer] shows the buyer who has over 30 contracts with Microsoft, which identifies key partners and understands the distribution of contracts across sectors.
 The table highlights that Employment and Social Development Canada (ESDC) and National Defence account for over 50% of Microsoft's total awards.
@@
 #| label: tbl-buyer
-#| tbl-cap: The table summarizes top 10 buyers with contract count, total dollar amount, and average dollar amount per contract. 
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# Summarize the data and filter buyers with counts >= 25
+#| tbl-cap: Summarizes the buyers who has over 30 contract counts, with total dollar amount and average dollar amount per contract. 
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Summarize the data and filter buyers with counts >= 30
 summary_table <- cleaned_data %>%
@@
 
-```{r}
-#| label: fig-preparatory
-#| fig-cap: The histogram shows the distribution of preparatory period.
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# create plot for Distribution of Poll Scores by candidate
-ggplot(cleaned_data, aes(x = PhaseDays)) +
-  geom_histogram(
-    bins = 30,
-    fill = "salmon",
-    color = "black",
-    alpha = 0.7
-  ) +
-  labs(
-    x = "Preparatory Phase (day)",
-    y = "Frequency"
-  ) +
-  theme_minimal()
-```
-
-\newpage
-
 **ContractDays** is a numeric variable which is the days between $StartDate$ and $EndDate$.
@@
 More information about data analysis and figures are displayed in @sec-data-visualization.
+
+\newpage
 
@@
 
+\newpage
+
 # Results {#sec-result}
+
+The table compares two Bayesian models: the full model, `Bayesian Model_Full`, which includes all predictors (e.g., $Contract$, $Buyer$, $ContractDays$, and $PhaseDays$); and a simplified model, `Bayesian Model`, that adjusts for reduced predictors' complexity (e.g., $ContractType$, $BuyerCleaned$, $ContractDays$, and $PhaseDays$).
+Coefficients represent the estimated effect of each predictor on the contract amount, with credible intervals reflecting uncertainty.
+The results highlight the influence of contract-specific and buyer-specific variables on procurement amounts.
 
@@
 
-The table compares two Bayesian models: the full model, `Bayesian Model_Full`, which includes all predictors (e.g., $Contract$, $Buyer$, $ContractDays$, and $PhaseDays$); and a simplified model, `Bayesian Model`, that adjusts for reduced predictors' complexity (e.g., $ContractType$, $BuyerCleaned$, $ContractDays$, and $PhaseDays$).
-Coefficients represent the estimated effect of each predictor on the contract amount, with credible intervals reflecting uncertainty.
-The results highlight the influence of contract-specific and buyer-specific variables on procurement amounts.
-
-The Full Bayesian Model has an $R^2$ of 0.289, indicating that approximately 28.9% of the variance in the contract amount is explained by the model.
-In comparison, the Reduced Bayesian Model explains only 8.2% of the variance in the contract amount, suggesting that the inclusion of additional predictors (specific buyers and contract names) improves the model’s explanatory power.
-However, both models have very low adjusted $R^2$, which suggests that the model is not capturing much of the unexplained variance in the data.
-
-The Full Bayesian Model has a Log-Likelihood of $-15,351.795$ and an ELPD of $-15,519.3$, while the Reduced Model has a Log-Likelihood of $-15,398.022$ and an ELPD of $-15,444.8$.
-These values suggest that the Full Bayesian Model fits the data slightly better than the Reduced Model, as it produces higher Log-Likelihood and ELPD values.
-
-The Full Bayesian Model has a lower LOOIC ($31,038.7$) and WAIC ($30,992.3$) compared to the Reduced Model, indicating that the Full Bayesian Model is better at out-of-sample predictive performance.
-Lower values of LOOIC and WAIC are indicative of better model fit and predictive accuracy.
-
-The Full Bayesian Model has a lower RMSE ($2,714,467.04$) compared to the Reduced Model ($2,971,135.35$), indicating better predictive performance.
-The lower RMSE value suggests that the Full Bayesian Model is more accurate in predicting the total contract amount.
+The Full Bayesian Model explains 28.9% of the variance in contract amounts, as indicated by the $R^2$ value of 0.289.
+In comparison, the Reduced Model explains only 8.2% of the variance, highlighting the contribution of additional predictors (e.g., specific buyers and contract names) in improving the model's explanatory power.
+However, both models show low adjusted $R^2$ values, suggesting that the models do not capture a large portion of the unexplained variance in the data.
+This points to the need for further refinement, potentially by incorporating additional factors that influence procurement decisions.
+
+The Full Bayesian Model has a Log-Likelihood of $-15,351.795$ and an ELPD of $-15,519.3$, while the Reduced Model's Log-Likelihood is $-15,398.022$ and ELPD is $-15,444.8$.
+The Full Model performs slightly better, as indicated by its higher Log-Likelihood and ELPD values.
+Both of these metrics suggest that the Full Model fits the data more accurately, as higher Log-Likelihood and ELPD values correspond to better data fit and prediction accuracy.
+
+The Full Bayesian Model has a lower LOOIC ($31,038.7$) and WAIC ($30,992.3$) compared to the Reduced Model, suggesting better out-of-sample predictive performance.
+LOOIC and WAIC are used to estimate the out-of-sample prediction accuracy, with lower values indicating better model performance.
+These results further support the conclusion that the Full Model is superior in predicting contract amounts.
+
+The Full Bayesian Model also outperforms the Reduced Model in terms of RMSE, with a value of $2,714,467.04$ compared to $2,971,135.35$ for the Reduced Model.
+The lower RMSE indicates that the Full Model is more accurate in predicting the total contract amount, making it a better tool for forecasting future contract values.
 
 More results about the Bayesian model can be found in @sec-model-details.
+This includes further exploration of model convergence, posterior predictive checks, and alternative specifications that could be explored in future research.
 
@@
 ## Graphs {#sec-data-visualization}
+
+```{r}
+#| label: fig-amount
+#| fig-cap: Total contract amounts awarded to Microsoft ($million) by year (since 2020). The data is grouped by award year, highlighting an increasing trends in Microsoft's federal contracts.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Extract year from AwardDate and filter the year from 2020 to 2024
+year_data <- cleaned_data %>%
+  mutate(
+    Year = as.integer(format(as.Date(AwardDate), "%Y"))
+    )
+
+# Sum the total amount group by year
+yearly_amount <- year_data %>%
+  group_by(Year) %>%
+  summarize(TotalAmount = sum(Amount, na.rm = TRUE)) %>%
+  mutate(TotalAmount = round(TotalAmount / 1000000)) # Convert to millions
+
+# Plot the amount in each year
+ggplot(yearly_amount, aes(x = as.factor(Year), y = TotalAmount)) +
+  geom_bar(stat = "identity", fill = "darkorange", color = "black", alpha = 0.7) +
+  labs(
+    title = "Total Contract Amounts by Year for Microsoft",
+    x = "Year",
+    y = "Total Amount ($million)"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.y = element_text(size = 8), 
+    axis.title.y = element_text(size = 10),
+    axis.title.x = element_text(size = 10),
+    axis.text.x = element_text(size = 8),
+    plot.title = element_text(size = 12)
+    )
+```
+
+```{r}
+#| label: fig-preparatory
+#| fig-cap: Distribution of preparatory period.
+This shows the frequency of contracts grouped by the number of days spent in the preparatory phase, illustrating the variability in preparatory phase durations across the dataset.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# create plot for Distribution of Poll Scores by candidate
+ggplot(cleaned_data, aes(x = PhaseDays)) +
+  geom_histogram(
+    bins = 30,
+    fill = "salmon",
+    color = "black",
+    alpha = 0.7
+  ) +
+  labs(
+    x = "Preparatory Phase (day)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+```
+
 ```{r}
 #| label: fig-contract
-#| fig-cap: The histogram shows the distribution of performance period, which demonstrates the days between start and end date.
+#| fig-cap: Distribution of contract performance phase durations.
+The histogram shows the frequency of contracts grouped by the number of days in the performance phase, which demonstrates the days between start and end date.
 #| echo: false
@@
 #| label: fig-cor
-#| fig-cap: Correlation matrix showing the relationships between numerical predictor variables and the outcome Amount variable
+#| fig-cap: Correlation matrix of numeric variables.
+This shows a weak negative correlations between $PreparatoryPhase$ and $ContractDays$; and weak positive correlations between predictors and $Amount$.
 #| echo: false
@@
 
+\newpage
+
+## Data Visualization Analysis
+
 ```{r}
 #| label: fig-box
-#| fig-cap: The Boxplots shows the relationship of ContractDays and PhaseDays with ContractType. 
+#| fig-cap: The relationship of ContractDays and PhaseDays by ContractType. 
 #| fig-subcap: ["Contract Duration in each Contract Type","Phase Duration in each Contract Type"]
@@
 
-\newpage
-
-## Data Visualization Analysis
-
 The boxplot @fig-box-1 shows the Contract Duration across various Contract Types.
@@
 #| label: fig-date
-#| fig-cap: Heatmap of Contracts by Four-Month Periods (2020–2024).
+#| fig-cap: The distribution of federal contracts by Four-Month Periods (2020–2024).
 The heatmap shows the frequency of federal procurement contracts awarded to Microsoft in each four-month period, with darker colors representing higher contract frequencies.
@@
 #| label: fig-mobilization
-#| fig-cap: The scatter plot shows the relationship between Award Date and Start Date for the top 10 federal procurement contracts types.
+#| fig-cap: The relationship between Award Date and Start Date for the top 10 federal procurement contracts types.
 Each point represents a contract, with colors indicating different contract types.
@@
 #| label: fig-performance
-#| fig-cap: The scatter plot shows the relationship between Start Date and End Date for the top 10 Microsoft’s federal procurement contracts types.
+#| fig-cap: The relationship between Start Date and End Date for the top 10 Microsoft’s federal procurement contracts types.
 Each point represents a contract, and the color reflects the contract type.
